### 2024-01-05 R FIle.qmd
adding files to repo

---
title: "testibng"
format: html
editor: visual
---

### Loading Packages

```{r}
library(tidytuesdayR)
library(tidyverse)
```

### Gathering Data

```{r}
# Union Data
tuesdata <- tidytuesdayR::tt_load(2023, week = 36)

demographics <- tuesdata$demographics
wages <- tuesdata$wages
states <- tuesdata$states

# Election Data
tuesdata <- tidytuesdayR::tt_load(2023, week = 45)

house <- tuesdata$house
```

### Cleaning Data

First, lets get an idea of what our data looks like.

```{r}
head(demographics)
head(wages)
head(states)
head(house)
```

From the head() function we can already see some tidying that needs to be done, mostly in the `house` table.

#### Items that need to be done

-   Sentence case

### EDA

```{r}
states |> 
  ggplot(aes(x = year, y = p_covered, color = state)) +
  geom_point()
```

```{r}
house <- house |>
  mutate(state = str_to_sentence(state))
```

```{r}
states |> 
  filter(sector == "Total") |> 
  left_join(house, by = c("year", "state")) |> 
  filter(party == "DEMOCRAT" | party == "REPUBLICAN") |> 
  filter(year > 2012) |> 
  ggplot(aes(x = party, y = p_members)) +
  geom_boxplot()
```

```{r}
test <- states |> 
  filter(sector == "Total") |> 
  left_join(house, by = c("year", "state")) |> 
  filter(party == "DEMOCRAT" | party == "REPUBLICAN") |> 
  filter(year > 2012)
```

```{r}
t.test(p_members ~ party, data = test, conf.level = .9)
```

### 2024-01-05 R FIle.qmd
added some EDA

@@
 From the head() function we can already see some tidying that needs to be done, mostly in the `house` table.
+First, to make data analysis easier we will need to combine the `states` and `house` table; this will be done on the year and state columns.
+however, the state column in the house data set is in uppercase, compared to sentence case in the states table.
+Next, R has read district in as a string column rather than a numeric column.
+
+Next we will look at a quick summary of the data to find very obvious trends.
+
+```{r}
+summary(demographics)
+```
+
+```{r}
+summary(wages)
+```
+```{r}
+summary(states)
+```
+```{r}
+summary(house)
+```
 
@@
 
--   Sentence case
+-   Sentence case to house table
+-   fix `district` in the house table
 

### 2024-01-06 R FIle.qmd
basic data cleaning

@@
 
+#### Examining data shape
+
 First, lets get an idea of what our data looks like.
@@
 Next, R has read district in as a string column rather than a numeric column.
+Examining the data set documentation, this is because 0 is actually "district at large" rather than a district number.
+This will be taken into account in data analysis.
 
@@
 ```
+From the summary, we can see the demographics has most of its variables skewed.
+This will be examined in closer detail in the exploratory data analysis.
+only two columns have NA values and each of those columns NA values represent less than 10% of the data.
+There is not an immediate cause to take action on these NA values, although they will be examined in greater detail in the EDA.
 
@@
 ```
+The `wages` table has no NA values. There is also nothing in the data documentation that would indicate a numeric value of some kind is being used as an NA value.
+
 ```{r}
@@
 ```
+Again, there are no NA values in the States data set, although we do see some significantly skewed data.
+
 ```{r}
@@
 
-#### Items that need to be done
+in the house data set, we can see a significant amount of NA values in the runoff column
 
--   Sentence case to house table
--   fix `district` in the house table
+#### Cleaning data script
+
+From a quick look at the shapes and distribution of the tables, the house table will be modified.
+
+```{r}
+house <- house |>
+  mutate(state = str_to_sentence(state)) |> # change from all caps to sentence case
+```
 

### 2024-01-06 R FIle.qmd
testing

@@
   mutate(state = str_to_sentence(state)) |>  # change from all caps to sentence case
+  mutate(regions = state.region[match(state_po, state.abb)]) # add regions 
+```
+
+#### Combining house and states dataframes
+The states data set has five entries for each year for each of the five `sectors`. For this project we will use the `total` sector. A decades column will also be added.
+
+```{r}
+full_data <- states |> 
+  filter(sector == "Total") |> 
+  left_join(house, by = c("year", "state")) |> 
+  filter(party == "DEMOCRAT" | party == "REPUBLICAN") |> 
+  mutate(decade = paste0((year - year %% 10), "s"),
+         party_binary = ifelse(party == "DEMOCRAT", 0, 1)) 
 ```
@@
 ```{r}
-states |> 
-  ggplot(aes(x = year, y = p_covered, color = state)) +
-  geom_point()
+full_data |> 
+  ggplot(aes(x = year, y = p_covered, color = regions)) +
+  geom_point() +
+  geom_smooth()
 ```
@@
 ```{r}
-house <- house |>
-  mutate(state = str_to_sentence(state))
+full_data |> 
+  ggplot(aes(x = party, y = p_members)) +
+  geom_boxplot() +
+  facet_wrap(vars(decade))
 ```
@@
 ```{r}
-states |> 
-  filter(sector == "Total") |> 
-  left_join(house, by = c("year", "state")) |> 
-  filter(party == "DEMOCRAT" | party == "REPUBLICAN") |> 
-  filter(year > 2012) |> 
-  ggplot(aes(x = party, y = p_members)) +
-  geom_boxplot()
+t.test(p_members ~ party, data = full_data, conf.level = .9)
+```
+```{r}
+full_data_after_2008 <- full_data |> 
+  filter(year > 2008)
+t.test(p_members ~ party, data = full_data_after_2008, conf.level = .9)
+```
+## Modeling
+```{r}
+model <- glm(party_binary ~ p_members, family = "binomial", data = full_data)
+summary(model)
 ```
 
-```{r}
-test <- states |> 
-  filter(sector == "Total") |> 
-  left_join(house, by = c("year", "state")) |> 
-  filter(party == "DEMOCRAT" | party == "REPUBLICAN") |> 
-  filter(year > 2012)
-```
-
-```{r}
-t.test(p_members ~ party, data = test, conf.level = .9)
-```
+## Conclusion

### 2024-01-07 Code/R FIle.qmd
fixed file path for wage data

--- 2024-01-06 R FIle.qmd
@@
 ---
+
+# useful git commands
+
+**This command will update all the branches on your computer** git remote update origin --prune
+
+showing off branch
+
+test
 
@@
 house <- tuesdata$house
+
+# Wage Data
+tuesdata <- tidytuesdayR::tt_load(2021, week = 7)
+
+income_distribution <- tuesdata$income_distribution
+
+# Minimum Wage Data
+minimum_wage <- read_csv("../Data/minimum_wage_data.csv")
+
+rm(tuesdata)
 ```
@@
 ```
+
 From the summary, we can see the demographics has most of its variables skewed.
@@
 ```
+
 The `wages` table has no NA values. There is also nothing in the data documentation that would indicate a numeric value of some kind is being used as an NA value.
@@
 ```
+
 Again, there are no NA values in the States data set, although we do see some significantly skewed data.
@@
 #### Combining house and states dataframes
+
 The states data set has five entries for each year for each of the five `sectors`. For this project we will use the `total` sector. A decades column will also be added.
@@
 ```
+
 ```{r}
@@
 ```
+
 ## Modeling
+
 ```{r}

### 2024-01-07 Code/R FIle.qmd
renamed data sets in code
saved data locally

@@
 # Union Data
-tuesdata <- tidytuesdayR::tt_load(2023, week = 36)
 
-demographics <- tuesdata$demographics
-wages <- tuesdata$wages
-states <- tuesdata$states
+union_demographics <- read_rds("../Data/union_demographics.rds")
+union_wages <- read_rds("../Data/union_wages.rds")
+union_states <- read_rds("../Data/union_states.rds")
 
 # Election Data
-tuesdata <- tidytuesdayR::tt_load(2023, week = 45)
-
-house <- tuesdata$house
+house <- read_rds("../Data/house.rds")
 
@@
 ```{r}
-full_data <- states |> 
+full_data <- union_states |> 
   filter(sector == "Total") |> 
@@
   mutate(decade = paste0((year - year %% 10), "s"),
-         party_binary = ifelse(party == "DEMOCRAT", 0, 1)) 
+         party_binary = ifelse(party == "DEMOCRAT", 0, 1))  |> 
+  seelect 
 ```

### 2024-01-07 Code/R FIle.qmd
fxed qmd
cleaned data

@@
 ```{r}
-library(tidytuesdayR)
 library(tidyverse)
@@
 # Wage Data
-tuesdata <- tidytuesdayR::tt_load(2021, week = 7)
-
-income_distribution <- tuesdata$income_distribution
+income_distribution <- read_rds("../Data/income_distribution.rds")
 
@@
 minimum_wage <- read_csv("../Data/minimum_wage_data.csv")
-
-rm(tuesdata)
 ```
@@
 ```{r}
-head(demographics)
-head(wages)
-head(states)
+head(union_demographics)
+head(union_wages)
+head(union_states)
 head(house)
+head(income_distribution)
+head(minimum_wage)
 ```
@@
 house <- house |>
-  mutate(state = str_to_sentence(state)) |>  # change from all caps to sentence case
-  mutate(regions = state.region[match(state_po, state.abb)]) # add regions 
+  mutate(state = str_to_sentence(state)) |> # change from all caps to sentence case) 
+  select(!(state_fips:office)) |> #Remove unwanted columns
+  select(!(stage:candidate)) |> 
+  select(!(writein:fusion_ticket))
+```
+
+```{r}
+union_demographics <- union_demographics |> 
+  select(-sample_size) |> 
+  mutate(union_type = facet, .keep = "unused")
+```
+
+```{r}
+union_wages <- union_wages |> 
+  select(-sample_size,
+         -at_cap) |> 
+  mutate(percent_difference = union_wage_premium_raw, .keep = "unused") |> 
+  mutate(percent_difference_adjusted = union_wage_premium_adjusted, .keep = "unused") |> 
+  mutate(union_type = facet, .keep = "unused")
+```
+
+```{r}
+union_states <- union_states |> 
+  select(-state_census_code,
+         -observations)
+```
+
+```{r}
+income_distribution <- income_distribution |> 
+  select(-number) |> 
+  select(!ends_with("moe")) |> 
+  select(year:income_mean) |> 
+  distinct(year, race, income_median, income_mean, .keep_all = TRUE) 
+```
+
+```{r}
+minimum_wage <- minimum_wage |> 
+  select(Year:CPI.Average)
 ```
@@
 The states data set has five entries for each year for each of the five `sectors`. For this project we will use the `total` sector. A decades column will also be added.
+
+```{r}
+test <- minimum_wage
+```
 

### 2024-01-07 Code/R FIle.qmd
better cleaned data; renamed variables

@@
 # Union Data
-
-union_demographics <- read_rds("../Data/union_demographics.rds")
 union_wages <- read_rds("../Data/union_wages.rds")
@@
 ```{r}
-head(union_demographics)
 head(union_wages)
@@
 Next we will look at a quick summary of the data to find very obvious trends.
-
-```{r}
-summary(demographics)
-```
 
@@
 ```{r}
-union_demographics <- union_demographics |> 
-  select(-sample_size) |> 
-  mutate(union_type = facet, .keep = "unused")
-```
-
-```{r}
 union_wages <- union_wages |> 
@@
          -at_cap) |> 
-  mutate(percent_difference = union_wage_premium_raw, .keep = "unused") |> 
-  mutate(percent_difference_adjusted = union_wage_premium_adjusted, .keep = "unused") |> 
+  mutate(wage_percent_difference = union_wage_premium_raw, .keep = "unused") |> 
+  mutate(wage_percent_difference_adjusted = union_wage_premium_adjusted, .keep = "unused") |> 
   mutate(union_type = facet, .keep = "unused")
@@
   select(-state_census_code,
-         -observations)
+         -observations) |> 
+  mutate(employment_count = employment * 1000, .keep = "unused") |> 
+  mutate(union_employment_count = members * 1000, .keep = "unused") |> 
+  mutate(collective_bargaion_employment_count = covered * 1000, .keep = "unused") |> 
+  rename(
+    percent_union = p_members,
+    percent_collective_bargain = p_covered
+  )
 ```
@@
   select(year:income_mean) |> 
-  distinct(year, race, income_median, income_mean, .keep_all = TRUE) 
+  distinct(year, race, income_median, income_mean, .keep_all = TRUE) |> 
+  rename(median_income = income_median,
+         mean_income = income_mean)
 ```
@@
 minimum_wage <- minimum_wage |> 
-  select(Year:CPI.Average)
+  select(Year:CPI.Average) |> 
+  rename(year = Year,
+         state = State,
+         state_min_wage = State.Minimum.Wage,
+         state_min_wage_2020_adj = State.Minimum.Wage.2020.Dollars,
+         fed_min_wage = Federal.Minimum.Wage,
+         fed_min_wage_2020_adj = Federal.Minimum.Wage.2020.Dollars,
+         effective_min_wage = Effective.Minimum.Wage,
+         effective_min_wage_2020_adj = Effective.Minimum.Wage.2020.Dollars,
+         CPI_avg = CPI.Average
+  )
 ```

### 2024-01-07 Code/R FIle.qmd
combined state data

@@
 ```{r}
-summary(wages)
+summary(union_wages)
 ```
@@
 ```{r}
-summary(states)
+summary(union_states)
 ```
@@
 house <- house |>
+  group_by(year, state, district) |> 
+  filter(candidatevotes == max(candidatevotes)) |> 
+  ungroup() |> 
   mutate(state = str_to_sentence(state)) |> # change from all caps to sentence case) 
@@
   select(!(stage:candidate)) |> 
-  select(!(writein:fusion_ticket)) 
+  select(!(writein:fusion_ticket)) |> 
+  rename(state_abbreviation = state_po)
 ```
@@
   mutate(union_employment_count = members * 1000, .keep = "unused") |> 
-  mutate(collective_bargaion_employment_count = covered * 1000, .keep = "unused") |> 
+  mutate(collective_bargain_employment_count = covered * 1000, .keep = "unused") |> 
   rename(
@@
 
-#### Combining house and states dataframes
+```{r}
+head(union_wages)
+head(union_states)
+head(house)
+head(income_distribution)
+head(minimum_wage)
+```
+```{r}
+range(union_wages$year) #ntl
+range(union_states$year) # state
+range(house$year) # state
+range(income_distribution$year) # ntl
+range(minimum_wage$year) # state
+```
+
+### Combining dataframes
 
@@
 
+#### State Dataframes
 ```{r}
-test <- minimum_wage
+full_state_data <- union_states |> 
+  filter(sector == "Total") |> 
+  select(-sector,
+         -state_abbreviation) |> 
+  left_join(house, by = c("year", "state")) |> 
+  filter(party == "DEMOCRAT" | party == "REPUBLICAN") |> 
+  mutate(decade = paste0((year - year %% 10), "s"),
+         party_binary = ifelse(party == "DEMOCRAT", 1, 0)) |> 
+  left_join(minimum_wage, by = c("year", "state")) |> 
+  group_by(year, state) |> 
+  mutate(party_state_avaerage = mean(party_binary))
 ```
@@
 ```{r}
-full_data <- union_states |> 
-  filter(sector == "Total") |> 
-  left_join(house, by = c("year", "state")) |> 
-  filter(party == "DEMOCRAT" | party == "REPUBLICAN") |> 
-  mutate(decade = paste0((year - year %% 10), "s"),
-         party_binary = ifelse(party == "DEMOCRAT", 0, 1))  |> 
-  seelect 
+state_column_list = c("year", "decade", "state", "state_abbreviation",
+                      "regions", "district", "party", "party_binary",
+                      "employment_count", "union_employment_count",
+                      "collective_bargain_employment_count", "percent_union",
+                      "percent_collective_bargain", "state_min_wage",
+                      "state_min_wage_2020_adj", "fed_min_wage",
+                      "fed_min_wage_2020_adj", "effective_min_wage",
+                      "effective_min_wage_2020_adj", "CPI_avg")
+full_state_data <- full_state_data[, state_column_list]
+```
+
+```{r}
+range(full_state_data$year)
 ```
@@
 
-```{r}
-full_data |> 
-  ggplot(aes(x = year, y = p_covered, color = regions)) +
-  geom_point() +
-  geom_smooth()
-```
-
-```{r}
-full_data |> 
-  ggplot(aes(x = party, y = p_members)) +
-  geom_boxplot() +
-  facet_wrap(vars(decade))
-```
-
-```{r}
-t.test(p_members ~ party, data = full_data, conf.level = .9)
-```
-
-```{r}
-full_data_after_2008 <- full_data |> 
-  filter(year > 2008)
-t.test(p_members ~ party, data = full_data_after_2008, conf.level = .9)
-```
-
 ## Modeling
 
-```{r}
-model <- glm(party_binary ~ p_members, family = "binomial", data = full_data)
-summary(model)
-```
-
 ## Conclusion

### 2024-01-07 Code/R FIle.qmd

@@
 union_wages <- union_wages |> 
+  filter(facet == "all wage and salary workers") |> 
+  distinct(year, .keep_all = TRUE) |> 
   select(-sample_size,
@@
 
+#### National Dataframes
+
+```{r}
+full_ntl_data <- income_distribution |> 
+  left_join(union_wages, by = "year")
+```
+```{r}
+range(full_ntl_data$year)
+```
+
 ### EDA

### 2024-01-08 Code/R FIle.qmd

@@
   select(-sector,
-         -state_abbreviation) |> 
-  left_join(house, by = c("year", "state")) |> 
+         -state) |> 
+  left_join(house, by = c("year", "state_abbreviation")) |> 
   filter(party == "DEMOCRAT" | party == "REPUBLICAN") |> 
@@
   group_by(year, state) |> 
-  mutate(party_state_avaerage = mean(party_binary))
+  mutate(party_state_avaerage = mean(party_binary)) |>
+  ungroup()
 ```
@@
 
+```{r}
+full_state_data %>% 
+  ungroup %>% 
+  count(year)
+```
+
 #### National Dataframes
+
+##### Creating Adjusted Wages Columns
+
+```{r}
+state_to_ntl <- full_state_data |>
+  ungroup() %>% 
+  mutate(adjusted_wage_prop = fed_min_wage / fed_min_wage_2020_adj, .after = year) |>
+  distinct(year, .keep = TRUE) 
+#  select(year:adjusted_wage_prop)
+```
 
@@
 full_ntl_data <- income_distribution |> 
-  left_join(union_wages, by = "year")
+  left_join(union_wages, by = "year") |>
+  mutate(decade = paste0((year - year %% 10), "s"), .after = year)
 ```
+
 ```{r}
@@
 
+```{r}
+summary(full_ntl_data)
+summary(full_state_data)
+```
+
+#### Graphing Full State Data
+
 ## Modeling

### 2024-01-08 Code/R FIle.qmd
fixed header

@@
 ---
-title: "testibng"
+title: "WIDS 2024 Final Project"
+author: "Michael Taffe, Melanie Klein, Ramiro Arevalo"
 format: html
@@
 #### State Dataframes
+
 ```{r}

### 2024-01-08 Code/R FIle.qmd
added more code comments, recoded party column in house data, moved filter for party from full_state_data to house clean script

@@
 ```{r}
+# look at columns
 head(union_wages)
@@
 ```{r}
+# look for NAs and skews
 summary(union_wages)
@@
 house <- house |>
-  group_by(year, state, district) |> 
-  filter(candidatevotes == max(candidatevotes)) |> 
+  group_by(year, state, district) |> # group for next line
+  filter(candidatevotes == max(candidatevotes)) |> #only include candidates that won
   ungroup() |> 
@@
   select(!(writein:fusion_ticket)) |> 
-  rename(state_abbreviation = state_po)
+  rename(state_abbreviation = state_po) |>  # rename state abbreviation column
+  mutate(party = case_when(
+    str_detect(party, "DEMOCRAT") ~ "DEMOCRAT",
+    str_detect(party, "REPUBLICAN") ~ "REUBLICAN",
+    str_detect(party, "INDEPENDENT") ~ "INDEPENDENT"
+  )) # Recode of party variable
 ```
@@
   select(-state_census_code,
-         -observations) |> 
-  mutate(employment_count = employment * 1000, .keep = "unused") |> 
+         -observations) |> # remove unwanted columns
+  mutate(employment_count = employment * 1000, .keep = "unused") |> # column measured in 1000s, this fixes it
   mutate(union_employment_count = members * 1000, .keep = "unused") |> 
-  mutate(collective_bargain_employment_count = covered * 1000, .keep = "unused") |> 
+  mutate(collective_bargain_employment_count = covered * 1000,
+         .keep = "unused") |> 
   rename(
@@
     percent_collective_bargain = p_covered
-  )
+  ) #renamed columns
 ```
@@
 income_distribution <- income_distribution |> 
-  select(-number) |> 
+  select(-number) |> # removing unwanted columns
   select(!ends_with("moe")) |> 
   select(year:income_mean) |> 
-  distinct(year, race, income_median, income_mean, .keep_all = TRUE) |> 
+  distinct(year, race, income_median, income_mean, .keep_all = TRUE) |> # only include one line per year. needed bc of multiple sectors that are not being used in data analysis
   rename(median_income = income_median,
-         mean_income = income_mean)
+         mean_income = income_mean) # renaming columns
 ```
@@
 minimum_wage <- minimum_wage |> 
-  select(Year:CPI.Average) |> 
+  select(Year:CPI.Average) |> # selecting relevant columns
   rename(year = Year,
@@
          CPI_avg = CPI.Average
-  )
+  ) # renaming variables
 ```
@@
 ```{r}
+# seeing what we have left
 head(union_wages)
@@
 ```{r}
+# seeing what years we have
 range(union_wages$year) #ntl
@@
   left_join(house, by = c("year", "state_abbreviation")) |> 
-  filter(party == "DEMOCRAT" | party == "REPUBLICAN") |> 
   mutate(decade = paste0((year - year %% 10), "s"),
@@
   ungroup()
+```
+
+```{r}
+union_states |> 
+  filter(sector == "Total") |> 
+  select(-sector,
+         -state) |> 
+  left_join(house, by = c("year", "state_abbreviation")) |> 
+  count(year)
 ```
@@
 full_state_data <- full_state_data[, state_column_list]
+```
+
+```{r}
+union_states |> 
+  filter(sector == "Total") |> 
+  select(-sector,
+         -state) |> 
+  left_join(house, by = c("year", "state_abbreviation")) |> 
+  filter(party == "DEMOCRAT" | party == "REPUBLICAN" | party == "FOGLIETTA (DEMOCRAT)") |>
+  count(year)
 ```

### 2024-01-08 Code/R FIle.qmd

@@
          CPI_avg = CPI.Average
-  ) # renaming variables
+  ) |> # renaming variables
+  filter(!state %in% c("U.S. Virgin Islands", "Puerto Rico", "Guam")) |>  # remove territory
+  mutate(state_abbreviation = 
+           ifelse(is.na(state.abb[match(state, state.name)]),
+                  "DC",
+                  state.abb[match(state, state.name)]),
+         .keep = "unused", .after = year) # replaces state with state abbreviation, NA values coded as DC as DC is the only NA value 
 ```
@@
          party_binary = ifelse(party == "DEMOCRAT", 1, 0)) |> 
-  left_join(minimum_wage, by = c("year", "state")) |> 
+  left_join(minimum_wage, by = c("year", "state_abbreviation")) |> 
   group_by(year, state) |> 
   mutate(party_state_avaerage = mean(party_binary)) |>
-  ungroup()
-```
-
-```{r}
-union_states |> 
-  filter(sector == "Total") |> 
-  select(-sector,
-         -state) |> 
-  left_join(house, by = c("year", "state_abbreviation")) |> 
-  count(year)
+  ungroup() |> 
+  mutate(state = state.name[match(state_abbreviation, state.abb)], # fix state encoding
+         regions = state.region[match(state_abbreviation, state.abb)]) # fix region encoding
 ```
@@
                       "effective_min_wage_2020_adj", "CPI_avg")
-full_state_data <- full_state_data[, state_column_list]
-```
-
-```{r}
-union_states |> 
-  filter(sector == "Total") |> 
-  select(-sector,
-         -state) |> 
-  left_join(house, by = c("year", "state_abbreviation")) |> 
-  filter(party == "DEMOCRAT" | party == "REPUBLICAN" | party == "FOGLIETTA (DEMOCRAT)") |>
-  count(year)
+full_state_data <- full_state_data[, state_column_list] # reorder columns
+full_state_data$party[is.na(full_state_data$party)] <- "No election" # replace party NAs
+full_state_data$district[is.na(full_state_data$district)] <- "No election" # replace district NAs
+full_state_data$party_binary[is.na(full_state_data$party_binary)] <- "No election" # replace district NAs
 ```
@@
 ```{r}
-full_state_data %>% 
-  ungroup %>% 
-  count(year)
+summary(full_state_data)
 ```

### 2024-01-08 Code/R FIle.qmd
added names for state specific exploratory graphs

@@
 #### Graphing Full State Data
+##### For facets, use best judgement if by state, decade, year, or region
+##### maybe make all states light grey and just choose outline outlier states and maybe regional averages
+
+##### Party by state and year
+##### Min Wage by state and year (include fed (facet or use by decades where necessary)
+##### Adjusted Min Wage by state and year (include adjusted fed (facet or use by decades where necessary)
+##### Union employment, CB employment, and full employment (facet by decade)
+##### percent union and percent collective bargaining
 

### 2024-01-09 Code/R FIle.qmd
started graphs for states

--- 2024-01-08 Code/R FIle.qmd
@@
 #### Graphing Full State Data
+##### For facets, use best judgement if by state, decade, year, or region
+##### maybe make all states light grey and just choose outline outlier states and maybe regional averages
+
+##### Party by state and year
+```{r}
+ggplot(full_state_data, aes(x = state_abbreviation, y = )) +
+```
+
+##### Min Wage by state and year (include fed (facet or use by decades where necessary)
+# figure out spacing of graph
+# can't do a line for federal minimum wage unless faceted by year instead of decade
+```{r}
+# abbreviation and decade
+ggplot(full_state_data, aes(x = state_abbreviation)) +
+  geom_point(aes(y = state_min_wage)) +
+  geom_hline(aes(yintercept = fed_min_wage), color = "red") +
+  facet_wrap(~decade, nrow = 5) +
+  theme(axis.text.x = element_text(angle=90)) +
+  labs(x = "State", 
+       y = "State Minimum Wage")
+
+# region and decade
+ggplot(full_state_data, aes(x = regions)) +
+  geom_point(aes(y = state_min_wage)) +
+  geom_hline(aes(yintercept = fed_min_wage), color = "red") +
+  facet_wrap(~decade, nrow = 5) +
+  labs(x = "State", 
+       y = "State Minimum Wage")
+
+# region and year
+ggplot(full_state_data, aes(x = regions)) +
+  geom_point(aes(y = state_min_wage)) +
+  geom_hline(aes(yintercept = fed_min_wage), color = "red") +
+  facet_wrap(~year, nrow = 10) +
+  labs(x = "State", 
+       y = "State Minimum Wage")
+```
+
+##### Adjusted Min Wage by state and year (include adjusted fed (facet or use by decades where necessary)
+# figure out spacing of graph
+# can't do a line for adj federal minimum wage unless faceted by year instead of decade
+# add a label for the line saying federal minimum wage
+```{r}
+# abbreviation and decade
+ggplot(full_state_data, aes(x = state_abbreviation)) +
+  geom_point(aes(y = state_min_wage_2020_adj)) +
+  geom_hline(aes(yintercept = fed_min_wage_2020_adj), color = "red") +
+  facet_wrap(~decade, nrow = 5) +
+  theme(axis.text.x = element_text(angle=90)) +
+  labs(x = "State", 
+       y = "State Minimum Wage in 2020 Dollars")
+
+# region and decade
+ggplot(full_state_data, aes(x = regions)) +
+  geom_point(aes(y = state_min_wage_2020_adj)) +
+  geom_hline(aes(yintercept = fed_min_wage_2020_adj), color = "red") +
+  facet_wrap(~decade, nrow = 5) +
+  labs(x = "Region", 
+       y = "State Minimum Wage in 2020 Dollars")
+
+# region and year
+ggplot(full_state_data, aes(x = regions)) +
+  geom_point(aes(y = state_min_wage_2020_adj)) +
+  geom_hline(aes(yintercept = fed_min_wage_2020_adj), color = "red") +
+  facet_wrap(~year, nrow = 10) +
+  labs(x = "Region", 
+       y = "State Minimum Wage in 2020 Dollars")
+```
+
+##### Union employment, CB employment, and full employment (facet by decade)
+# should we take the average value per decade?
+# add labels for each type of employment
+# fix spacing because union employment count is not visible
+```{r}
+# state and decade
+ggplot(full_state_data, aes(x = state_abbreviation)) +
+  geom_point(aes(y = employment_count), color = "red") +
+  geom_point(aes(y = union_employment_count), color = "blue") +
+  geom_point(aes(y = collective_bargain_employment_count), color = "green") +
+  facet_wrap(~decade, nrow = 5) +
+  labs(x = "State", 
+       y = "Number of People Employed")
+
+# region and decade
+ggplot(full_state_data, aes(x = regions)) +
+  geom_point(aes(y = employment_count), color = "red") +
+  geom_point(aes(y = union_employment_count), color = "blue") +
+  geom_point(aes(y = collective_bargain_employment_count), color = "green") +
+  facet_wrap(~decade, nrow = 5) +
+  labs(x = "Region", 
+       y = "Number of People Employed")
+```
+
+##### percent union and percent collective bargaining
+```{r}
+
+```
 

### 2024-01-09 Code/R FIle.qmd
state graphs

@@
 # should we take the average value per decade?
-# add labels for each type of employment
+# add a key for the color of the points
 # fix spacing because union employment count is not visible
@@
 ##### percent union and percent collective bargaining
+# should we take the average value per decade?
+# add a key for the color of the points
 ```{r}
+# state and decade
+ggplot(full_state_data, aes(x = state_abbreviation)) +
+  geom_point(aes(y = percent_union), color = "red") +
+  geom_point(aes(y = percent_collective_bargain), color = "blue") +
+  facet_wrap(~decade, nrow = 5) +
+  labs(x = "State", 
+       y = "Number of People Employed")
 
+# region and decade
+ggplot(full_state_data, aes(x = regions)) +
+  geom_point(aes(y = percent_union), color = "red") +
+  geom_point(aes(y = percent_collective_bargain), color = "blue") +
+  facet_wrap(~decade, nrow = 5) +
+  labs(x = "Region", 
+       y = "Number of People Employed")
 ```

### 2024-01-09 Code/R FIle.qmd
effective minimum wage graph

--- 2024-01-08 Code/R FIle.qmd
@@
 #### Graphing Full State Data
+
 ##### For facets, use best judgement if by state, decade, year, or region
+
 ##### maybe make all states light grey and just choose outline outlier states and maybe regional averages
@@
 ##### Party by state and year
+
+```{r}
+ggplot(full_state_data, aes(x = state_abbreviation, y = )) +
+```
+
 ##### Min Wage by state and year (include fed (facet or use by decades where necessary)
+
+# figure out spacing of graph
+
+# can't do a line for federal minimum wage unless faceted by year instead of decade
+
+```{r}
+effective_min_wage_adj_graph1 <- full_state_data |> 
+  filter(!is.na(regions)) |> 
+  filter(effective_min_wage_2020_adj != fed_min_wage_2020_adj) |> 
+  filter(state_abbreviation != "HI") |> 
+  ggplot(aes(x = year)) +
+  geom_line(aes(y = effective_min_wage_2020_adj, color = state_abbreviation)) +
+  scale_color_manual(values = c(rep("gray", 34))) +
+  geom_line(aes(y = fed_min_wage), color = "red", linewidth = .75)
+
+HI_effective_min_wage <- full_state_data |> 
+  filter(!is.na(regions)) |> 
+  filter(effective_min_wage_2020_adj != fed_min_wage_2020_adj) |> 
+  filter(state_abbreviation == "HI") |> 
+  mutate(effective_min_wage_2020_adj = effective_min_wage_2020_adj - .05)
+
+CT_effective_min_wage <- full_state_data |> 
+  filter(!is.na(regions)) |> 
+  filter(effective_min_wage_2020_adj != fed_min_wage_2020_adj) |> 
+  filter(state_abbreviation == "CT")
+
+AK_effective_min_wage <- full_state_data |> 
+  filter(!is.na(regions)) |> 
+  filter(effective_min_wage_2020_adj != fed_min_wage_2020_adj) |> 
+  filter(state_abbreviation == "AK")
+
+CA_effective_min_wage <- full_state_data |> 
+  filter(!is.na(regions)) |> 
+  filter(effective_min_wage_2020_adj != fed_min_wage_2020_adj) |> 
+  filter(state_abbreviation == "CA")
+
+effective_min_wage_adj_graph1 +
+  geom_line(data = HI_effective_min_wage, aes(y = effective_min_wage_2020_adj), color = "blue",linewidth = .7) +
+  annotate("text", x=1994, y=9.5, label= "Hawaii", color = "blue") +
+  geom_line(data = CT_effective_min_wage, aes(y = effective_min_wage_2020_adj), color = "#288225", linewidth = .7) +
+  annotate("text", x=1985, y=7.5, label= "Connecticut", color = "#288225") +
+  geom_line(data = AK_effective_min_wage, aes(y = effective_min_wage_2020_adj), color = "orange", linewidth = .7) +
+  annotate("text", x=1985, y=10, label= "Alaska", color = "orange") +
+  geom_line(data = CA_effective_min_wage, aes(y = effective_min_wage_2020_adj), color = "purple", linewidth = .7) +
+  annotate("text", x=2016, y=12, label= "California", color = "purple") +
+  annotate("text", x=2000, y=5.5, label= "Federal", color = "red") +
+  theme_minimal() +
+  theme(legend.position = "none") +
+  labs(title = "State effective minimum wage by Year (adjusted for 2020 $)",
+       subtitle = "For states with an adjusted effective minimum wage above the federal level",
+       x = "Year",
+       y = "Effective minimum wage (2020 USD)")
+
+rm(HI_effective_min_wage, CT_effective_min_wage,
+   AK_effective_min_wage, CA_effective_min_wage)
+```
+
+```{r}
+# abbreviation and decade
+ggplot(full_state_data, aes(x = state_abbreviation)) +
+  geom_point(aes(y = state_min_wage)) +
+  geom_hline(aes(yintercept = fed_min_wage), color = "red") +
+  facet_wrap(~decade, nrow = 5) +
+  theme(axis.text.x = element_text(angle=90)) +
+  labs(x = "State", 
+       y = "State Minimum Wage")
+
+# region and decade
+ggplot(full_state_data, aes(x = regions)) +
+  geom_point(aes(y = state_min_wage)) +
+  geom_hline(aes(yintercept = fed_min_wage), color = "red") +
+  facet_wrap(~decade, nrow = 5) +
+  labs(x = "State", 
+       y = "State Minimum Wage")
+
+# region and year
+ggplot(full_state_data, aes(x = regions)) +
+  geom_point(aes(y = state_min_wage)) +
+  geom_hline(aes(yintercept = fed_min_wage), color = "red") +
+  facet_wrap(~year, nrow = 10) +
+  labs(x = "State", 
+       y = "State Minimum Wage")
+```
+
 ##### Adjusted Min Wage by state and year (include adjusted fed (facet or use by decades where necessary)
+
+# figure out spacing of graph
+
+# can't do a line for adj federal minimum wage unless faceted by year instead of decade
+
+# add a label for the line saying federal minimum wage
+
+```{r}
+# abbreviation and decade
+ggplot(full_state_data, aes(x = state_abbreviation)) +
+  geom_point(aes(y = state_min_wage_2020_adj)) +
+  geom_hline(aes(yintercept = fed_min_wage_2020_adj), color = "red") +
+  facet_wrap(~decade, nrow = 5) +
+  theme(axis.text.x = element_text(angle=90)) +
+  labs(x = "State", 
+       y = "State Minimum Wage in 2020 Dollars")
+
+# region and decade
+ggplot(full_state_data, aes(x = regions)) +
+  geom_point(aes(y = state_min_wage_2020_adj)) +
+  geom_hline(aes(yintercept = fed_min_wage_2020_adj), color = "red") +
+  facet_wrap(~decade, nrow = 5) +
+  labs(x = "Region", 
+       y = "State Minimum Wage in 2020 Dollars")
+
+# region and year
+ggplot(full_state_data, aes(x = regions)) +
+  geom_point(aes(y = state_min_wage_2020_adj)) +
+  geom_hline(aes(yintercept = fed_min_wage_2020_adj), color = "red") +
+  facet_wrap(~year, nrow = 10) +
+  labs(x = "Region", 
+       y = "State Minimum Wage in 2020 Dollars")
+```
+
 ##### Union employment, CB employment, and full employment (facet by decade)
+
+# should we take the average value per decade?
+
+# add a key for the color of the points
+
+# fix spacing because union employment count is not visible
+
+```{r}
+# state and decade
+ggplot(full_state_data, aes(x = state_abbreviation)) +
+  geom_point(aes(y = employment_count), color = "red") +
+  geom_point(aes(y = union_employment_count), color = "blue") +
+  geom_point(aes(y = collective_bargain_employment_count), color = "green") +
+  facet_wrap(~decade, nrow = 5) +
+  labs(x = "State", 
+       y = "Number of People Employed")
+
+# region and decade
+ggplot(full_state_data, aes(x = regions)) +
+  geom_point(aes(y = employment_count), color = "red") +
+  geom_point(aes(y = union_employment_count), color = "blue") +
+  geom_point(aes(y = collective_bargain_employment_count), color = "green") +
+  facet_wrap(~decade, nrow = 5) +
+  labs(x = "Region", 
+       y = "Number of People Employed")
+```
+
 ##### percent union and percent collective bargaining
+
+# should we take the average value per decade?
+
+# add a key for the color of the points
+
+```{r}
+# state and decade
+ggplot(full_state_data, aes(x = state_abbreviation)) +
+  geom_point(aes(y = percent_union), color = "red") +
+  geom_point(aes(y = percent_collective_bargain), color = "blue") +
+  facet_wrap(~decade, nrow = 5) +
+  labs(x = "State", 
+       y = "Number of People Employed")
+
+# region and decade
+ggplot(full_state_data, aes(x = regions)) +
+  geom_point(aes(y = percent_union), color = "red") +
+  geom_point(aes(y = percent_collective_bargain), color = "blue") +
+  facet_wrap(~decade, nrow = 5) +
+  labs(x = "Region", 
+       y = "Number of People Employed")
+```
 

### 2024-01-09 Code/R FIle.qmd
more graphs. better code

@@
 library(tidyverse)
+library(usmap)
 ```
@@
 
-##### For facets, use best judgement if by state, decade, year, or region
+* For facets, use best judgement if by state, decade, year, or region
 
-##### maybe make all states light grey and just choose outline outlier states and maybe regional averages
+* maybe make all states light grey and just choose outline outlier states and maybe regional averages
 
@@
 ```{r}
-ggplot(full_state_data, aes(x = state_abbreviation, y = )) +
+library(gganimate)
+library(gifski)
+library(transformr)
+```
+
+```{r}
+map_data <- full_state_data |> 
+  select(c("year", "state", "state_abbreviation", "party_state_avaerage")) |> 
+  distinct(year, state, .keep_all = TRUE) |> 
+  filter(year %% 4==0) |> 
+  filter(!is.na(party_state_avaerage))
+
+plot_usmap(data = map_data, values = "party_state_avaerage", color = "black") +
+  facet_wrap(~year, nrow = 2, ncol= 6) +
+  theme(legend.position = "top") +
+  scale_fill_gradient(low = "red", high = "blue")
 ```
@@
 ```{r}
-effective_min_wage_adj_graph1 <- full_state_data |> 
+state_minimum_wage_data <- full_state_data |> 
   filter(!is.na(regions)) |>
-  filter(effective_min_wage_2020_adj != fed_min_wage_2020_adj) |> 
-  filter(state_abbreviation != "HI") |> 
+  filter(year <=2020 ) |> 
+  mutate(
+    map_color = case_when(
+      state_abbreviation == "CA" ~ "CA",
+      state_abbreviation == "AK" ~ "AK",
+      state_abbreviation == "OK" ~ "OK",
+      state_abbreviation == "FL" ~ "FL",
+      .default = NA
+    )
+  ) 
+state_minimum_wage_data |> 
   ggplot(aes(x = year)) +
-  geom_line(aes(y = effective_min_wage_2020_adj, color = state_abbreviation)) +
-  scale_color_manual(values = c(rep("gray", 34))) +
-  geom_line(aes(y = fed_min_wage), color = "red", linewidth = .75)
-
-HI_effective_min_wage <- full_state_data |> 
-  filter(!is.na(regions)) |> 
-  filter(effective_min_wage_2020_adj != fed_min_wage_2020_adj) |> 
-  filter(state_abbreviation == "HI") |> 
-  mutate(effective_min_wage_2020_adj = effective_min_wage_2020_adj - .05)
-
-CT_effective_min_wage <- full_state_data |> 
-  filter(!is.na(regions)) |> 
-  filter(effective_min_wage_2020_adj != fed_min_wage_2020_adj) |> 
-  filter(state_abbreviation == "CT")
-
-AK_effective_min_wage <- full_state_data |> 
-  filter(!is.na(regions)) |> 
-  filter(effective_min_wage_2020_adj != fed_min_wage_2020_adj) |> 
-  filter(state_abbreviation == "AK")
-
-CA_effective_min_wage <- full_state_data |> 
-  filter(!is.na(regions)) |> 
-  filter(effective_min_wage_2020_adj != fed_min_wage_2020_adj) |> 
-  filter(state_abbreviation == "CA")
-
-effective_min_wage_adj_graph1 +
-  geom_line(data = HI_effective_min_wage, aes(y = effective_min_wage_2020_adj), color = "blue",linewidth = .7) +
-  annotate("text", x=1994, y=9.5, label= "Hawaii", color = "blue") +
-  geom_line(data = CT_effective_min_wage, aes(y = effective_min_wage_2020_adj), color = "#288225", linewidth = .7) +
-  annotate("text", x=1985, y=7.5, label= "Connecticut", color = "#288225") +
-  geom_line(data = AK_effective_min_wage, aes(y = effective_min_wage_2020_adj), color = "orange", linewidth = .7) +
-  annotate("text", x=1985, y=10, label= "Alaska", color = "orange") +
-  geom_line(data = CA_effective_min_wage, aes(y = effective_min_wage_2020_adj), color = "purple", linewidth = .7) +
-  annotate("text", x=2016, y=12, label= "California", color = "purple") +
-  annotate("text", x=2000, y=5.5, label= "Federal", color = "red") +
+  geom_line(aes(y = state_min_wage_2020_adj, group = state_abbreviation),
+            color = "grey75", alpha =0.6) +
+  geom_line(data = subset(state_minimum_wage_data, !is.na(map_color)),
+            aes(y = state_min_wage_2020_adj, color = map_color),
+            linewidth = .75) +
+  geom_line(aes(y = fed_min_wage), color = "red", linewidth = .75) +
   theme_minimal() +
-  theme(legend.position = "none") +
-  labs(title = "State effective minimum wage by Year (adjusted for 2020 $)",
+  labs(title = "State Minimum Wage by Year (Adjusted for 2020 $)",
        subtitle = "For states with an adjusted effective minimum wage above the federal level",
        x = "Year",
-       y = "Effective minimum wage (2020 USD)")
-
-rm(HI_effective_min_wage, CT_effective_min_wage,
-   AK_effective_min_wage, CA_effective_min_wage)
+       y = "Minimum wage (2020 USD)")
 ```
@@
 # add a label for the line saying federal minimum wage
+
+```{r}
+effective_minimum_wage_data <- full_state_data |> 
+  filter(!is.na(regions)) |> 
+  filter(effective_min_wage_2020_adj != fed_min_wage_2020_adj) |> 
+  mutate(
+    map_color = case_when(
+      state_abbreviation == "CA" ~ "CA",
+      state_abbreviation == "AK" ~ "AK",
+      state_abbreviation == "CT" ~ "CT",
+      state_abbreviation == "HI" ~ "HI",
+      .default = NA
+    )
+  ) 
+effective_minimum_wage_data |> 
+  ggplot(aes(x = year)) +
+  geom_line(aes(y = effective_min_wage_2020_adj, group = state_abbreviation),
+            color = "grey75", alpha =0.6) +
+  geom_line(data = subset(effective_minimum_wage_data, !is.na(map_color)),
+            aes(y = effective_min_wage_2020_adj, color = map_color),
+            linewidth = .75) +
+  geom_line(aes(y = fed_min_wage), color = "red", linewidth = .75) +
+  theme_minimal() +
+  labs(title = "State Effective Minimum Wage by Year (Adjusted for 2020 $)",
+       subtitle = "For states with an adjusted effective minimum wage above the federal level",
+       x = "Year",
+       y = "Effective minimum wage (2020 USD)")
+```
 

### 2024-01-09 Code/R FIle.qmd

@@
 ```{r}
-state_minimum_wage_data <- full_state_data |> 
-  filter(!is.na(regions)) |>
-  filter(year <=2020 ) |> 
+state_minimum_wage_data <- full_state_data |> # This object will be used to create the graph and also in the data subset
+  filter(!is.na(regions)) |> #This just removes DC
+  filter(year <=2020 ) |>  #wage data only exists before 2020
   mutate(
     map_color = case_when(
-      state_abbreviation == "CA" ~ "CA",
+      state_abbreviation == "CA" ~ "CA", # States we would like to highlight. first "" is the state abb, second "" is what will appear in legend
       state_abbreviation == "AK" ~ "AK",
@@
   ) 
-state_minimum_wage_data |> 
+state_minimum_wage_data |> # use created table here
   ggplot(aes(x = year)) +
   geom_line(aes(y = state_min_wage_2020_adj, group = state_abbreviation),
-            color = "grey75", alpha =0.6) +
-  geom_line(data = subset(state_minimum_wage_data, !is.na(map_color)),
+            color = "grey75", alpha =0.6) + # using group means the groupings wont appear in legend
+  geom_line(data = subset(state_minimum_wage_data, # use created object here
+                          !is.na(map_color)), #subset used to keep out non highlighted states
             aes(y = state_min_wage_2020_adj, color = map_color),
             linewidth = .75) +
-  geom_line(aes(y = fed_min_wage), color = "red", linewidth = .75) +
+  geom_line(aes(y = fed_min_wage), color = "red", linewidth = .75) + # federal minimum wage
   theme_minimal() +
@@
        y = "Minimum wage (2020 USD)")
-```
 
-```{r}
-# abbreviation and decade
-ggplot(full_state_data, aes(x = state_abbreviation)) +
-  geom_point(aes(y = state_min_wage)) +
-  geom_hline(aes(yintercept = fed_min_wage), color = "red") +
-  facet_wrap(~decade, nrow = 5) +
-  theme(axis.text.x = element_text(angle=90)) +
-  labs(x = "State", 
-       y = "State Minimum Wage")
-
-# region and decade
-ggplot(full_state_data, aes(x = regions)) +
-  geom_point(aes(y = state_min_wage)) +
-  geom_hline(aes(yintercept = fed_min_wage), color = "red") +
-  facet_wrap(~decade, nrow = 5) +
-  labs(x = "State", 
-       y = "State Minimum Wage")
-
-# region and year
-ggplot(full_state_data, aes(x = regions)) +
-  geom_point(aes(y = state_min_wage)) +
-  geom_hline(aes(yintercept = fed_min_wage), color = "red") +
-  facet_wrap(~year, nrow = 10) +
-  labs(x = "State", 
-       y = "State Minimum Wage")
+rm(state_minimum_wage_data) # remove object to make graph
 ```
@@
        y = "Effective minimum wage (2020 USD)")
-```
 
-```{r}
-# abbreviation and decade
-ggplot(full_state_data, aes(x = state_abbreviation)) +
-  geom_point(aes(y = state_min_wage_2020_adj)) +
-  geom_hline(aes(yintercept = fed_min_wage_2020_adj), color = "red") +
-  facet_wrap(~decade, nrow = 5) +
-  theme(axis.text.x = element_text(angle=90)) +
-  labs(x = "State", 
-       y = "State Minimum Wage in 2020 Dollars")
-
-# region and decade
-ggplot(full_state_data, aes(x = regions)) +
-  geom_point(aes(y = state_min_wage_2020_adj)) +
-  geom_hline(aes(yintercept = fed_min_wage_2020_adj), color = "red") +
-  facet_wrap(~decade, nrow = 5) +
-  labs(x = "Region", 
-       y = "State Minimum Wage in 2020 Dollars")
-
-# region and year
-ggplot(full_state_data, aes(x = regions)) +
-  geom_point(aes(y = state_min_wage_2020_adj)) +
-  geom_hline(aes(yintercept = fed_min_wage_2020_adj), color = "red") +
-  facet_wrap(~year, nrow = 10) +
-  labs(x = "Region", 
-       y = "State Minimum Wage in 2020 Dollars")
+rm(effective_minimum_wage_data)
 ```
@@
 
-```{r}
-# state and decade
-ggplot(full_state_data, aes(x = state_abbreviation)) +
-  geom_point(aes(y = employment_count), color = "red") +
-  geom_point(aes(y = union_employment_count), color = "blue") +
-  geom_point(aes(y = collective_bargain_employment_count), color = "green") +
-  facet_wrap(~decade, nrow = 5) +
-  labs(x = "State", 
-       y = "Number of People Employed")
-
-# region and decade
-ggplot(full_state_data, aes(x = regions)) +
-  geom_point(aes(y = employment_count), color = "red") +
-  geom_point(aes(y = union_employment_count), color = "blue") +
-  geom_point(aes(y = collective_bargain_employment_count), color = "green") +
-  facet_wrap(~decade, nrow = 5) +
-  labs(x = "Region", 
-       y = "Number of People Employed")
-```
-
 ##### percent union and percent collective bargaining
@@
 ```{r}
-# state and decade
-ggplot(full_state_data, aes(x = state_abbreviation)) +
-  geom_point(aes(y = percent_union), color = "red") +
-  geom_point(aes(y = percent_collective_bargain), color = "blue") +
-  facet_wrap(~decade, nrow = 5) +
-  labs(x = "State", 
-       y = "Number of People Employed")
+per_union_wage_data <- full_state_data |> # This object will be used to create the graph and also in the data subset
+  filter(!is.na(regions)) |> #This just removes DC
+  filter(year <=2020 ) |>  #wage data only exists before 2020
+  mutate(
+    map_color = case_when(
+      state_abbreviation == "CA" ~ "CA", # States we would like to highlight. first "" is the state abb, second "" is what will appear in legend
+      state_abbreviation == "AK" ~ "AK",
+      state_abbreviation == "OK" ~ "OK",
+      state_abbreviation == "FL" ~ "FL",
+      .default = NA
+    )
+  ) 
+per_union_wage_data |> # use created table here
+  ggplot(aes(x = year)) +
+  geom_line(aes(y = percent_union, group = state_abbreviation),
+            color = "grey75", alpha =0.6) + # using group means the groupings wont appear in legend
+  geom_line(data = subset(per_union_wage_data, # use created object here
+                          !is.na(map_color)), #subset used to keep out non highlighted states
+            aes(y = percent_union, color = map_color),
+            linewidth = .75) +
+  theme_minimal() +
+  labs(title = "States by % of labor market covered by union",
+       x = "Year",
+       y = "% of labor market in union")
 
-# region and decade
-ggplot(full_state_data, aes(x = regions)) +
-  geom_point(aes(y = percent_union), color = "red") +
-  geom_point(aes(y = percent_collective_bargain), color = "blue") +
-  facet_wrap(~decade, nrow = 5) +
-  labs(x = "Region", 
-       y = "Number of People Employed")
+rm(per_union_wage_data)
+```
+```{r}
+per_collective_wage_data <- full_state_data |> # This object will be used to create the graph and also in the data subset
+  filter(!is.na(regions)) |> #This just removes DC
+  filter(year <=2020 ) |>  #wage data only exists before 2020
+  mutate(
+    map_color = case_when(
+      state_abbreviation == "CA" ~ "CA", # States we would like to highlight. first "" is the state abb, second "" is what will appear in legend
+      state_abbreviation == "AK" ~ "AK",
+      state_abbreviation == "HI" ~ "HI",
+      state_abbreviation == "FL" ~ "FL",
+      .default = NA
+    )
+  ) 
+per_collective_wage_data |> # use created table here
+  ggplot(aes(x = year)) +
+  geom_line(aes(y = percent_collective_bargain, group = state_abbreviation),
+            color = "grey75", alpha =0.6) + # using group means the groupings wont appear in legend
+  geom_line(data = subset(per_collective_wage_data, # use created object here
+                          !is.na(map_color)), #subset used to keep out non highlighted states
+            aes(y = percent_collective_bargain, color = map_color),
+            linewidth = .75) +
+  theme_minimal() +
+  labs(title = "States by % of labor market covered by collective Bargaining Agreement",
+       x = "Year",
+       y = "% of labor market in union")
+
+rm(per_collective_wage_data)
 ```

### 2024-01-09 Code/R FIle.qmd
fixed typo

@@
 map_data <- full_state_data |> 
-  select(c("year", "state", "state_abbreviation", "party_state_avaerage")) |> 
+  select(c("year", "state", "state_abbreviation", "party_state_average")) |> 
   distinct(year, state, .keep_all = TRUE) |> 

### 2024-01-09 Code/R FIle.qmd
fixed average spelling

@@
   group_by(year, state) |> 
-  mutate(party_state_avaerage = mean(party_binary)) |>
+  mutate(party_state_average = mean(party_binary)) |>
   ungroup() |> 
@@
                       "regions", "district", "party", "party_binary",
+                      "party_state_average",
                       "employment_count", "union_employment_count",
@@
   filter(year %% 4==0) |> 
-  filter(!is.na(party_state_avaerage))
+  filter(!is.na(party_state_average))
 
-plot_usmap(data = map_data, values = "party_state_avaerage", color = "black") +
+plot_usmap(data = map_data, values = "party_state_average", color = "black") +
   facet_wrap(~year, nrow = 2, ncol= 6) +

### 2024-01-09 Code/R FIle.qmd
deleted some commented notes

@@
 
-* For facets, use best judgement if by state, decade, year, or region
-
-* maybe make all states light grey and just choose outline outlier states and maybe regional averages
-
 ##### Party by state and year
@@
 ##### Min Wage by state and year (include fed (facet or use by decades where necessary)
-
-# figure out spacing of graph
-
-# can't do a line for federal minimum wage unless faceted by year instead of decade
-
 ```{r}
@@
 ##### Adjusted Min Wage by state and year (include adjusted fed (facet or use by decades where necessary)
-
-# figure out spacing of graph
-
-# can't do a line for adj federal minimum wage unless faceted by year instead of decade
-
-# add a label for the line saying federal minimum wage
-
 ```{r}
@@
 
-# should we take the average value per decade?
-
-# add a key for the color of the points
-
-# fix spacing because union employment count is not visible
-
 ##### percent union and percent collective bargaining
-
-# should we take the average value per decade?
-
-# add a key for the color of the points
-
 ```{r}

### 2024-01-09 Code/R FIle.qmd
removed state_abbreviation from map data

@@
 map_data <- full_state_data |> 
-  select(c("year", "state", "state_abbreviation", "party_state_average")) |> 
+  select(c("year", "state", "party_state_average")) |> 
   distinct(year, state, .keep_all = TRUE) |> 

### 2024-01-09 Code/R FIle.qmd
updated x axis to be every 4 years

@@
   filter(!is.na(regions)) |> #This just removes DC
-  filter(year <=2020 ) |>  #wage data only exists before 2020
+  filter(year <=2020 & year >=1984) |>  #wage data only exists before 2020
   mutate(
@@
   theme_minimal() +
+  scale_x_continuous(breaks = seq(min(state_minimum_wage_data$year), max(state_minimum_wage_data$year), by = 4)) + 
   labs(title = "State Minimum Wage by Year (Adjusted for 2020 $)",
@@
   filter(effective_min_wage_2020_adj != fed_min_wage_2020_adj) |> 
+  filter(year <=2020 & year >=1984) |>
   mutate(
@@
   theme_minimal() +
+  scale_x_continuous(breaks = seq(min(effective_minimum_wage_data$year), max(effective_minimum_wage_data$year), by = 4)) + 
   labs(title = "State Effective Minimum Wage by Year (Adjusted for 2020 $)",
@@
   filter(!is.na(regions)) |> #This just removes DC
-  filter(year <=2020 ) |>  #wage data only exists before 2020
+  filter(year <= 2020 & year >= 1984) |>  #wage data only exists before 2020
   mutate(
@@
   theme_minimal() +
+  scale_x_continuous(breaks = seq(min(per_union_wage_data$year), max(per_union_wage_data$year), by = 4)) + 
   labs(title = "States by % of labor market covered by union",
@@
   theme_minimal() +
+  scale_x_continuous(breaks = seq(min(per_collective_wage_data$year), max(per_collective_wage_data$year), by = 4)) + 
   labs(title = "States by % of labor market covered by collective Bargaining Agreement",

### 2024-01-09 Code/R FIle.qmd
updated national dataset to have employment counts and adjusted wages and income for 2020

@@
 ```{r}
-state_to_ntl <- full_state_data |>
-  ungroup() %>% 
-  mutate(adjusted_wage_prop = fed_min_wage / fed_min_wage_2020_adj, .after = year) |>
-  distinct(year, .keep = TRUE) 
-#  select(year:adjusted_wage_prop)
+state_to_ntl_adjusted <- full_state_data |>
+  mutate(prop_2020_adj = fed_min_wage_2020_adj / fed_min_wage, .after = year) |>
+  distinct(year, .keep_all = TRUE) |>
+  select(year, prop_2020_adj)
+```
+
+```{r}
+state_to_ntl_employment <- full_state_data |>
+  group_by(year) |>
+  summarise_at(vars(employment_count:collective_bargain_employment_count), sum)
 ```
@@
   left_join(union_wages, by = "year") |>
-  mutate(decade = paste0((year - year %% 10), "s"), .after = year)
+  left_join(state_to_ntl_adjusted, by = "year") |>
+  left_join(state_to_ntl_employment, by = "year") |>
+  mutate(decade = paste0((year - year %% 10), "s"), .after = year) |>
+  mutate(mean_median_income_difference = mean_income - median_income, .after = mean_income) |>
+   mutate(median_income_2020_adj = median_income * prop_2020_adj,
+         mean_income_2020_adj = mean_income * prop_2020_adj,
+         mean_median_income_difference_2020_adj = mean_median_income_difference * prop_2020_adj,
+         wage_2020_adj = wage * prop_2020_adj,
+         union_wage_2020_adj = union_wage * prop_2020_adj,
+         nonunion_wage_2020_adj = nonunion_wage * prop_2020_adj)
 ```

### 2024-01-10 Code/R FIle.qmd
cleaned up package loading, data loading, and data cleaning

@@
 
-# useful git commands
+# 1. Comparing interactions between collective bargaining rates, wages, and House of Representative elections
 
-**This command will update all the branches on your computer** git remote update origin --prune
+** Insert abstract of project, where we got our data, what our data looks like, and other information here
 
-showing off branch
-
-test
-
-### Loading Packages
+# 2. Data Analysis 
+## 2.1 Getting Set Up
 
 ```{r}
-library(tidyverse)
-library(usmap)
+suppressPackageStartupMessages(library(tidyverse))
+library(usmap) # Used for map plots
 ```
 
-### Gathering Data
-
 ```{r}
-# Union Data
+# Union Data: obtained from Tidy Tuesday
 union_wages <- read_rds("../Data/union_wages.rds")
@@
 
-# Election Data
+# Election Data: Obtained from Tidy Tuesday
 house <- read_rds("../Data/house.rds")
 
-# Wage Data
+# Wage Data: Obtained from Tidy Tuesday
 income_distribution <- read_rds("../Data/income_distribution.rds")
 
-# Minimum Wage Data
-minimum_wage <- read_csv("../Data/minimum_wage_data.csv")
+# Minimum Wage Data: Obtained from U.S. Department of Labor
+minimum_wage <- read_csv("../Data/minimum_wage_data.csv",show_col_types = FALSE)
 ```
 
-### Cleaning Data
+## 2.2 Basic Data Cleaning
+### 2.2.1 Examining data shape
 
-#### Examining data shape
-
-First, lets get an idea of what our data looks like.
+Before we do anything to our data, we will examine our data to ensure everything loaded in correctly
 
 ```{r}
-# look at columns
 head(union_wages)
@@
 
-#### Cleaning data script
+### 2.2.2 Cleaning data script
 
@@
   rename(median_income = income_median,
-         mean_income = income_mean) # renaming columns
+         mean_income = income_mean) |> # renaming columns
+  filter(race == "All Races") |> 
+  select(-race)
 ```
@@
 
-### Combining dataframes
+### 2.2.3 Merging dataframes
 
@@
 
-#### State Dataframes
+#### 2.2.3.1 State Dataframes
 
@@
 ```{r}
+head(full_state_data)
 summary(full_state_data)
@@
 
-#### National Dataframes
-
-##### Creating Adjusted Wages Columns
+#### 2.2.3.2 National Dataframes
 
@@
          union_wage_2020_adj = union_wage * prop_2020_adj,
-         nonunion_wage_2020_adj = nonunion_wage * prop_2020_adj)
+         nonunion_wage_2020_adj = nonunion_wage * prop_2020_adj) |> 
+  select(-prop_2020_adj)
+```
+
+```{r}
+ntl_column_list <- c("year", "decade", "mean_income",
+                     "mean_income_2020_adj", "median_income",
+                     "median_income_2020_adj", "mean_median_income_difference",
+                     "mean_median_income_difference_2020_adj", "wage",
+                     "wage_2020_adj", "union_wage", "union_wage_2020_adj",
+                     "nonunion_wage", "nonunion_wage_2020_adj",
+                     "wage_percent_difference", 
+                     "wage_percent_difference_adjusted",
+                     "employment_count", "union_employment_count", 
+                     "collective_bargain_employment_count", "union_type")
+full_ntl_data <- full_ntl_data[, ntl_column_list] # reorder columns
+```
+
+```{r}
+rm(house, income_distribution, minimum_wage, state_to_ntl_adjusted,
+   state_to_ntl_employment, union_states, union_wages, state_column_list,
+   ntl_column_list)
 ```
@@
 
-### EDA
-
-```{r}
-summary(full_ntl_data)
-summary(full_state_data)
-```
+## 2.3 EDA
 
@@
 ##### Party by state and year
-
-```{r}
-library(gganimate)
-library(gifski)
-library(transformr)
-```
 

### 2024-01-10 Code/R FIle.qmd

@@
          union_wage_2020_adj = union_wage * prop_2020_adj,
-         nonunion_wage_2020_adj = nonunion_wage * prop_2020_adj) |> 
+         nonunion_wage_2020_adj = nonunion_wage * prop_2020_adj,
+         percent_union = union_employment_count / employment_count,
+         percent_collective_bargain = collective_bargain_employment_count / employment_count) |> 
   select(-prop_2020_adj)
@@
                      "employment_count", "union_employment_count", 
-                     "collective_bargain_employment_count", "union_type")
+                     "collective_bargain_employment_count",
+                     "percent_union", "percent_collective_bargain",
+                     "union_type")
 full_ntl_data <- full_ntl_data[, ntl_column_list] # reorder columns
@@
 
-#### Graphing Full State Data
+For our exploratory data analysis, we will look at interactions between collective bargaining rates, wages, and house election. We will also examine basic distribution
 
-##### Party by state and year
+### 2.3.1 Collective Bargaining
 
 ```{r}
-map_data <- full_state_data |> 
-  select(c("year", "state", "party_state_average")) |> 
-  distinct(year, state, .keep_all = TRUE) |> 
-  filter(year %% 4==0) |> 
-  filter(!is.na(party_state_average))
+per_union_wage_data <- full_state_data |> # This object will be used to create the graph and also in the data subset
+  filter(!is.na(regions)) |> #This just removes DC
+  filter(year <= 2020 & year >= 1984) |>  #wage data only exists before 2020
+  mutate(
+    map_color = case_when(
+      state_abbreviation == "CA" ~ "CA", # States we would like to highlight. first "" is the state abb, second "" is what will appear in legend
+      state_abbreviation == "AK" ~ "AK",
+      state_abbreviation == "OK" ~ "OK",
+      state_abbreviation == "FL" ~ "FL",
+      .default = NA
+    )
+  ) 
+per_union_wage_data |> # use created table here
+  ggplot(aes(x = year)) +
+  geom_line(aes(y = percent_union, group = state_abbreviation),
+            color = "grey75", alpha =0.6) + # using group means the groupings wont appear in legend
+  geom_line(data = subset(per_union_wage_data, # use created object here
+                          !is.na(map_color)), #subset used to keep out non highlighted states
+            aes(y = percent_union, color = map_color),
+            linewidth = .75) +
+  geom_line(data = subset(full_ntl_data, year >= 1984),
+            aes(y = percent_union), color = "orange", linewidth = 1.5) +
+  annotate("text", x=1988, y=.15, label= "Federal Avg", color = "orange", size = 5) +
+  theme_minimal() +
+  scale_x_continuous(breaks = seq(min(per_union_wage_data$year), max(per_union_wage_data$year), by = 4)) + 
+  labs(title = "States by % of labor market covered by union",
+       x = "Year",
+       y = "% of labor market in union")
 
-plot_usmap(data = map_data, values = "party_state_average", color = "black") +
-  facet_wrap(~year, nrow = 2, ncol= 6) +
-  theme(legend.position = "top") +
-  scale_fill_gradient(low = "red", high = "blue")
+rm(per_union_wage_data)
 ```
 
-##### Min Wage by state and year (include fed (facet or use by decades where necessary)
+```{r}
+per_collective_wage_data <- full_state_data |> # This object will be used to create the graph and also in the data subset
+  filter(!is.na(regions)) |> #This just removes DC
+  filter(year <= 2020 & year >= 1984) |>  #wage data only exists before 2020
+  mutate(
+    map_color = case_when(
+      state_abbreviation == "CA" ~ "CA", # States we would like to highlight. first "" is the state abb, second "" is what will appear in legend
+      state_abbreviation == "AK" ~ "AK",
+      state_abbreviation == "HI" ~ "HI",
+      state_abbreviation == "FL" ~ "FL",
+      .default = NA
+    )
+  ) 
+per_collective_wage_data |> # use created table here
+  ggplot(aes(x = year)) +
+  geom_line(aes(y = percent_collective_bargain, group = state_abbreviation),
+            color = "grey75", alpha =0.6) + # using group means the groupings wont appear in legend
+  geom_line(data = subset(per_collective_wage_data, # use created object here
+                          !is.na(map_color)), #subset used to keep out non highlighted states
+            aes(y = percent_collective_bargain, color = map_color),
+            linewidth = .75) +
+  geom_line(data = subset(full_ntl_data, year >= 1984),
+            aes(y = percent_collective_bargain), color = "orange", linewidth = 1.5) +
+  annotate("text", x=1988, y=.17, label= "Federal Avg", color = "orange", size = 5) +
+  theme_minimal() +
+  scale_x_continuous(breaks = seq(min(per_collective_wage_data$year), max(per_collective_wage_data$year), by = 4)) + 
+  labs(title = "States by % of labor market covered by collective Bargaining Agreement",
+       x = "Year",
+       y = "% of labor market in union")
+
+rm(per_collective_wage_data)
+```
+
+### 2.3.2 Wages
+
 ```{r}
@@
 
-##### Adjusted Min Wage by state and year (include adjusted fed (facet or use by decades where necessary)
 ```{r}
@@
 ```
+```{r}
+full_ntl_data |> 
+  filter(year >= 1984) |> 
+  ggplot(aes(x = year)) +
+  geom_line(aes(y = mean_income_2020_adj), color = "red") +
+  geom_line(aes(y = median_income_2020_adj), color = "blue") +
+  theme_minimal()
+```
 
-##### Union employment, CB employment, and full employment (facet by decade)
+### 2.3.3 House Elections
 
-##### percent union and percent collective bargaining
 ```{r}
-per_union_wage_data <- full_state_data |> # This object will be used to create the graph and also in the data subset
-  filter(!is.na(regions)) |> #This just removes DC
-  filter(year <= 2020 & year >= 1984) |>  #wage data only exists before 2020
-  mutate(
-    map_color = case_when(
-      state_abbreviation == "CA" ~ "CA", # States we would like to highlight. first "" is the state abb, second "" is what will appear in legend
-      state_abbreviation == "AK" ~ "AK",
-      state_abbreviation == "OK" ~ "OK",
-      state_abbreviation == "FL" ~ "FL",
-      .default = NA
-    )
-  ) 
-per_union_wage_data |> # use created table here
+map_data <- full_state_data |> 
+  select(c("year", "state", "party_state_average")) |> 
+  distinct(year, state, .keep_all = TRUE) |> 
+  filter(year %% 4==0) |> 
+  filter(!is.na(party_state_average))
+
+plot_usmap(data = map_data, values = "party_state_average", color = "black") +
+  facet_wrap(~year, nrow = 2, ncol= 6) +
+  theme(legend.position = "top") +
+  scale_fill_gradient(low = "red", high = "blue")
+
+rm(map_data)
+```
+
+### 2.3.4 Collective Bargaining and Wages
+
+```{r}
+full_ntl_data |> 
+  filter(year >= 1984) |> 
   ggplot(aes(x = year)) +
-  geom_line(aes(y = percent_union, group = state_abbreviation),
-            color = "grey75", alpha =0.6) + # using group means the groupings wont appear in legend
-  geom_line(data = subset(per_union_wage_data, # use created object here
-                          !is.na(map_color)), #subset used to keep out non highlighted states
-            aes(y = percent_union, color = map_color),
-            linewidth = .75) +
-  theme_minimal() +
-  scale_x_continuous(breaks = seq(min(per_union_wage_data$year), max(per_union_wage_data$year), by = 4)) + 
-  labs(title = "States by % of labor market covered by union",
-       x = "Year",
-       y = "% of labor market in union")
-
-rm(per_union_wage_data)
+  geom_line(aes(y = union_wage_2020_adj), color = "blue") +
+  geom_line(aes(y = nonunion_wage_2020_adj), color = "red") + 
+  theme_minimal()
 ```
@@
 ```{r}
-per_collective_wage_data <- full_state_data |> # This object will be used to create the graph and also in the data subset
-  filter(!is.na(regions)) |> #This just removes DC
-  filter(year <= 2020 & year >= 1984) |>  #wage data only exists before 2020
-  mutate(
-    map_color = case_when(
-      state_abbreviation == "CA" ~ "CA", # States we would like to highlight. first "" is the state abb, second "" is what will appear in legend
-      state_abbreviation == "AK" ~ "AK",
-      state_abbreviation == "HI" ~ "HI",
-      state_abbreviation == "FL" ~ "FL",
-      .default = NA
-    )
-  ) 
-per_collective_wage_data |> # use created table here
-  ggplot(aes(x = year)) +
-  geom_line(aes(y = percent_collective_bargain, group = state_abbreviation),
-            color = "grey75", alpha =0.6) + # using group means the groupings wont appear in legend
-  geom_line(data = subset(per_collective_wage_data, # use created object here
-                          !is.na(map_color)), #subset used to keep out non highlighted states
-            aes(y = percent_collective_bargain, color = map_color),
-            linewidth = .75) +
-  theme_minimal() +
-  scale_x_continuous(breaks = seq(min(per_collective_wage_data$year), max(per_collective_wage_data$year), by = 4)) + 
-  labs(title = "States by % of labor market covered by collective Bargaining Agreement",
-       x = "Year",
-       y = "% of labor market in union")
+full_ntl_data |> 
+  filter(year >= 1984) |> 
+  ggplot(aes(x = percent_collective_bargain)) +
+  geom_point(aes(y = wage), color = "blue") +
+  geom_point(aes(y = union_wage), color = "red") +
+  facet_wrap(~decade) 
+```
 
-rm(per_collective_wage_data)
+### 2.3.5 Collective Bargaining and House Elections
+
+```{r}
+full_state_data |> 
+  filter(year >= 1984) |> 
+  ggplot(aes(x = percent_collective_bargain, y = party_state_average)) +
+  geom_point(color = "blue") +
+  geom_smooth(method = "lm") +
+  facet_wrap(~decade) 
 ```
+
+### 2.3.6 Wages and House elections
 

### 2024-01-10 Code/R FIle.qmd
worked more on graphs

@@
 
-** Insert abstract of project, where we got our data, what our data looks like, and other information here
+**Insert abstract of project, where we got our data, what our data looks like, and other information here**
 
 # 2. Data Analysis
+
 ## 2.1 Getting Set Up
@@
 ## 2.2 Basic Data Cleaning
+
 ### 2.2.1 Examining data shape
@@
 
-For our exploratory data analysis, we will look at interactions between collective bargaining rates, wages, and house election. We will also examine basic distribution
+For our exploratory data analysis, we will look at interactions between collective bargaining rates, wages, and house election. We will also examine basic distribution.
 
@@
       state_abbreviation == "FL" ~ "FL",
+      state_abbreviation == "HI" ~ "HI",
       .default = NA
@@
   scale_x_continuous(breaks = seq(min(per_union_wage_data$year), max(per_union_wage_data$year), by = 4)) + 
-  labs(title = "States by % of labor market covered by union",
+  labs(title = "States By % of Labor Market Covered By Union",
        x = "Year",
-       y = "% of labor market in union")
+       y = "% of Labor Market In Union",
+       color = "Selected States")
 
@@
 ```
+Our first graph shows what is commonly known about unions in the United States: Union membership has been on the decline since at least the 1980s.
+Unfortunately our data does not extend to the last two years where some reports indicate union memberships are on the rise.
+Early on, we can see that southern and more Republican leading states have lower union rates than more democrat and northern / western states.
+There is a very small bump in membership in 2008, which we hypothesize happened in relation to the 2008 financial crisis.
 
@@
   scale_x_continuous(breaks = seq(min(per_collective_wage_data$year), max(per_collective_wage_data$year), by = 4)) + 
-  labs(title = "States by % of labor market covered by collective Bargaining Agreement",
+  labs(title = "States by % of Labor Market Covered By a Collective Bargaining Agreement",
        x = "Year",
-       y = "% of labor market in union")
+       y = "% of Labor Market in a Collective Bargaining Agreement")
 
@@
 ```
+This graph is almost the same exact as the first, although the rates are slightly higher because this graph includes all collective bargaining agreements.
 
@@
             linewidth = .75) +
-  geom_line(aes(y = fed_min_wage), color = "red", linewidth = .75) + # federal minimum wage
+  geom_line(aes(y = fed_min_wage_2020_adj), color = "red", linewidth = .75) + # federal minimum wage
   theme_minimal() +
@@
        x = "Year",
-       y = "Minimum wage (2020 USD)")
+       y = "Minimum Wage (2020 USD)")
 
@@
 ```
+This graph can be used to draw a number of interesting inferences.
+First, there is a decently dark line at the bottom that indicates a number of states still have no minimum wage.
+This will be examined in greater detailed in the next graph.
+Next, we can see tht there is a significan shift post 2016 of raising the minimum wage at the state level.
+Before that time almost no state had a adjusted minimum wage over $10.
+Now, there are numerous states that surpass this value.
+Next, we can see that the federal minimum wage has been slowly losing value due to inflation since it was last raised in 2009.
 
@@
             linewidth = .75) +
-  geom_line(aes(y = fed_min_wage), color = "red", linewidth = .75) +
+  geom_line(aes(y = fed_min_wage_2020_adj), color = "red", linewidth = .75) +
   theme_minimal() +
@@
 ```
+This graph removes all data points where the state minimum wage falls below the federal amount.
+From this we can see that more and more states are taking it into their own hands to protect workers rights.
+We can see a large jump in 2007 which may be related to the financial crisis.
+
 ```{r}
@@
   geom_line(aes(y = median_income_2020_adj), color = "blue") +
-  theme_minimal()
+  theme_minimal() +
+  labs(
+    title = "Mean and Median income in the United States Between 1984 and 2019",
+    subtitle = "Using incomes adjusted to 2020 rates",
+    x = "Year",
+    y = "Income in USD (Adjusted for Inflation to 2020 Rates"
+  )
 ```
+
+```{r}
+full_ntl_data |> 
+  filter(year >= 1984) |> 
+  ggplot(aes(x = year)) +
+  geom_line(aes(y = mean_income), color = "red") +
+  geom_line(aes(y = median_income), color = "blue") +
+  theme_minimal() +
+  labs(
+    title = "Mean and Median income in the United States Between 1984 and 2019",
+    subtitle = "Using incomes adjusted to 2020 rates",
+    x = "Year",
+    y = "Income in USD (Adjusted for Inflation to 2020 Rates"
+  )
+```
+
+The two above graphs show that when accounting for inflation, Americans are earning less per year now than they did 30 years ago.
 
@@
 ```
+We can see that, especially in the northern-midwest states, that the country has grown to be more republican in he last 30 years.
+However, in the last 10 years this has begun to shift again.
+During this whole time, it appears that New England has become more democrat.
 
@@
 ```
+From our income graphs, this graph is rather unintuitive, as we would expect adjusted wages to be decreasing in this time.
+There are a number of potential causes for this.
+First, Americans work less hours today than they did in 1980.
+Second, and more likely, is that the income variables we used includes those who are not working.
+As the U.S.
+experiences a graying of society, this could be driving down mean incomes.
 
@@
   ggplot(aes(x = percent_collective_bargain)) +
-  geom_point(aes(y = wage), color = "blue") +
-  geom_point(aes(y = union_wage), color = "red") +
+  geom_point(aes(y = wage_2020_adj), color = "blue") +
+  geom_jitter(aes(y = union_wage_2020_adj), color = "red") +
   facet_wrap(~decade) 
 ```
+
+This graph is incredibly interesting.
+First, a quick view of this grpah shows the exact opposite of our hypothesis.
+Looking at each decade we can see collective bargaining rates decrease and wage increase.
+Then, in each decade that relationship continues.
+Still, we can see that in every decade, a collective bargaining agreement will lead to an increase in wages for a worker over a non-collective-bargaining-jobs.
 
@@
   geom_point(color = "blue") +
-  geom_smooth(method = "lm") +
-  facet_wrap(~decade) 
+  geom_smooth(method = "lm", color = "red") +
+  facet_wrap(~decade) +
+  labs(
+    title = "Percent collective Bargaining vs Proportion of House Delegation that is Democratic",
+    x = "Percent of workers covered by Collective Bargaining",
+    y = "Proportion of House Delegation that is Democration"
+  )
 ```
 
-### 2.3.6 Wages and House elections
+This graph demonstrates how political collective bargaining has become over the last 40 years.
+In the 1980s, there was almost no relationship between collective bargaining agreements and party affiliation.
+However, by 2020, there is now a significant bias
+
+```{r}
+full_state_data |> 
+  filter(year <2020 & year >= 1984) |> 
+  filter(year %% 4==0) |>
+  ggplot(aes(x = percent_collective_bargain, y = party_state_average)) +
+  geom_jitter(aes(color = regions), alpha = 0.5, width = 0.1, height = 0.1) +
+  geom_smooth(method = "lm", color = "black") +
+  facet_wrap(~decade) +
+  labs(
+    title = "Percent collective Bargaining vs Proportion of House Delegation that is Democratic",
+    x = "Percent of workers covered by Collective Bargaining",
+    y = "Proportion of House Delegation that is Democration"
+  )
+```
 

### 2024-01-10 Code/Presentation.qmd
created final presentation

---
title: "WIDS 2024 Final Presentation"
author: "Michael Taffe, Melanie Klein, Ramiro Arevalo"
format: 
  revealjs:
    theme: simple
editor: visual
---

## Project Abstract

## About the Data

* **Union data**
  * Gathered from Tidy Tuesday, although the original 
* **House election data**
  *
* **Wage**
  *

## EDA

When you click the **Render** button a presentation will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

## Models

## Conclusion

## Where we would take our project in the future

### 2024-01-11 Code/R FIle.qmd
worked on models

@@
 
-in the house data set, we can see a significant amount of NA values in the runoff column
+In the house data set, we can see a significant amount of NA values in the runoff column
 
@@
        x = "Year",
-       y = "% of Labor Market in a Collective Bargaining Agreement")
+       y = "% of Labor Market in a Collective Bargaining Agreement",
+       color = "Selected States")
 
@@
        x = "Year",
-       y = "Minimum Wage (2020 USD)")
+       y = "Minimum Wage (2020 USD)",
+       color = "Selected States")
 
@@
        x = "Year",
-       y = "Effective minimum wage (2020 USD)")
+       y = "Effective minimum wage (2020 USD)",
+       color = "Selected States")
 
@@
   labs(
-    title = "Mean and Median income in the United States Between 1984 and 2019",
-    subtitle = "Using incomes adjusted to 2020 rates",
+    title = "Mean and Median Income in the United States Between 1984 and 2019",
+    subtitle = "Using mean (red) and median (blue) incomes adjusted to 2020 rates",
     x = "Year",
@@
   labs(
-    title = "Mean and Median income in the United States Between 1984 and 2019",
-    subtitle = "Using incomes adjusted to 2020 rates",
+    title = "Mean and Median Income in the United States Between 1984 and 2019",
+    subtitle = "Using mean (red) and median (blue) incomes adjusted to 2020 rates",
     x = "Year",
-    y = "Income in USD (Adjusted for Inflation to 2020 Rates"
+    y = "Income in USD (Adjusted for Inflation to 2020 Rates)"
   )
@@
 ```
-We can see that, especially in the northern-midwest states, that the country has grown to be more republican in he last 30 years.
+We can see that, especially in the northern-midwest states, that the country has grown to be more republican in the last 30 years.
 However, in the last 10 years this has begun to shift again.
@@
 This graph is incredibly interesting.
-First, a quick view of this grpah shows the exact opposite of our hypothesis.
+First, a quick view of this graph shows the exact opposite of our hypothesis.
 Looking at each decade we can see collective bargaining rates decrease and wage increase.
@@
   labs(
-    title = "Percent collective Bargaining vs Proportion of House Delegation that is Democratic",
-    x = "Percent of workers covered by Collective Bargaining",
-    y = "Proportion of House Delegation that is Democration"
+    title = "Percent Collective Bargaining vs Proportion of House Delegation that is Democratic",
+    x = "Percent of Workers Covered by Collective Bargaining",
+    y = "Proportion of House Delegation that is Democratic"
   )
@@
 In the 1980s, there was almost no relationship between collective bargaining agreements and party affiliation.
-However, by 2020, there is now a significant bias
+However, by 2020, there is now a significant bias.
 
@@
   labs(
-    title = "Percent collective Bargaining vs Proportion of House Delegation that is Democratic",
-    x = "Percent of workers covered by Collective Bargaining",
-    y = "Proportion of House Delegation that is Democration"
+    title = "Percent Collective Bargaining vs Proportion of House Delegation that is Democratic",
+    x = "Percent of Workers Covered by Collective Bargaining",
+    y = "Proportion of House Delegation that is Democratic",
+    color = "Region"
+  )
+```
+
+This graph demonstrates the same as the previous graph, but with a linear regression line for each region.
+
+```{r}
+full_state_data |> 
+  filter(year <2020 & year >= 1984) |> 
+  filter(year %% 4==0) |>
+  ggplot(aes(x = percent_collective_bargain, y = party_state_average)) +
+  geom_jitter(aes(color = regions), alpha = 0.1, width = 0.1, height = 0.1) +
+  geom_smooth(method = "lm", aes(group = regions, color = regions), se = FALSE) +
+  facet_wrap(~decade) +
+  labs(
+    title = "Percent Collective Bargaining vs Proportion of House Delegation that is Democratic",
+    x = "Percent of Workers Covered by Collective Bargaining",
+    y = "Proportion of House Delegation that is Democratic",
+    color = "Region"
   )
@@
 
+T-test between union and non-union wages
+```{r}
+# 1984 to 2019
+t.test(full_ntl_data$union_wage_2020_adj, full_ntl_data$nonunion_wage_2020_adj)
+
+# 1980s
+full_ntl_data_1980s <- full_ntl_data |>
+  filter(year >= 1984 & year < 1990)
+t.test(full_ntl_data_1980s$union_wage_2020_adj, full_ntl_data_1980s$nonunion_wage_2020_adj)
+
+# 1990s
+full_ntl_data_1990s <- full_ntl_data |>
+  filter(year >= 1990 & year < 2000)
+t.test(full_ntl_data_1990s$union_wage_2020_adj, full_ntl_data_1990s$nonunion_wage_2020_adj)
+
+# 2000s
+full_ntl_data_2000s <- full_ntl_data |>
+  filter(year >= 2000 & year < 2010)
+t.test(full_ntl_data_2000s$union_wage_2020_adj, full_ntl_data_2000s$nonunion_wage_2020_adj)
+
+# 2010s
+full_ntl_data_2010s <- full_ntl_data |>
+  filter(year >= 2010 & year < 2020)
+t.test(full_ntl_data_2010s$union_wage_2020_adj, full_ntl_data_2010s$nonunion_wage_2020_adj)
+```
+
+T test between median and mean incomes
+```{r}
+# 1984 to 2019
+t.test(full_ntl_data$median_income_2020_adj, full_ntl_data$mean_income_2020_adj)
+
+# 1980s
+full_ntl_data_1980s <- full_ntl_data |>
+  filter(year >= 1984 & year < 1990)
+t.test(full_ntl_data_1980s$median_income_2020_adj, full_ntl_data_1980s$mean_income_2020_adj)
+
+# 1990s
+full_ntl_data_1990s <- full_ntl_data |>
+  filter(year >= 1990 & year < 2000)
+t.test(full_ntl_data_1990s$median_income_2020_adj, full_ntl_data_1990s$mean_income_2020_adj)
+
+# 2000s
+full_ntl_data_2000s <- full_ntl_data |>
+  filter(year >= 2000 & year < 2010)
+t.test(full_ntl_data_2000s$median_income_2020_adj, full_ntl_data_2000s$mean_income_2020_adj)
+
+# 2010s
+full_ntl_data_2010s <- full_ntl_data |>
+  filter(year >= 2010 & year < 2020)
+t.test(full_ntl_data_2010s$median_income_2020_adj, full_ntl_data_2010s$mean_income_2020_adj)
+```
+
+Regression between percent union and party state average
+```{r}
+# 1984 to 2019
+summary(lm(full_state_data$party_state_average~full_state_data$percent_union))
+
+# 1980s
+full_state_data_1980s <- full_state_data |>
+  filter(year >= 1984 & year < 1990)
+summary(lm(full_state_data_1980s$party_state_average~full_state_data_1980s$percent_union))
+
+# 1990s
+full_state_data_1990s <- full_state_data |>
+  filter(year >= 1990 & year < 2000)
+summary(lm(full_state_data_1990s$party_state_average~full_state_data_1990s$percent_union))
+
+# 20000s
+full_state_data_2000s <- full_state_data |>
+  filter(year >= 2000 & year < 2010)
+summary(lm(full_state_data_2000s$party_state_average~full_state_data_2000s$percent_union))
+
+# 2010s
+full_state_data_2010s <- full_state_data |>
+  filter(year >= 2010 & year < 2020)
+summary(lm(full_state_data_2010s$party_state_average~full_state_data_2010s$percent_union))
+```
+
 ## Conclusion

### 2024-01-11 Documents/Paper.qmd
code about loading and cleaning data

---
title: "Your Substantive Title Here"
author: 
  - Your Name1^[American University]
  - Your Name2^[American University]
format: 
  pdf:
    number-sections: true
    toc: false
    geometry: 
      - top=3cm
date: 2023-09-27
date-format: iso
documentclass: article
editor: source
abstract: "This is our informative abstract of fewer than 200 words. It describes what we investigate, how we investigate it, and what we find."
bibliography: main.bib
---

```{r setup}
#| echo: false
#| message: false

# This chunk loads any packages we need.
# Because "message: false", any R messages from this chunk do not appear in our paper.

library(tidyverse)
library(usmap)
```

```{r}
#| echo: false
#| message: false

# This chunk might read our data.
# It might clean the data, create new variables, etc.
# Now our data are ready for our paper.

# Because echo: false, this chunk itself is not shown.

# Union Data: obtained from Tidy Tuesday
union_wages <- read_rds("../Data/union_wages.rds")
union_states <- read_rds("../Data/union_states.rds")

# Election Data: Obtained from Tidy Tuesday
house <- read_rds("../Data/house.rds")

# Wage Data: Obtained from Tidy Tuesday
income_distribution <- read_rds("../Data/income_distribution.rds")

# Minimum Wage Data: Obtained from U.S. Department of Labor
minimum_wage <- read_csv("../Data/minimum_wage_data.csv",show_col_types = FALSE)

# Clean house election data
house <- house |>
  group_by(year, state, district) |> # group for next line
  filter(candidatevotes == max(candidatevotes)) |> #only include candidates that won
  ungroup() |> 
  mutate(state = str_to_sentence(state)) |> # change from all caps to sentence case) 
  select(!(state_fips:office)) |> #Remove unwanted columns
  select(!(stage:candidate)) |> 
  select(!(writein:fusion_ticket)) |> 
  rename(state_abbreviation = state_po) |>  # rename state abbreviation column
  mutate(party = case_when(
    str_detect(party, "DEMOCRAT") ~ "DEMOCRAT",
    str_detect(party, "REPUBLICAN") ~ "REUBLICAN",
    str_detect(party, "INDEPENDENT") ~ "INDEPENDENT"
  )) # Recode of party variable

# Clean union wages data
union_wages <- union_wages |> 
  filter(facet == "all wage and salary workers") |> 
  distinct(year, .keep_all = TRUE) |> 
  select(-sample_size,
         -at_cap) |> 
  mutate(wage_percent_difference = union_wage_premium_raw, .keep = "unused") |> 
  mutate(wage_percent_difference_adjusted = union_wage_premium_adjusted, .keep = "unused") |> 
  mutate(union_type = facet, .keep = "unused")

# Clean union states data
union_states <- union_states |> 
  select(-state_census_code,
         -observations) |> # remove unwanted columns
  mutate(employment_count = employment * 1000, .keep = "unused") |> # column measured in 1000s, this fixes it
  mutate(union_employment_count = members * 1000, .keep = "unused") |> 
  mutate(collective_bargain_employment_count = covered * 1000,
         .keep = "unused") |> 
  rename(
    percent_union = p_members,
    percent_collective_bargain = p_covered
  ) #renamed columns

# Clean income distribution data
income_distribution <- income_distribution |> 
  select(-number) |> # removing unwanted columns
  select(!ends_with("moe")) |> 
  select(year:income_mean) |> 
  distinct(year, race, income_median, income_mean, .keep_all = TRUE) |> # only include one line per year. needed bc of multiple sectors that are not being used in data analysis
  rename(median_income = income_median,
         mean_income = income_mean) |> # renaming columns
  filter(race == "All Races") |> 
  select(-race)

# Clean minimum wage data
minimum_wage <- minimum_wage |> 
  select(Year:CPI.Average) |> # selecting relevant columns
  rename(year = Year,
         state = State,
         state_min_wage = State.Minimum.Wage,
         state_min_wage_2020_adj = State.Minimum.Wage.2020.Dollars,
         fed_min_wage = Federal.Minimum.Wage,
         fed_min_wage_2020_adj = Federal.Minimum.Wage.2020.Dollars,
         effective_min_wage = Effective.Minimum.Wage,
         effective_min_wage_2020_adj = Effective.Minimum.Wage.2020.Dollars,
         CPI_avg = CPI.Average
  ) |> # renaming variables
  filter(!state %in% c("U.S. Virgin Islands", "Puerto Rico", "Guam")) |>  # remove territory
  mutate(state_abbreviation = 
           ifelse(is.na(state.abb[match(state, state.name)]),
                  "DC",
                  state.abb[match(state, state.name)]),
         .keep = "unused", .after = year) # replaces state with state abbreviation, NA values coded as DC as DC is the only NA value 
```

```{r readdata}
#| eval: false
#| echo: false

# Because eval: false, this chunk is not run.

# Joining data sets to create full state data
full_state_data <- union_states |> 
  filter(sector == "Total") |> 
  select(-sector,
         -state) |> 
  left_join(house, by = c("year", "state_abbreviation")) |> 
  mutate(decade = paste0((year - year %% 10), "s"),
         party_binary = ifelse(party == "DEMOCRAT", 1, 0)) |> 
  left_join(minimum_wage, by = c("year", "state_abbreviation")) |> 
  group_by(year, state) |> 
  mutate(party_state_average = mean(party_binary)) |>
  ungroup() |> 
  mutate(state = state.name[match(state_abbreviation, state.abb)], # fix state encoding
         regions = state.region[match(state_abbreviation, state.abb)]) # fix region encoding

state_column_list = c("year", "decade", "state", "state_abbreviation",
                      "regions", "district", "party", "party_binary",
                      "party_state_average",
                      "employment_count", "union_employment_count",
                      "collective_bargain_employment_count", "percent_union",
                      "percent_collective_bargain", "state_min_wage",
                      "state_min_wage_2020_adj", "fed_min_wage",
                      "fed_min_wage_2020_adj", "effective_min_wage",
                      "effective_min_wage_2020_adj", "CPI_avg")
full_state_data <- full_state_data[, state_column_list] # reorder columns
full_state_data$party[is.na(full_state_data$party)] <- "No election" # replace party NAs
full_state_data$district[is.na(full_state_data$district)] <- "No election" # replace district NAs
full_state_data$party_binary[is.na(full_state_data$party_binary)] <- "No election" # replace district NAs

# Joining data sets to create full national data
state_to_ntl_adjusted <- full_state_data |>
  mutate(prop_2020_adj = fed_min_wage_2020_adj / fed_min_wage, .after = year) |>
  distinct(year, .keep_all = TRUE) |>
  select(year, prop_2020_adj)

state_to_ntl_employment <- full_state_data |>
  group_by(year) |>
  summarise_at(vars(employment_count:collective_bargain_employment_count), sum)

full_ntl_data <- income_distribution |> 
  left_join(union_wages, by = "year") |>
  left_join(state_to_ntl_adjusted, by = "year") |>
  left_join(state_to_ntl_employment, by = "year") |>
  mutate(decade = paste0((year - year %% 10), "s"), .after = year) |>
  mutate(mean_median_income_difference = mean_income - median_income, .after = mean_income) |>
   mutate(median_income_2020_adj = median_income * prop_2020_adj,
         mean_income_2020_adj = mean_income * prop_2020_adj,
         mean_median_income_difference_2020_adj = mean_median_income_difference * prop_2020_adj,
         wage_2020_adj = wage * prop_2020_adj,
         union_wage_2020_adj = union_wage * prop_2020_adj,
         nonunion_wage_2020_adj = nonunion_wage * prop_2020_adj,
         percent_union = union_employment_count / employment_count,
         percent_collective_bargain = collective_bargain_employment_count / employment_count) |> 
  select(-prop_2020_adj)

ntl_column_list <- c("year", "decade", "mean_income",
                     "mean_income_2020_adj", "median_income",
                     "median_income_2020_adj", "mean_median_income_difference",
                     "mean_median_income_difference_2020_adj", "wage",
                     "wage_2020_adj", "union_wage", "union_wage_2020_adj",
                     "nonunion_wage", "nonunion_wage_2020_adj",
                     "wage_percent_difference", 
                     "wage_percent_difference_adjusted",
                     "employment_count", "union_employment_count", 
                     "collective_bargain_employment_count",
                     "percent_union", "percent_collective_bargain",
                     "union_type")
full_ntl_data <- full_ntl_data[, ntl_column_list] # reorder columns

rm(house, income_distribution, minimum_wage, state_to_ntl_adjusted,
   state_to_ntl_employment, union_states, union_wages, state_column_list,
   ntl_column_list)
```

# Introduction

In this section, we introduce the reader to the phenomenon we investigate. We describe the way in which our analysis contributes to an important intellectual debate, or how it answers a pressing political or social question. We introduce our research question, hypotheses, data, and results. We signpost for the reader what's coming in the rest of the paper.

We remember that our paper is not a mystery novel. We note our core results early and often.

Throughout our paper, we use active, first-person language and avoid the passive voice. For example, we write "we examine the relationship between $X$ and $Y$"; we do not write "the relationship between $X$ and $Y$ was examined." Where we do the analysis, we speak about it transparently. We use the present tense; for example, "In this paper, we argue \ldots" and "Paper XYZ demonstrates the relationship between \ldots".

# [Our Substance and Context Section Title Here]

Here we go deeper into the intellectual debate, the political and social context of our investigation. To give the reader a clear sense of why we are writing this paper, we describe the relevant scholarly, technical, or popular literature. We give this section a meaningful _substantive_ title; it is not entitled "Literature Review", for example. We cite at least three published, peer-reviewed scholarly works. For example, we could cite @mooree20 or @moorav12^[There should always be a space before the "(" in a citation date.], which we discussed in class.^[To cite a paper within parentheses, use, e.g., [@moore12].] We only cite others' work in our paper when it enhances the reader's understanding of what we, the authors of this paper, are doing.  We connect everything we cite to _our_ investigation; this is our original research, not a book report or an annotated bibliography.

We do not cite paper titles or journal names, unless our paper is about someone else's paper or about the set of articles in a journal. We do not cite authors' first, given names. We can refer to either what an author does, or what a paper does, but we should be consistent. For example, "@moorav12 argue that we should \ldots" refers to what the authors do; "@moorav12 argues that we should \ldots" refers to what the paper -- @moorav12 -- does.

In order to integrate citations into the References section below, we add entries into our file `main.bib`. This is a plain-text file that we edit in RStudio. We store `main.bib` in the same folder as our paper's `.qmd` and `.pdf` files. Its entries are formatted so that they can be knit to `.pdf`; see [https://j.mp/2UzTXEZ](https://www.overleaf.com/learn/latex/Bibliography_management_with_bibtex#The_bibliography_file) for example entries for articles, books, and miscellaneous. We can get these entries automatically from Google Scholar by turning on BibTeX in the Google Scholar Settings - Bibliography Manager. Perhaps we use a tool like free, open-source BibDesk to help us manage the `.bib` file.

# Data and Methods {#sec-data}

This section describes the data we analyze. We describe the source of the data,
and its primary features. We cite our data. We describe the methods we use to
answer our question and to test our hypotheses.

If our data were `cars`, loaded in a chunk above, we note that our data have `r nrow(cars)` observations. Our unit of analysis is the cars; each row represents a different car that was measured.

We refer to concepts and label them appropriately. We state our outcome and how it is measured: "Our outcome is stopping distance measured in feet." We state our key predictor and how it is measured: "Our key predictor is the car's speed, measured in miles per hour."

We almost never refer to specific variable, object, function, or data frame names (such as `var_x`, `ourdata`, `this_useful_func`, or `df`). These particular names are almost never of interest or use to the reader.

We explain important decisions and codings in this section. "We collect our data from source X. We code the outcome as 1 if the registrant turned out in the 2022 election, 0 if they did not." A table can serve as an efficient way to detail several such decisions.

Where there are less-critical details that we implement to improve our analysis or presentation, we do not explain them in the paper. Our paper does not say, for example, "we reorder the levels of the factor variable from alphabetical 'high, low, medium' to the sensible 'low, medium, high.'" Of course, we implement that reordering, but it should not appear in our paper.

We cite the software we use. For example, we conduct our analysis using R version `r paste(R.version$major, R.version$minor, sep = ".")` [@rcoreteam23]. We rely on several elements of the `tidyverse` [@wickhamtidyverse]. Of course, we do not cite software that we do not use.

# [Our Results Section Title Here]

Here, we explain and interpret our results. We try to learn as much as we can about our question as possible, given the data and analysis. We present our results clearly. We interpret them for the reader with precision and circumspection. We avoid making claims that are not substantiated by our data. We are careful about causality. When we describe associations, we avoid language like "effects" and "increases"; we only describe "effects" or "impacts" when we have a causally well-identified research design.

Note that this section may be integrated into @sec-data, if joining the two improves the overall presentation.

## Predicting Distance with Speed

Our results for the `cars` data include estimating the linear model

$$\text{Distance}_i = \beta_0 + \beta_1 (\text{Speed}_i) + \epsilon_i.$$
Perhaps we start by plotting the data, as in @fig-cars.

```{r fig-cars}
#| echo: false
#| fig-cap: Distance on Speed

ggplot(cars, aes(speed, dist)) +
  geom_point()
```

The data may be roughly linear, though there may be some non-linearity we should incorporate.

```{r linearmodel}
#| echo: false

# Estimate a linear model:
lm_out <- lm(dist ~ speed, data = cars)
# Extract the coefficient on speed:
cars_speed_coef <- coef(lm_out)["speed"]
```

Below we show the model estimates. The first table uses `xtable()`, the second uses `stargazer()` [@hlavac18].

```{r linearxtable}
#| echo: false
#| message: false
#| results: 'asis'

# We can print regression tables with xtable or stargazer:
regr_table <- xtable::xtable(lm_out,
                             digits = 2,
                             caption = "Our Informative Caption")

print(regr_table, comment = FALSE)
```

```{r linearstargazer}
#| echo: false
#| message: false
#| results: 'asis'

# We can print regression tables with xtable or stargazer:
stargazer::stargazer(lm_out, 
                     title = "Our Informative Title",
                     dep.var.caption = "Outcome",
                     digits = 2,
                     header = FALSE)
```

Using the `cars` data, we find that each unit of speed is associated with `r round(cars_speed_coef, 1)` more units of distance. We draw out what this really means, and what it implies. For example, if a typical difference among our observations is 7 units of speed, then our model estimates that a typical difference in distance among our observations is $7 \times `r round(cars_speed_coef, 1)` = `r 7 * round(cars_speed_coef, 1)`$ units of distance. We describe the substantive relevance of this number.

We do not report estimates like `p = 3.242e-15`, since these are computational zeros. Instead, we write $p < 0.001$ or $p \approx 0$, as appropriate.

We do not report quantities to unhelpful degrees of precision. Although there were 112,030,874 votes cast from voting-eligible population of 242,690,810 in the U.S. in 2022, it is not helpful to report turnout as `r (112030874 / 242690810) * 100`%; writing `r round((112030874 / 242690810) * 100, 1)`% suffices.

## Comparing Distances between High- and Low-Speed Cars

```{r ttest}
#| echo: FALSE

# Create a binary variable for "high-speed":
cars <- cars |> mutate(hs = ifelse(speed > mean(speed), 1, 0))

# Conduct t-test:
t_out <- t.test(dist ~ hs, cars)

t_lower_ci <- t_out$conf.int[1] |> round(2)
t_upper_ci <- t_out$conf.int[2] |> round(2)
```

To report the results of a $t$-test, we do so in text, and perhaps in a well-formatted table as well, such as @tbl-ttesttable. Here, as above, we report the important details in text. For example, when we define "high-speed" cars as those traveling above the mean speed, the difference between the high-speed and low-speed group means is `r t_out$estimate[2] - t_out$estimate[1]`, with a 95% confidence interval that covers $(`r -t_upper_ci`, `r -t_lower_ci`)$.

```{r tbl-ttesttable}
#| echo: FALSE
#| warning: FALSE
#| tbl-cap: Distance by Speed Group

parameters::model_parameters(t_out) |> 
  parameters::print_md(
    footer = "")
```

If I have tests of two outcomes from the same data, I can bind them together, as in @tbl-ttesttable2:

```{r tbl-ttesttable2}
#| echo: FALSE
#| warning: FALSE
#| tbl-cap: Distance and Square-root Distance by Speed Group

cars <- cars |> mutate(dist_sqrt = (sqrt(dist)))
t_out_2 <- t.test(dist_sqrt ~ hs, data = cars)

tab <- rbind(
  model_parameters(t_out),
  model_parameters(t_out_2)
)

tab |> 
  print_md(
    footer = "")
```

# Discussion

We remind the reader what this paper was about, why it was important, and what we found. We reflect on limitations of the data or methods. If we have specific advice for someone picking up where we leave off, we provide that guidance. We avoid making trite statements like "more research should be done".

\clearpage

# References

### 2024-01-11 Documents/Paper.qmd
added names to report

@@
 author: 
-  - Your Name1^[American University]
-  - Your Name2^[American University]
+  - Melanie Klein^[American University]
+  - Michael Taffe^[American University]
+  - Ramiro Arevalo^[American University]
+
 format: 

### 2024-01-11 Code/R FIle.qmd
fixed graph labels

@@
     title = "Mean and Median Income in the United States Between 1984 and 2019",
-    subtitle = "Using mean (red) and median (blue) incomes adjusted to 2020 rates",
+    subtitle = "Using mean (red) and median (blue) incomes",
     x = "Year",
-    y = "Income in USD (Adjusted for Inflation to 2020 Rates)"
+    y = "Income in USD"
   )

### 2024-01-11 Code/Presentation.qmd
filled out data manipulation and EDA of presentation

@@
     theme: simple
+    monofont: "Courier New"
+    monofontoptions: "size=10%"
 editor: visual
@@
 
-## About the Data
+## About the Data pt. 1 {.smaller}
 
 * **Union data**
-  * Gathered from Tidy Tuesday, although the original 
+  * Gathered from Tidy Tuesday, although the original data comes from the Union Membership and Coverage Database at Georgia state university
+  * For our project, the relevant tables include one that details union membership counts by state by year and another that tracks union and non union wages by year.
 * **House election data**
-  *
+  * Also gathered from Tidy Tuesday's repository, this data originally came from the MIT Election Data and Science Lab
+  * The variables we selected from the data set were year, state, district, and election winner
+
+## About the Data pt. 2 {.smaller}
 * **Wage**
-  *
+  * Our wage data was obtained through the U.S. Department of Labor and Tidy Tuesday (originally from U.S. census data)
+  * This data tracked the mean and median income nationally in the United States
+  * Another table tracked minimum wage at the state level
 
-## EDA
+## Data manipulation pt. 1
+* All of our data had undergone basic cleaning before we downloaded it so all of the columns were correctly formatted and there was almost 0 NA values
+* each data set was pared down so that only the variables relevant to our project was completed.
+* for our house election data, the table was manipulated so that only rows of candidates who won their election remained.
+  * other minor changes included renaming variables and recoding a few outlier observations in the `party` variable
+* For the union data, we filtered the datasets down to not include union sectors (ex. construction-unions) and added columns to adjust for inflation
+* For our wage data, we filtered out race, as this was outside of the scope of our study.
 
-When you click the **Render** button a presentation will be generated that includes both content and the output of embedded code. You can embed code like this:
+## Data Manipulation pt. 2
+* Our last data manipulation was combining all of our tables into two data sets: one that stores data at a national level, and another that stores data at a state level.
+* some of the variables were only provided at the state level so they were summarized and added to the national data set.
 
-```{r}
-1 + 1
-```
+## EDA pt. 1: Overview
+
+* EDA was split into five sections: wage data, union data, house data, wage data's associations with union data, and union data's association with house data.
+We did not have the relevent wage data to compare wages and house elections
+
+## EDA pt. 2: Wage Data
+:::: {.columns}
+
+::: {.column width="60%"}
+* Minimum wages, which had been stagnant since 2009, began to grow again in 2016, although the federal rate has remained stagnant
+* Incomes, when adjusted for inflation, have been decreasing since 1984. Surprisingly, our data shows income inequality has remained stagnant over the last forty years.
+:::
+
+::: {.column width="40%"}
+column 2
+:::
+
+::::
+
+## EDA pt. 3: Union Data
+:::: {.columns}
+
+::: {.column width="60%"}
+* Union and collective bargaining rates have collapsed over the last 40 years, although may just beginning to see a come back in the last two.
+* Southern states have much less union activity than northern and western states.
+:::
+
+::: {.column width="40%"}
+column 2
+:::
+
+::::
+
+## EDA pt. 4: House Data 
+:::: {.columns}
+
+::: {.column width="60%"}
+* The country has generally grown more republican in the last 40 years, although data in the last 10 years show the pendulem might now be swinging the other way
+  * Northern-midwest states have had the biggest shift to the right. New England is the only region that has seen a shift left. 
+:::
+
+::: {.column width="40%"}
+column 2
+:::
+
+::::
+
+## EDA pt. 5: Wage Data's Interactions With Union Data
+:::: {.columns}
+
+::: {.column width="60%"}
+* From the graph, it appears that a reduction in the amount of employees covered by a union leads to an increase in wages.
+However, for individual workers, being part of a union remains a way to increase wages.
+* We do not have data to compare wages and union rates by state, which would likely tell a different story.
+:::
+
+::: {.column width="40%"}
+column 2
+:::
+
+::::
+
+## EDA pt. 6: House Election Data's Interactions With Union Data
+:::: {.columns}
+
+::: {.column width="60%"}
+* Unions have become a very political topic in the last 40 years
+* This trend appears to be driven by the northeast and south
+:::
+
+::: {.column width="40%"}
+column 2
+:::
+
+::::
 
@@
 ## Where we would take our project in the future
+
+##

### 2024-01-11 Code/R FIle.qmd

@@
 
+```{r}
+summary(income_distribution)
+```
+In this table, we can see that there are a few NA values in the tables, and some of the variables have a noticable skew. After some basic data cleaning is done we will see if these values remain
+
+```{r}
+summary(minimum_wage)
+```
+Only one column has NA values in this data frame, and in that column only 15 out of 2862 observations are missing values. Again, we will see if these values remain after our data cleaning
+
 ### 2.2.2 Cleaning data script
@@
   rename(state_abbreviation = state_po) |>  # rename state abbreviation column
-  mutate(party = case_when(
+  mutate(party = case_when( # Recode of party variable
     str_detect(party, "DEMOCRAT") ~ "DEMOCRAT",
@@
     str_detect(party, "INDEPENDENT") ~ "INDEPENDENT"
-  )) # Recode of party variable
+  )) |> 
+  filter(state_abbreviation != "DC") # Removes the single line of DC data
 ```
@@
 ```
+```{r}
+colSums(is.na(house))
+colSums(is.na(income_distribution))
+colSums(is.na(minimum_wage))
+colSums(is.na(union_wages))
+colSums(is.na(union_states))
+```
+There are no more NA values
 
@@
 
-The states data set has five entries for each year for each of the five `sectors`. For this project we will use the `total` sector. A decades column will also be added.
+The states data set has five entries for each year for each of the five `sectors`.
+For this project we will use the `total` sector.
+A decades column will also be added.
+Finally, we will filter all of the data to only include data between 1984 and 2019 as these are the common data years between all of the tables
 
@@
   mutate(state = state.name[match(state_abbreviation, state.abb)], # fix state encoding
-         regions = state.region[match(state_abbreviation, state.abb)]) # fix region encoding
+         regions = state.region[match(state_abbreviation, state.abb)]) |>  # fix region encoding
+  filter(year >= 1984 & year <= 2019) |> 
+  filter(state_abbreviation != "DC") # Removes the single line of DC data
 ```
@@
 ```{r}
-range(full_state_data$year)
-```
-
-```{r}
 head(full_state_data)
@@
 ```
+After combining the dataset, we can see that not all of the variables are tracked every year, causing there to be NA values. We will keep this in mind as we continue our data analysis.
 
@@
          percent_collective_bargain = collective_bargain_employment_count / employment_count) |> 
-  select(-prop_2020_adj)
+  select(-prop_2020_adj) |> 
+  filter(year >= 1984 & year <= 2019)
 ```
@@
 ```{r}
-range(full_ntl_data$year)
+head(full_ntl_data)
+summary(full_ntl_data)
 ```
+There are no NA values in our national data set.
 
@@
 per_union_wage_data <- full_state_data |> # This object will be used to create the graph and also in the data subset
-  filter(!is.na(regions)) |> #This just removes DC
-  filter(year <= 2020 & year >= 1984) |>  #wage data only exists before 2020
   mutate(
@@
   scale_x_continuous(breaks = seq(min(per_union_wage_data$year), max(per_union_wage_data$year), by = 4)) + 
-  labs(title = "States By % of Labor Market Covered By Union",
+  labs(title = "States by % of Labor Market Covered by Union",
        x = "Year",
@@
 ```
+
 Our first graph shows what is commonly known about unions in the United States: Union membership has been on the decline since at least the 1980s.
-Unfortunately our data does not extend to the last two years where some reports indicate union memberships are on the rise.
-Early on, we can see that southern and more Republican leading states have lower union rates than more democrat and northern / western states.
+Unfortunately our data does not extend to the last four years where some reports indicate union membership is on the rise.
+Early on, we can see that southern and more Republican leaning states have lower union rates than more Democrat and northern / western states.
 There is a very small bump in membership in 2008, which we hypothesize happened in relation to the 2008 financial crisis.
@@
 per_collective_wage_data <- full_state_data |> # This object will be used to create the graph and also in the data subset
-  filter(!is.na(regions)) |> #This just removes DC
-  filter(year <= 2020 & year >= 1984) |>  #wage data only exists before 2020
   mutate(
@@
   scale_x_continuous(breaks = seq(min(per_collective_wage_data$year), max(per_collective_wage_data$year), by = 4)) + 
-  labs(title = "States by % of Labor Market Covered By a Collective Bargaining Agreement",
+  labs(title = "States by % of Labor Market Covered by a Collective Bargaining Agreement",
        x = "Year",
@@
 ```
-This graph is almost the same exact as the first, although the rates are slightly higher because this graph includes all collective bargaining agreements.
+
+This graph is almost the same as the previous plot, although the rates are slightly higher because this graph includes all collective bargaining agreements.
 
@@
 ```{r}
+full_ntl_data |> 
+  filter(year >= 1984) |> 
+  ggplot(aes(x = year)) +
+  geom_line(aes(y = mean_income), color = "red") +
+  annotate("text", x=1995, y=85000, label= "Mean income", color = "red", size = 5) +
+  geom_line(aes(y = median_income), color = "blue") +
+  annotate("text", x=1995, y=65000, label= "Median income", color = "blue", size = 5) +
+  scale_x_continuous(breaks = seq(min(full_ntl_data$year), max(full_ntl_data$year), by = 4)) + 
+  theme_minimal() +
+  labs(
+    title = "Mean and Median Income in the United States Between 1984 and 2019",
+    x = "Year",
+    y = "Income in USD (Adjusted to 2020 values)"
+  )
+```
+
+```{r}
+full_ntl_data |> 
+  filter(year >= 1984) |> 
+  ggplot(aes(x = year)) +
+  geom_line(aes(y = mean_income_2020_adj), color = "red") +
+  annotate("text", x=1995, y=145000, label= "Mean income", color = "red", size = 5) +
+  geom_line(aes(y = median_income_2020_adj), color = "blue") +
+  annotate("text", x=1995, y=115000, label= "Median income", color = "blue", size = 5) +
+  scale_x_continuous(breaks = seq(min(full_ntl_data$year), max(full_ntl_data$year), by = 4)) + 
+  theme_minimal() +
+  labs(
+    title = "Mean and Median Income in the United States Between 1984 and 2019",
+    x = "Year",
+    y = "Income in USD (Adjusted to 2020 values)"
+  )
+```
+
+The two above graphs show that when accounting for inflation, Americans are earning less per year now than they did 40 years ago.
+As we will see later on when examining wage data at a national level, there are numerous external reasons for this.
+Most notably, an aging population that is no longer working is likely driving the average down.
+It is possible that the average worker is earning more but we do not have the data to verify this.
+
+```{r}
 state_minimum_wage_data <- full_state_data |> # This object will be used to create the graph and also in the data subset
-  filter(!is.na(regions)) |> #This just removes DC
-  filter(year <=2020 & year >=1984) |>  #wage data only exists before 2020
   mutate(
@@
             linewidth = .75) +
-  geom_line(aes(y = fed_min_wage_2020_adj), color = "red", linewidth = .75) + # federal minimum wage
+  geom_line(aes(y = fed_min_wage_2020_adj), color = "orange", linewidth = 1.5) + # federal minimum wage
+  annotate("text", x=1990, y=6, label= "Federal Avg", color = "orange", size = 5) +
   theme_minimal() +
@@
        x = "Year",
-       y = "Minimum Wage (2020 USD)",
+       y = "Minimum Wage (Adjusted to 2020 values)",
        color = "Selected States")
@@
 ```
+
 This graph can be used to draw a number of interesting inferences.
@@
 This will be examined in greater detailed in the next graph.
-Next, we can see tht there is a significan shift post 2016 of raising the minimum wage at the state level.
+Next, we can see that there is a significant shift post 2014 of raising the minimum wage at the state level.
 Before that time almost no state had a adjusted minimum wage over $10.
 Now, there are numerous states that surpass this value.
-Next, we can see that the federal minimum wage has been slowly losing value due to inflation since it was last raised in 2009.
+Final, we can see that the federal minimum wage has been slowly losing value due to inflation since it was last raised in 2009.
 
@@
 effective_minimum_wage_data <- full_state_data |> 
-  filter(!is.na(regions)) |> 
   filter(effective_min_wage_2020_adj != fed_min_wage_2020_adj) |> 
-  filter(year <=2020 & year >=1984) |>
   mutate(
@@
             linewidth = .75) +
-  geom_line(aes(y = fed_min_wage_2020_adj), color = "red", linewidth = .75) +
+  geom_line(aes(y = fed_min_wage_2020_adj), color = "orange", linewidth = 1.5) +
+  annotate("text", x=1990, y=6, label= "Federal Avg", color = "orange", size = 5) +
   theme_minimal() +
@@
        x = "Year",
-       y = "Effective minimum wage (2020 USD)",
+       y = "Effective Minimum Wage (Adjusted to 2020 values)",
        color = "Selected States")
@@
 ```
+
 This graph removes all data points where the state minimum wage falls below the federal amount.
-From this we can see that more and more states are taking it into their own hands to protect workers rights.
+From this we can see that more states are taking it into their own hands to protect workers rights.
 We can see a large jump in 2007 which may be related to the financial crisis.
-
-```{r}
-full_ntl_data |> 
-  filter(year >= 1984) |> 
-  ggplot(aes(x = year)) +
-  geom_line(aes(y = mean_income_2020_adj), color = "red") +
-  geom_line(aes(y = median_income_2020_adj), color = "blue") +
-  theme_minimal() +
-  labs(
-    title = "Mean and Median Income in the United States Between 1984 and 2019",
-    subtitle = "Using mean (red) and median (blue) incomes adjusted to 2020 rates",
-    x = "Year",
-    y = "Income in USD (Adjusted for Inflation to 2020 Rates"
-  )
-```
-
-```{r}
-full_ntl_data |> 
-  filter(year >= 1984) |> 
-  ggplot(aes(x = year)) +
-  geom_line(aes(y = mean_income), color = "red") +
-  geom_line(aes(y = median_income), color = "blue") +
-  theme_minimal() +
-  labs(
-    title = "Mean and Median Income in the United States Between 1984 and 2019",
-    subtitle = "Using mean (red) and median (blue) incomes",
-    x = "Year",
-    y = "Income in USD"
-  )
-```
-
-The two above graphs show that when accounting for inflation, Americans are earning less per year now than they did 30 years ago.
+There is another spike around 2014.
 
@@
 plot_usmap(data = map_data, values = "party_state_average", color = "black") +
-  facet_wrap(~year, nrow = 2, ncol= 6) +
-  theme(legend.position = "top") +
-  scale_fill_gradient(low = "red", high = "blue")
-
+  facet_wrap(~year, nrow = 4, ncol= 3) +
+  theme(legend.position = "bottom") +
+  scale_fill_gradient(low = "red", high = "blue", name = "Proportion") +
+  labs(
+    title = "Proportion of House Delegation that is Democratic"
+  )
 rm(map_data)
 ```
-We can see that, especially in the northern-midwest states, that the country has grown to be more republican in the last 30 years.
+
+We can see that, especially in the northern-midwest states, that the country has grown to be more republican in the last 40 years.
 However, in the last 10 years this has begun to shift again.
-During this whole time, it appears that New England has become more democrat.
+During this whole time, it appears that New England has become more Democrat.
 
@@
 full_ntl_data |> 
-  filter(year >= 1984) |> 
   ggplot(aes(x = year)) +
   geom_line(aes(y = union_wage_2020_adj), color = "blue") +
+  annotate("text", x=1993, y=28, label= "Union Wages", color = "blue", size = 5) +
   geom_line(aes(y = nonunion_wage_2020_adj), color = "red") +
-  theme_minimal()
+  annotate("text", x=1993, y=22, label= "Non Union Wages", color = "red", size = 5) +
+  scale_x_continuous(breaks = seq(min(full_ntl_data$year), max(full_ntl_data$year), by = 4)) + 
+  theme_minimal() +
+  labs(
+    title = "Union vs. Non Union Wages Over Time",
+    x = "Year",
+    y = "Wages (Adjusted to 2020 values)"
+  )
 ```
+
 From our income graphs, this graph is rather unintuitive, as we would expect adjusted wages to be decreasing in this time.
@@
 full_ntl_data |> 
-  filter(year >= 1984) |> 
   ggplot(aes(x = percent_collective_bargain)) +
-  geom_point(aes(y = wage_2020_adj), color = "blue") +
+  geom_point(aes(y = nonunion_wage_2020_adj), color = "blue") +
   geom_jitter(aes(y = union_wage_2020_adj), color = "red") +
-  facet_wrap(~decade) 
+  theme_light() +
+  facet_wrap(~decade) +
+  labs(
+    title = "Union vs. Non-Union Wages by Decade",
+    subtitle = "Red represents union wages and blue represents non-union wages",
+    x = "Percent of Workers Under Collective Bargaining Agreement",
+    y = "Wages (Adjusted to 2020 values)"
+
+  )
 ```
@@
 First, a quick view of this graph shows the exact opposite of our hypothesis.
-Looking at each decade we can see collective bargaining rates decrease and wage increase.
+Looking at each decade we can see collective bargaining rates decreasing and wage increasing.
 Then, in each decade that relationship continues.
-Still, we can see that in every decade, a collective bargaining agreement will lead to an increase in wages for a worker over a non-collective-bargaining-jobs.
+Still, we can see that in every decade, a collective bargaining agreement will lead to an increase in wages for a worker over non-collective-bargaining-jobs.
 
 ### 2.3.5 Collective Bargaining and House Elections
-
-```{r}
-full_state_data |> 
-  filter(year >= 1984) |> 
-  ggplot(aes(x = percent_collective_bargain, y = party_state_average)) +
-  geom_point(color = "blue") +
-  geom_smooth(method = "lm", color = "red") +
-  facet_wrap(~decade) +
-  labs(
-    title = "Percent Collective Bargaining vs Proportion of House Delegation that is Democratic",
-    x = "Percent of Workers Covered by Collective Bargaining",
-    y = "Proportion of House Delegation that is Democratic"
-  )
-```
 
@@
 ```{r}
+#| message: FALSE
 full_state_data |> 
-  filter(year <2020 & year >= 1984) |> 
   filter(year %% 4==0) |>
@@
   facet_wrap(~decade) +
+  theme_light() +
   labs(
-    title = "Percent Collective Bargaining vs Proportion of House Delegation that is Democratic",
+    title = "Percent Collective Bargaining vs Proportion of House Delegation That is Democratic",
     x = "Percent of Workers Covered by Collective Bargaining",
@@
 
-This graph demonstrates the same as the previous graph, but with a linear regression line for each region.
-
 ```{r}
+#| message: FALSE
 full_state_data |> 
-  filter(year <2020 & year >= 1984) |> 
   filter(year %% 4==0) |>
   ggplot(aes(x = percent_collective_bargain, y = party_state_average)) +
-  geom_jitter(aes(color = regions), alpha = 0.1, width = 0.1, height = 0.1) +
+  geom_jitter(aes(color = regions), alpha = 0.15, width = 0.1, height = 0.1) +
   geom_smooth(method = "lm", aes(group = regions, color = regions), se = FALSE) +
@@
   labs(
-    title = "Percent Collective Bargaining vs Proportion of House Delegation that is Democratic",
+    title = "Percent Collective Bargaining vs Proportion of House Delegation That is Democratic",
+    subtitle = "Regressed by Region",
     x = "Percent of Workers Covered by Collective Bargaining",
@@
 ```
+These two graphs show very interesting results.
+First, we can see how unions have become politicized in the last 40 years.
+In the 1980s, there was almost no correlation between the amount of workers in a collective bargaining agreement and what the house delegation looked like in that state.
+By the 2010s, this had changed.
+From the second graph, we can see that the North Central and South regions are the main drivers behind this, seen in the steepening of the regression slope.
 
@@
 T-test between union and non-union wages
+
 ```{r}
@@
 T test between median and mean incomes
+
 ```{r}
@@
 Regression between percent union and party state average
+
 ```{r}
-# 1984 to 2019
-summary(lm(full_state_data$party_state_average~full_state_data$percent_union))
-
 # 1980s
@@
   filter(year >= 1984 & year < 1990)
-summary(lm(full_state_data_1980s$party_state_average~full_state_data_1980s$percent_union))
-
-# 1990s
-full_state_data_1990s <- full_state_data |>
-  filter(year >= 1990 & year < 2000)
-summary(lm(full_state_data_1990s$party_state_average~full_state_data_1990s$percent_union))
-
-# 20000s
-full_state_data_2000s <- full_state_data |>
-  filter(year >= 2000 & year < 2010)
-summary(lm(full_state_data_2000s$party_state_average~full_state_data_2000s$percent_union))
+CB_party_1980s_regression <- summary(lm(full_state_data_1980s$party_state_average~full_state_data_1980s$percent_collective_bargain))$coefficients
 
@@
 
+```{r}
+CB_party_1980s_regression <- CB_party_1980s_regression |> 
+  as_tibble() |> 
+  mutate(Reggression_Run = c("Intercept", "percent_collective_bargain"))
+
+CB_party_1980s_regression
+```
+
 ## Conclusion

### 2024-01-11 Code/R FIle.qmd
worked on regression model

@@
 
-T-test between union and non-union wages
+Regression between percent collective bargaining and party state average
+```{r}
+decade_region <- list(
+  c("1980s", "Northeast"),
+  c("1980s", "West"),
+  c("1980s", "North Central"),
+  c("1980s", "South"),
+  c("2010s", "Northeast"),
+  c("2010s", "West"),
+  c("2010s", "North Central"),
+  c("2010s", "South")
+)
 
-```{r}
-# 1984 to 2019
-t.test(full_ntl_data$union_wage_2020_adj, full_ntl_data$nonunion_wage_2020_adj)
+regression_decade_region <- function(decade1, region1) {
+  full_state_data_filtered <- subset(full_state_data, regions == region1 & decade == decade1)
+  model <- lm(full_state_data_filtered$party_state_average~full_state_data_filtered$percent_collective_bargain) 
+  summary(model)$coefficients |> 
+  as_tibble() |> 
+  mutate(Reggression_Run = c("Intercept", "percent_collective_bargain"),
+         Decade = c(decade1, decade1),
+         Region = c(region1, region1),
+         .before = Estimate)
+}
 
-# 1980s
-full_ntl_data_1980s <- full_ntl_data |>
+full_state_data_temp <- full_state_data |>
   filter(year >= 1984 & year < 1990)
-t.test(full_ntl_data_1980s$union_wage_2020_adj, full_ntl_data_1980s$nonunion_wage_2020_adj)
+CB_party_regression <- summary(lm(full_state_data_temp$party_state_average~full_state_data_temp$percent_collective_bargain))$coefficients |> 
+  as_tibble() |> 
+  mutate(Reggression_Run = c("Intercept", "percent_collective_bargain"),
+         Decade = c("1980s", "1980s"),
+         Region = c("all", "all"),
+         .before = Estimate)
 
-# 1990s
-full_ntl_data_1990s <- full_ntl_data |>
-  filter(year >= 1990 & year < 2000)
-t.test(full_ntl_data_1990s$union_wage_2020_adj, full_ntl_data_1990s$nonunion_wage_2020_adj)
+for (x in decade_region) {
+  CB_party_regression <- full_join(CB_party_regression, regression_decade_region(x[1], x[2]),
+                                   by = join_by(Reggression_Run, Decade, Region, Estimate, `Std. Error`, `t value`, `Pr(>|t|)`))
+}
 
-# 2000s
-full_ntl_data_2000s <- full_ntl_data |>
-  filter(year >= 2000 & year < 2010)
-t.test(full_ntl_data_2000s$union_wage_2020_adj, full_ntl_data_2000s$nonunion_wage_2020_adj)
+full_state_data_temp <- full_state_data |>
+  filter(year >= 2010 & year < 2020)
+CB_party_temp <- summary(lm(full_state_data_temp$party_state_average~full_state_data_temp$percent_collective_bargain))$coefficients |> 
+  as_tibble() |> 
+  mutate(Reggression_Run = c("Intercept", "percent_collective_bargain"),
+         Decade = c("2010s", "2010s"),
+         Region = c("all", "all"),
+         .before = Estimate)
 
-# 2010s
-full_ntl_data_2010s <- full_ntl_data |>
-  filter(year >= 2010 & year < 2020)
-t.test(full_ntl_data_2010s$union_wage_2020_adj, full_ntl_data_2010s$nonunion_wage_2020_adj)
-```
+CB_party_regression <- full_join(CB_party_regression, CB_party_temp,
+                                   by = join_by(Reggression_Run, Decade, Region, Estimate, `Std. Error`, `t value`, `Pr(>|t|)`))
 
-T test between median and mean incomes
-
-```{r}
-# 1984 to 2019
-t.test(full_ntl_data$median_income_2020_adj, full_ntl_data$mean_income_2020_adj)
-
-# 1980s
-full_ntl_data_1980s <- full_ntl_data |>
-  filter(year >= 1984 & year < 1990)
-t.test(full_ntl_data_1980s$median_income_2020_adj, full_ntl_data_1980s$mean_income_2020_adj)
-
-# 1990s
-full_ntl_data_1990s <- full_ntl_data |>
-  filter(year >= 1990 & year < 2000)
-t.test(full_ntl_data_1990s$median_income_2020_adj, full_ntl_data_1990s$mean_income_2020_adj)
-
-# 2000s
-full_ntl_data_2000s <- full_ntl_data |>
-  filter(year >= 2000 & year < 2010)
-t.test(full_ntl_data_2000s$median_income_2020_adj, full_ntl_data_2000s$mean_income_2020_adj)
-
-# 2010s
-full_ntl_data_2010s <- full_ntl_data |>
-  filter(year >= 2010 & year < 2020)
-t.test(full_ntl_data_2010s$median_income_2020_adj, full_ntl_data_2010s$mean_income_2020_adj)
-```
-
-Regression between percent union and party state average
-
-```{r}
-# 1980s
-full_state_data_1980s <- full_state_data |>
-  filter(year >= 1984 & year < 1990)
-CB_party_1980s_regression <- summary(lm(full_state_data_1980s$party_state_average~full_state_data_1980s$percent_collective_bargain))$coefficients
-
-# 2010s
-full_state_data_2010s <- full_state_data |>
-  filter(year >= 2010 & year < 2020)
-summary(lm(full_state_data_2010s$party_state_average~full_state_data_2010s$percent_union))
-```
-
-```{r}
-CB_party_1980s_regression <- CB_party_1980s_regression |> 
-  as_tibble() |> 
-  mutate(Reggression_Run = c("Intercept", "percent_collective_bargain"))
-
-CB_party_1980s_regression
+CB_party_regression |>
+  filter(Reggression_Run == "percent_collective_bargain")
 ```

### 2024-01-11 Code/R FIle.qmd
fixed theme

@@
   facet_wrap(~decade) +
+  theme_light() +
   labs(

### 2024-01-12 Code/R FIle.qmd
created graph for regression

@@
 
-Regression between percent collective bargaining and party state average
+t test between union and non union wages
+```{r}
+# 1980s
+full_ntl_data_1980s <- full_ntl_data |>
+  filter(decade == "1980s")
+t.test(full_ntl_data_1980s$union_wage_2020_adj, full_ntl_data_1980s$nonunion_wage_2020_adj)
+
+# 2010s
+full_ntl_data_2010s <- full_ntl_data |>
+  filter(decade == "2010s")
+t.test(full_ntl_data_2010s$union_wage_2020_adj, full_ntl_data_2010s$nonunion_wage_2020_adj)
+```
+
+Linear regression of percent collective bargaining and party state average
 ```{r}
@@
 CB_party_regression <- full_join(CB_party_regression, CB_party_temp,
-                                   by = join_by(Reggression_Run, Decade, Region, Estimate, `Std. Error`, `t value`, `Pr(>|t|)`))
+                                   by = join_by(Reggression_Run, Decade, Region, Estimate, `Std. Error`, `t value`, `Pr(>|t|)`)) |>
+  filter(Reggression_Run == "percent_collective_bargain") |>
+  rename(P_Value = "Pr(>|t|)")
+```
 
-CB_party_regression |>
-  filter(Reggression_Run == "percent_collective_bargain")
+Graph of regression estimate by decade and year
+```{r}
+ggplot(CB_party_regression, aes(x = Decade, y = Estimate, fill = Region)) +
+  geom_bar(stat = "identity", position = "dodge", color = ifelse(CB_party_regression$P_Value > 0.05, "black", NA), size = 0.2) +
+  theme_minimal() +
+  labs(title = "Linear Regression Estimate by Decade",
+       subtitle = "Model of Percent Collective Bargaining vs Proportion of House Delegation That is Democratic")
 ```

### 2024-01-12 Code/R FIle.qmd
fixed graph

@@
 ggplot(CB_party_regression, aes(x = Decade, y = Estimate, fill = Region)) +
-  geom_bar(stat = "identity", position = "dodge", color = ifelse(CB_party_regression$P_Value > 0.05, "black", NA), size = 0.2) +
+  geom_bar(stat = "identity", position = "dodge", color = ifelse(CB_party_regression$P_Value > 0.05, "black", NA), linewidth = 0.2) +
   theme_minimal() +

### 2024-01-12 Code/R FIle.qmd
fixed regression graph labels

--- 2024-01-12 Code/R FIle.qmd
@@
 ggplot(CB_party_regression, aes(x = Decade, y = Estimate, fill = Region)) +
-  geom_bar(stat = "identity", position = "dodge", color = ifelse(CB_party_regression$P_Value > 0.05, "black", NA), size = 0.2) +
+  geom_bar(stat = "identity", position = "dodge") +
+  geom_text(aes(label = sprintf("P=%.3f", P_Value)),
+             position = position_dodge(width = 0.9),
+             vjust = -0.2,
+             size = 3) +
   theme_minimal() +

### 2024-01-12 Code/Full R Code.qmd
moved files around

--- 2024-01-12 Code/R FIle.qmd
@@
   ) 
-per_union_wage_data |> # use created table here
+States_by_union <- per_union_wage_data |> # use created table here
   ggplot(aes(x = year)) +
@@
        color = "Selected States")
-
-rm(per_union_wage_data)
+States_by_union
+ggsave("../Documents/Graph_Images/States_by_union.png", width=7.29, height=5.41)
+rm(per_union_wage_data, States_by_union)
 ```
@@
   ) 
-per_collective_wage_data |> # use created table here
+States_by_CB <- per_collective_wage_data |> # use created table here
   ggplot(aes(x = year)) +
@@
        color = "Selected States")
-
-rm(per_collective_wage_data)
+States_by_CB
+ggsave("../Documents/Graph_Images/States_by_CB.png", width=7.29, height=5.41)
+rm(per_collective_wage_data, States_by_CB)
 ```
@@
 ```{r}
-full_ntl_data |> 
+Income_by_year <- full_ntl_data |> 
   filter(year >= 1984) |> 
@@
   )
+Income_by_year
+ggsave("../Documents/Graph_Images/Income_by_year.png", width=7.29, height=5.41)
+rm(Income_by_year)
 ```
@@
 ```{r}
-full_ntl_data |> 
+Income_by_year_adj <- full_ntl_data |> 
   filter(year >= 1984) |> 
@@
   labs(
-    title = "Mean and Median Income in the United States Between 1984 and 2019",
+    title = "Mean and Median Income in the United States Between 1984 and 2019 (2020 USD)",
     x = "Year",
@@
   )
+
+Income_by_year_adj
+ggsave("../Documents/Graph_Images/Income_by_year_adj.png", width=7.29, height=5.41)
+rm(Income_by_year_adj)
 ```
@@
   ) 
-state_minimum_wage_data |> # use created table here
+Min_wage_by_state <- state_minimum_wage_data |> # use created table here
   ggplot(aes(x = year)) +
@@
 
+Min_wage_by_state
+ggsave("../Documents/Graph_Images/Min_wage_by_state.png", width=7.29, height=5.41)
+rm(Min_wage_by_state)
 rm(state_minimum_wage_data) # remove object to make graph
@@
   ) 
-effective_minimum_wage_data |> 
+Eff_wage_by_state <- effective_minimum_wage_data |> 
   ggplot(aes(x = year)) +
@@
 
+Eff_wage_by_state
+ggsave("../Documents/Graph_Images/Eff_wage_by_state.png", width=7.29, height=5.41)
+rm(Eff_wage_by_state)
 rm(effective_minimum_wage_data)
@@
   filter(year %% 4==0) |> 
-  filter(!is.na(party_state_average))
+  filter(!is.na(party_state_average)) |> 
+  full_join(
+    tibble(
+      "year" = seq(from=1984, to=2016, by=4),
+      "state" = rep("District of Columbia", 9),
+      "party_state_average" = rep(.5, 9)),
+    by = c("year", "state", "party_state_average")
+    )
 
-plot_usmap(data = map_data, values = "party_state_average", color = "black") +
+House_by_year <- plot_usmap(data = map_data, values = "party_state_average", color = "black") +
   facet_wrap(~year, nrow = 4, ncol= 3) +
@@
   )
-rm(map_data)
+
+House_by_year
+ggsave("../Documents/Graph_Images/House_by_year.png", width=7.29, height=5.41)
+rm(map_data, House_by_year)
 ```
@@
 ```{r}
-full_ntl_data |> 
+Union_wage_by_year <- full_ntl_data |> 
   ggplot(aes(x = year)) +
@@
   )
+
+Union_wage_by_year
+ggsave("../Documents/Graph_Images/Union_wage_by_year.png", width=7.29, height=5.41)
+rm(Union_wage_by_year)
 ```
@@
 ```{r}
-full_ntl_data |> 
+Union_wage_by_decade <- full_ntl_data |> 
   ggplot(aes(x = percent_collective_bargain)) +
@@
     y = "Wages (Adjusted to 2020 values)"
+  )
 
-  )
+Union_wage_by_decade
+ggsave("../Documents/Graph_Images/Union_wage_by_decade.png", width=7.29, height=5.41)
+rm(Union_wage_by_decade)
 ```
@@
 #| message: FALSE
-full_state_data |> 
+CB_by_decade1 <- full_state_data |> 
   filter(year %% 4==0) |>
@@
   labs(
-    title = "Percent Collective Bargaining vs Proportion of House Delegation That is Democratic",
+    title = "% Collective Bargaining vs Proportion of House Delegation That is Democratic",
     x = "Percent of Workers Covered by Collective Bargaining",
@@
   )
+
+CB_by_decade1
+ggsave("../Documents/Graph_Images/CB_by_decade1.png", width=7.29, height=5.41)
+rm(CB_by_decade1)
 ```
@@
 #| message: FALSE
-full_state_data |> 
+CB_by_decade2 <- full_state_data |> 
   filter(year %% 4==0) |>
@@
   labs(
-    title = "Percent Collective Bargaining vs Proportion of House Delegation That is Democratic",
+    title = "% Collective Bargaining vs Proportion of House Delegation That is Democratic",
     subtitle = "Regressed by Region",
@@
   )
+
+CB_by_decade2
+ggsave("../Documents/Graph_Images/CB_by_decade2.png", width=7.29, height=5.41)
+rm(CB_by_decade2)
 ```
@@
 t.test(full_ntl_data_2010s$union_wage_2020_adj, full_ntl_data_2010s$nonunion_wage_2020_adj)
+rm(full_ntl_data_1980s, full_ntl_data_2010s)
 ```
@@
   rename(P_Value = "Pr(>|t|)")
+
+rm(CB_party_temp, x, regression_decade_region)
 ```
@@
 ```{r}
-ggplot(CB_party_regression, aes(x = Decade, y = Estimate, fill = Region)) +
-  geom_bar(stat = "identity", position = "dodge", color = ifelse(CB_party_regression$P_Value > 0.05, "black", NA), size = 0.2) +
+Regression_plot <- ggplot(CB_party_regression, aes(x = Decade, y = Estimate, fill = Region)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  geom_text(aes(label = sprintf("P=%.3f", P_Value)),
+             position = position_dodge(width = 0.9),
+             vjust = -0.2,
+             size = 3) +
   theme_minimal() +
@@
        subtitle = "Model of Percent Collective Bargaining vs Proportion of House Delegation That is Democratic")
+
+Regression_plot
+ggsave("../Documents/Graph_Images/Regression_plot.png", width=7.29, height=5.41)
+rm(Regression_plot, CB_party_regression, decade_region)
 ```

### 2024-01-12 Documents/Presentation.qmd
moved files around

---
title: "WIDS 2024 Final Presentation"
author: "Michael Taffe, Melanie Klein, Ramiro Arevalo"
format: 
  revealjs:
    theme: simple
    monofont: "Courier New"
    monofontoptions: "size=10%"
editor: visual
---

## Project Abstract

For our project, our goal was to examine the relationships between collective bargaining agreements, U.S. house election results, and incomes of Americans. The motivation for this project is the rise in wealth inequality in America, the rise of populism in America, and the recent surge in union activity. In our project, we found that all three metrics are related. Using regression testing we show that for unions and political leanings, this relationship has surged and deepened in the last forty years.

## Literature Review

## About the Data pt. 1 {.smaller}

-   **Union data**
    -   Gathered from Tidy Tuesday, although the original data comes from the Union Membership and Coverage Database at Georgia state university
    -   For our project, the relevant tables include one that details union membership counts by state by year and another that tracks union and non union wages by year.
-   **House election data**
    -   Also gathered from Tidy Tuesday's repository, this data originally came from the MIT Election Data and Science Lab
    -   The variables we selected from the data set were year, state, district, and election winner

## About the Data pt. 2 {.smaller}

-   **Wage**
    -   Our wage data was obtained through the U.S. Department of Labor and Tidy Tuesday (originally from U.S. census data)
    -   This data tracked the mean and median income nationally in the United States
    -   Another table tracked minimum wage at the state level

## Data manipulation pt. 1

-   All of our data had undergone basic cleaning before we downloaded it so all of the columns were correctly formatted and there was almost 0 NA values
-   each data set was pared down so that only the variables relevant to our project was completed.
-   for our house election data, the table was manipulated so that only rows of candidates who won their election remained.
    -   other minor changes included renaming variables and recoding a few outlier observations in the `party` variable
-   For the union data, we filtered the datasets down to not include union sectors (ex. construction-unions) and added columns to adjust for inflation
-   For our wage data, we filtered out race, as this was outside of the scope of our study.

## Data Manipulation pt. 2

-   Our last data manipulation was combining all of our tables into two data sets: one that stores data at a national level, and another that stores data at a state level.
-   some of the variables were only provided at the state level so they were summarized and added to the national data set.

## EDA pt. 1: Overview

-   EDA was split into five sections: wage data, union data, house data, wage data's associations with union data, and union data's association with house data. We did not have the relevent wage data to compare wages and house elections

## EDA pt. 2: Wage Data

::: columns
::: {.column width="60%"}
-   Minimum wages, which had been stagnant since 2009, began to grow again in 2016, although the federal rate has remained stagnant
-   Incomes, when adjusted for inflation, have been decreasing since 1984. Surprisingly, our data shows income inequality has remained stagnant over the last forty years.
:::

::: {.column width="40%"}

:::
:::

## EDA pt. 3: Union Data

::: columns
::: {.column width="60%"}
-   Union and collective bargaining rates have collapsed over the last 40 years, although may just beginning to see a come back in the last two.
-   Southern states have much less union activity than northern and western states.
:::

::: {.column width="40%"}
column 2
:::
:::

## EDA pt. 4: House Data

::: columns
::: {.column width="60%"}
-   The country has generally grown more republican in the last 40 years, although data in the last 10 years show the pendulem might now be swinging the other way
    -   Northern-midwest states have had the biggest shift to the right. New England is the only region that has seen a shift left.
:::

::: {.column width="40%"}
column 2
:::
:::

## EDA pt. 5: Wage Data's Interactions With Union Data

::: columns
::: {.column width="60%"}
-   From the graph, it appears that a reduction in the amount of employees covered by a union leads to an increase in wages. However, for individual workers, being part of a union remains a way to increase wages.
-   We do not have data to compare wages and union rates by state, which would likely tell a different story.
:::

::: {.column width="40%"}
column 2
:::
:::

## EDA pt. 6: House Election Data's Interactions With Union Data

::: columns
::: {.column width="60%"}
-   Unions have become a very political topic in the last 40 years
-   This trend appears to be driven by the northeast and south
:::

::: {.column width="40%"}
column 2
:::
:::

## Models

## Conclusion

## Where we would take our project in the future

##

### 2024-01-12 Documents/Presentation.qmd
populated presentation

@@
 
-## Data manipulation pt. 1
+## Data manipulation pt. 1 {.smaller}
 
@@
 
-## Data Manipulation pt. 2
+## Data manipulation pt. 2 {.smaller}
 
@@
 
-## EDA pt. 1: Overview
+## EDA pt. 1: Overview {.smaller}
 
 -   EDA was split into five sections: wage data, union data, house data, wage data's associations with union data, and union data's association with house data.
-We did not have the relevent wage data to compare wages and house elections
+We did not have the relevant wage data to compare wages and house elections
 
-## EDA pt. 2: Wage Data
+## EDA pt. 2: Wage Data {.smaller}
 
@@
 ::: {.column width="60%"}
+-   Incomes, when adjusted for inflation, have been decreasing since 1984. Surprisingly, our data shows income inequality has remained stagnant over the last forty years.
 -   Minimum wages, which had been stagnant since 2009, began to grow again in 2016, although the federal rate has remained stagnant
--   Incomes, when adjusted for inflation, have been decreasing since 1984. Surprisingly, our data shows income inequality has remained stagnant over the last forty years.
 :::
@@
 ::: {.column width="40%"}
-
+![Plot 1](./Graph_Images/Income_by_year_adj.png)
+![Plot 2](./Graph_Images/Eff_wage_by_state.png)
 :::
@@
 
-## EDA pt. 3: Union Data
+## EDA pt. 3: Union Data {.smaller}
 
@@
 ::: {.column width="40%"}
-column 2
+![Plot 3](./Graph_Images/States_by_CB.png)
 :::
@@
 
-## EDA pt. 4: House Data
+## EDA pt. 4: House Data {.smaller}
 
@@
 ::: {.column width="60%"}
--   The country has generally grown more republican in the last 40 years, although data in the last 10 years show the pendulem might now be swinging the other way
+-   The country has generally grown more republican in the last 40 years, although data in the last 10 years show the pendulum might now be swinging the other way
     -   Northern-midwest states have had the biggest shift to the right. New England is the only region that has seen a shift left.
@@
 ::: {.column width="40%"}
-column 2
+![Plot 4](./Graph_Images/House_by_year.png)
 :::
@@
 
-## EDA pt. 5: Wage Data's Interactions With Union Data
+## EDA pt. 5: Wage Data's Interactions With Union Data {.smaller}
 
@@
 ::: {.column width="40%"}
-column 2
+![Plot 5](./Graph_Images/Union_wage_by_decade.png)
 :::
@@
 
-## EDA pt. 6: House Election Data's Interactions With Union Data
+## EDA pt. 6: House Election Data's Interactions With Union Data {.smaller}
 
@@
 -   Unions have become a very political topic in the last 40 years
--   This trend appears to be driven by the northeast and south
+-   This trend appears to be driven by the north central and south regions.
 :::
@@
 ::: {.column width="40%"}
-column 2
+![Plot 6](./Graph_Images/CB_by_decade1.png)
+![Plot 7](./Graph_Images/CB_by_decade2.png)
+
 :::
@@
 
-## Models
+## Models {.smaller}
+::: columns
+::: {.column width="60%"}
+-   Regression
+- T-test between union and non union wages
+  - 1980s: t = 22.846, df = 9.4096, p-value = 1.462e-09
+    - mean of x = 26.306;  mean of y = 20.375
+    - 95% CI: (5.347, 6.514)
+  - 2010s: t = 22.846, df = 9.4096, p-value = 1.462e-09
+    - mean of x = 3.786;  mean of y = 25.664
+    - 95% CI: (3.786, 5.952)
+:::
 
-## Conclusion
+::: {.column width="40%"}
+![Plot 8](./Graph_Images/Regression_plot.png)
 
-## Where we would take our project in the future
+:::
+:::
 
-##
+## Conclusion {.smaller}
+
+## Where we would take our project in the future {.smaller}
+
+# Thank You
+
+## Bibliography {.smaller}

### 2024-01-12 Documents/Presentation.qmd

@@
 For our project, our goal was to examine the relationships between collective bargaining agreements, U.S.
-house election results, and incomes of Americans.
-The motivation for this project is the rise in wealth inequality in America, the rise of populism in America, and the recent surge in union activity.
+house election results, and incomes/wages of Americans.
+The motivation for this project is the rise in wealth inequality and populism in America and the recent surge in union activity.
 In our project, we found that all three metrics are related.
@@
 -   **Union data**
-    -   Gathered from Tidy Tuesday, although the original data comes from the Union Membership and Coverage Database at Georgia state university
+    -   Gathered from Tidy Tuesday, although the original data comes from the Union Membership and Coverage Database at Georgia state university.
     -   For our project, the relevant tables include one that details union membership counts by state by year and another that tracks union and non union wages by year.
 -   **House election data**
-    -   Also gathered from Tidy Tuesday's repository, this data originally came from the MIT Election Data and Science Lab
+    -   Also gathered from Tidy Tuesday's repository, this data originally came from the MIT Election Data and Science Lab.
     -   The variables we selected from the data set were year, state, district, and election winner
@@
 -   **Wage**
-    -   Our wage data was obtained through the U.S. Department of Labor and Tidy Tuesday (originally from U.S. census data)
-    -   This data tracked the mean and median income nationally in the United States
-    -   Another table tracked minimum wage at the state level
+    -   Our wage data was obtained through the U.S. Department of Labor and Tidy Tuesday (originally from U.S. census data).
+    -   This data tracked the mean and median income nationally in the United States.
+    -   Another table tracked minimum wage at the state level.
 
@@
 
--   All of our data had undergone basic cleaning before we downloaded it so all of the columns were correctly formatted and there was almost 0 NA values
+-   All of our data had undergone basic cleaning before we downloaded it so all of the columns were correctly formatted and there was almost 0 NA values.
 -   each data set was pared down so that only the variables relevant to our project was completed.
 -   for our house election data, the table was manipulated so that only rows of candidates who won their election remained.
-    -   other minor changes included renaming variables and recoding a few outlier observations in the `party` variable
--   For the union data, we filtered the datasets down to not include union sectors (ex. construction-unions) and added columns to adjust for inflation
+    -   other minor changes included renaming variables and recoding a few outlier observations in the `party` variable.
+-   For the union data, we filtered the datasets down to not include union sectors (ex. construction-unions) and added columns to adjust for inflation.
 -   For our wage data, we filtered out race, as this was outside of the scope of our study.
 
-## Data manipulation pt. 2 {.smaller}
+## Data manipulation pt. 2
 
@@
 
-## EDA pt. 1: Overview {.smaller}
+## EDA pt. 1: Overview
 
 -   EDA was split into five sections: wage data, union data, house data, wage data's associations with union data, and union data's association with house data.
-We did not have the relevant wage data to compare wages and house elections
+We did not have the relevant wage data to compare wages and house elections.
 
@@
 -   Incomes, when adjusted for inflation, have been decreasing since 1984. Surprisingly, our data shows income inequality has remained stagnant over the last forty years.
--   Minimum wages, which had been stagnant since 2009, began to grow again in 2016, although the federal rate has remained stagnant
+-   Minimum wages, which had been stagnant since 2009, began to grow again in 2016, although the federal rate has remained stagnant.
 :::
@@
 ::: {.column width="60%"}
--   The country has generally grown more republican in the last 40 years, although data in the last 10 years show the pendulum might now be swinging the other way
+-   The country has generally grown more republican in the last 40 years, although data in the last 10 years show the pendulum might now be swinging the other way.
     -   Northern-midwest states have had the biggest shift to the right. New England is the only region that has seen a shift left.
@@
 ::: {.column width="60%"}
--   Unions have become a very political topic in the last 40 years
+-   Unions have become a very political topic in the last 40 years.
 -   This trend appears to be driven by the north central and south regions.
@@
 ::: {.column width="40%"}
-![Plot 6](./Graph_Images/CB_by_decade1.png)
-![Plot 7](./Graph_Images/CB_by_decade2.png)
+![Plot 6](./Graph_Images/CB_by_decade1.png){width=80%}
+![Plot 7](./Graph_Images/CB_by_decade2.png){width=80%}
 
@@
 ::: {.column width="60%"}
--   Regression
+-   Our regression analysis shows that the south and north central regions in the united states are the main regions that have seen union membership rates and political leanings show increased correlation.
 - T-test between union and non union wages
@@
 
+From our data analysis it can be infered that collective bargaining agreements have become a very political concept.
+The charts in our exploratory data analysis and regression models show this.
+The relationship between unions and wages/income in the United States is much harder to draw inferences.
+While our t-tests and some charts show that unions can help workers earn more, other graphs show that unions have a negative effect on incomes.
+
 ## Where we would take our project in the future {.smaller}
+- Wage data at the state level
+- Data past 2019
+- A greater amount and complexity of models
 

### 2024-01-12 Documents/Paper.qmd
work on paper

--- 2024-01-11 Documents/Paper.qmd
@@
 ---
-title: "Your Substantive Title Here"
+title: "Comparing Interactions Between Collective Bargaining Rates, Wages, and House of Representative Elections"
 author: 
-  - Your Name1^[American University]
-  - Your Name2^[American University]
+  - Ramiro Arevalo^[American University]
+  - Melanie Klein^[American University]
+  - Michael Taffe^[American University]
+
 format: 
@@
       - top=3cm
-date: 2023-09-27
+date: 2024-01-12
 date-format: iso
@@
 abstract: "This is our informative abstract of fewer than 200 words. It describes what we investigate, how we investigate it, and what we find."
-bibliography: main.bib
+#bibliography: main.bib
 ---
@@
 #| message: false
-
-# This chunk loads any packages we need.
-# Because "message: false", any R messages from this chunk do not appear in our paper.
 
@@
 
-# This chunk might read our data.
-# It might clean the data, create new variables, etc.
-# Now our data are ready for our paper.
-
-# Because echo: false, this chunk itself is not shown.
+# read data, clean the data, create new variables, etc.
 
@@
 
-# Clean house election data
-house <- house |>
-  group_by(year, state, district) |> # group for next line
-  filter(candidatevotes == max(candidatevotes)) |> #only include candidates that won
-  ungroup() |> 
-  mutate(state = str_to_sentence(state)) |> # change from all caps to sentence case) 
-  select(!(state_fips:office)) |> #Remove unwanted columns
-  select(!(stage:candidate)) |> 
-  select(!(writein:fusion_ticket)) |> 
-  rename(state_abbreviation = state_po) |>  # rename state abbreviation column
-  mutate(party = case_when(
-    str_detect(party, "DEMOCRAT") ~ "DEMOCRAT",
-    str_detect(party, "REPUBLICAN") ~ "REUBLICAN",
-    str_detect(party, "INDEPENDENT") ~ "INDEPENDENT"
-  )) # Recode of party variable
-
-# Clean union wages data
-union_wages <- union_wages |> 
-  filter(facet == "all wage and salary workers") |> 
-  distinct(year, .keep_all = TRUE) |> 
-  select(-sample_size,
-         -at_cap) |> 
-  mutate(wage_percent_difference = union_wage_premium_raw, .keep = "unused") |> 
-  mutate(wage_percent_difference_adjusted = union_wage_premium_adjusted, .keep = "unused") |> 
-  mutate(union_type = facet, .keep = "unused")
-
-# Clean union states data
-union_states <- union_states |> 
-  select(-state_census_code,
-         -observations) |> # remove unwanted columns
-  mutate(employment_count = employment * 1000, .keep = "unused") |> # column measured in 1000s, this fixes it
-  mutate(union_employment_count = members * 1000, .keep = "unused") |> 
-  mutate(collective_bargain_employment_count = covered * 1000,
-         .keep = "unused") |> 
-  rename(
-    percent_union = p_members,
-    percent_collective_bargain = p_covered
-  ) #renamed columns
-
-# Clean income distribution data
-income_distribution <- income_distribution |> 
-  select(-number) |> # removing unwanted columns
-  select(!ends_with("moe")) |> 
-  select(year:income_mean) |> 
-  distinct(year, race, income_median, income_mean, .keep_all = TRUE) |> # only include one line per year. needed bc of multiple sectors that are not being used in data analysis
-  rename(median_income = income_median,
-         mean_income = income_mean) |> # renaming columns
-  filter(race == "All Races") |> 
-  select(-race)
-
-# Clean minimum wage data
-minimum_wage <- minimum_wage |> 
-  select(Year:CPI.Average) |> # selecting relevant columns
-  rename(year = Year,
-         state = State,
-         state_min_wage = State.Minimum.Wage,
-         state_min_wage_2020_adj = State.Minimum.Wage.2020.Dollars,
-         fed_min_wage = Federal.Minimum.Wage,
-         fed_min_wage_2020_adj = Federal.Minimum.Wage.2020.Dollars,
-         effective_min_wage = Effective.Minimum.Wage,
-         effective_min_wage_2020_adj = Effective.Minimum.Wage.2020.Dollars,
-         CPI_avg = CPI.Average
-  ) |> # renaming variables
-  filter(!state %in% c("U.S. Virgin Islands", "Puerto Rico", "Guam")) |>  # remove territory
-  mutate(state_abbreviation = 
-           ifelse(is.na(state.abb[match(state, state.name)]),
-                  "DC",
-                  state.abb[match(state, state.name)]),
-         .keep = "unused", .after = year) # replaces state with state abbreviation, NA values coded as DC as DC is the only NA value 
-```
-
-```{r readdata}
-#| eval: false
-#| echo: false
-
-# Because eval: false, this chunk is not run.
-
-# Joining data sets to create full state data
-full_state_data <- union_states |> 
-  filter(sector == "Total") |> 
-  select(-sector,
-         -state) |> 
-  left_join(house, by = c("year", "state_abbreviation")) |> 
-  mutate(decade = paste0((year - year %% 10), "s"),
-         party_binary = ifelse(party == "DEMOCRAT", 1, 0)) |> 
-  left_join(minimum_wage, by = c("year", "state_abbreviation")) |> 
-  group_by(year, state) |> 
-  mutate(party_state_average = mean(party_binary)) |>
-  ungroup() |> 
-  mutate(state = state.name[match(state_abbreviation, state.abb)], # fix state encoding
-         regions = state.region[match(state_abbreviation, state.abb)]) # fix region encoding
-
-state_column_list = c("year", "decade", "state", "state_abbreviation",
-                      "regions", "district", "party", "party_binary",
-                      "party_state_average",
-                      "employment_count", "union_employment_count",
-                      "collective_bargain_employment_count", "percent_union",
-                      "percent_collective_bargain", "state_min_wage",
-                      "state_min_wage_2020_adj", "fed_min_wage",
-                      "fed_min_wage_2020_adj", "effective_min_wage",
-                      "effective_min_wage_2020_adj", "CPI_avg")
-full_state_data <- full_state_data[, state_column_list] # reorder columns
-full_state_data$party[is.na(full_state_data$party)] <- "No election" # replace party NAs
-full_state_data$district[is.na(full_state_data$district)] <- "No election" # replace district NAs
-full_state_data$party_binary[is.na(full_state_data$party_binary)] <- "No election" # replace district NAs
-
-# Joining data sets to create full national data
-state_to_ntl_adjusted <- full_state_data |>
-  mutate(prop_2020_adj = fed_min_wage_2020_adj / fed_min_wage, .after = year) |>
-  distinct(year, .keep_all = TRUE) |>
-  select(year, prop_2020_adj)
-
-state_to_ntl_employment <- full_state_data |>
-  group_by(year) |>
-  summarise_at(vars(employment_count:collective_bargain_employment_count), sum)
-
-full_ntl_data <- income_distribution |> 
-  left_join(union_wages, by = "year") |>
-  left_join(state_to_ntl_adjusted, by = "year") |>
-  left_join(state_to_ntl_employment, by = "year") |>
-  mutate(decade = paste0((year - year %% 10), "s"), .after = year) |>
-  mutate(mean_median_income_difference = mean_income - median_income, .after = mean_income) |>
-   mutate(median_income_2020_adj = median_income * prop_2020_adj,
-         mean_income_2020_adj = mean_income * prop_2020_adj,
-         mean_median_income_difference_2020_adj = mean_median_income_difference * prop_2020_adj,
-         wage_2020_adj = wage * prop_2020_adj,
-         union_wage_2020_adj = union_wage * prop_2020_adj,
-         nonunion_wage_2020_adj = nonunion_wage * prop_2020_adj,
-         percent_union = union_employment_count / employment_count,
-         percent_collective_bargain = collective_bargain_employment_count / employment_count) |> 
-  select(-prop_2020_adj)
-
-ntl_column_list <- c("year", "decade", "mean_income",
-                     "mean_income_2020_adj", "median_income",
-                     "median_income_2020_adj", "mean_median_income_difference",
-                     "mean_median_income_difference_2020_adj", "wage",
-                     "wage_2020_adj", "union_wage", "union_wage_2020_adj",
-                     "nonunion_wage", "nonunion_wage_2020_adj",
-                     "wage_percent_difference", 
-                     "wage_percent_difference_adjusted",
-                     "employment_count", "union_employment_count", 
-                     "collective_bargain_employment_count",
-                     "percent_union", "percent_collective_bargain",
-                     "union_type")
-full_ntl_data <- full_ntl_data[, ntl_column_list] # reorder columns
-
-rm(house, income_distribution, minimum_wage, state_to_ntl_adjusted,
-   state_to_ntl_employment, union_states, union_wages, state_column_list,
-   ntl_column_list)
 ```
@@
 # Introduction
+
+Unions play a crucial role in negotiating wages and benefits for workers through collective bargaining agreements.
+The political landscape, particularly the House of Representatives, holds significant legislative authority in influencing workers' rights and labor policies.
+We examine the relationship between union employment, the composition of the House delegation, and economic variables such as income, minimum wages, and union wages.
+
+In our research, we seek to understand how union membership and economic factors influence the level of representation by the Democratic Party in the House of Representatives.
+We hypothesize that there will be an association between the Democratic Party and the employment of union and collective bargaining employees.
 
@@
 
-# [Our Substance and Context Section Title Here]
+# \[Our Substance and Context Section Title Here\]
 
@@
 To give the reader a clear sense of why we are writing this paper, we describe the relevant scholarly, technical, or popular literature.
-We give this section a meaningful _substantive_ title; it is not entitled "Literature Review", for example.
+We give this section a meaningful *substantive* title; it is not entitled "Literature Review", for example.
 We cite at least three published, peer-reviewed scholarly works.
-For example, we could cite @mooree20 or @moorav12^[There should always be a space before the "(" in a citation date.], which we discussed in class.^[To cite a paper within parentheses, use, e.g., [@moore12].] We only cite others' work in our paper when it enhances the reader's understanding of what we, the authors of this paper, are doing.
-We connect everything we cite to _our_ investigation; this is our original research, not a book report or an annotated bibliography.
+For example, we could cite @mooree20 or @moorav12[^1], which we discussed in class.[^2] We only cite others' work in our paper when it enhances the reader's understanding of what we, the authors of this paper, are doing.
+We connect everything we cite to *our* investigation; this is our original research, not a book report or an annotated bibliography.
+
+[^1]: There should always be a space before the "(" in a citation date.
+
+[^2]: To cite a paper within parentheses, use, e.g., [@moore12].
 
@@
 
-This section describes the data we analyze. We describe the source of the data,
-and its primary features. We cite our data. We describe the methods we use to
-answer our question and to test our hypotheses.
+This section describes the data we analyze.
+We describe the source of the data, and its primary features.
+We cite our data.
+We describe the methods we use to answer our question and to test our hypotheses.
 
@@
 
-# [Our Results Section Title Here]
+# \[Our Results Section Title Here\]
 
@@
 
-$$\text{Distance}_i = \beta_0 + \beta_1 (\text{Speed}_i) + \epsilon_i.$$
-Perhaps we start by plotting the data, as in @fig-cars.
+$$\text{Distance}_i = \beta_0 + \beta_1 (\text{Speed}_i) + \epsilon_i.$$ Perhaps we start by plotting the data, as in @fig-cars.
 
@@
 
-cars <- cars |> mutate(dist_sqrt = (sqrt(dist)))
-t_out_2 <- t.test(dist_sqrt ~ hs, data = cars)
-
-tab <- rbind(
-  model_parameters(t_out),
-  model_parameters(t_out_2)
-)
-
-tab |> 
-  print_md(
-    footer = "")
 ```

### 2024-01-12 Documents/Paper.qmd
updates in code presentation and paper

@@
 editor: source
-abstract: "This is our informative abstract of fewer than 200 words. It describes what we investigate, how we investigate it, and what we find."
+abstract: "For our project, our goal is to examine the relationships between collective bargaining agreements, U.S.
+house election results, and incomes/wages of Americans.
+The motivation for this project is the rise in wealth inequality and populism in America and the recent surge in union activity.
+In our project, we find that all three metrics are related.
+Using regression testing we show that for unions and political leanings, this relationship has surged and deepened in the last four decades from 1984 to 2019"
 #bibliography: main.bib
@@
 The political landscape, particularly the House of Representatives, holds significant legislative authority in influencing workers' rights and labor policies.
-We examine the relationship between union employment, the composition of the House delegation, and economic variables such as income, minimum wages, and union wages.
+We examine the relationship between unions and collective bargaining agreements, U.S.
+House of Representatives election results, and economic variables including the incomes and wages of Americans.
 
 In our research, we seek to understand how union membership and economic factors influence the level of representation by the Democratic Party in the House of Representatives.
-We hypothesize that there will be an association between the Democratic Party and the employment of union and collective bargaining employees.
+We hypothesize that there will be an association between the proportion of the House delegation that is Democratic and the employment of union and collective bargaining workers.
 

### 2024-01-12 Documents/Presentation.qmd
updates in code presentation and paper

@@
 
-For our project, our goal was to examine the relationships between collective bargaining agreements, U.S.
+For our project, our goal is to examine the relationships between collective bargaining agreements, U.S.
 house election results, and incomes/wages of Americans.
 The motivation for this project is the rise in wealth inequality and populism in America and the recent surge in union activity.
-In our project, we found that all three metrics are related.
+In our project, we find that all three metrics are related.
 Using regression testing we show that for unions and political leanings, this relationship has surged and deepened in the last forty years.
@@
 -   **Union data**
-    -   Gathered from Tidy Tuesday, although the original data comes from the Union Membership and Coverage Database at Georgia state university.
+    -   Gathered from Tidy Tuesday, although the original data comes from the Union Membership and Coverage Database at Georgia State University.
     -   For our project, the relevant tables include one that details union membership counts by state by year and another that tracks union and non union wages by year.
@@
 -   All of our data had undergone basic cleaning before we downloaded it so all of the columns were correctly formatted and there was almost 0 NA values.
--   each data set was pared down so that only the variables relevant to our project was completed.
--   for our house election data, the table was manipulated so that only rows of candidates who won their election remained.
-    -   other minor changes included renaming variables and recoding a few outlier observations in the `party` variable.
+-   Each data set was pared down so that only the variables relevant to our project was completed.
+-   For our house election data, the table was manipulated so that only rows of candidates who won their election remained.
+    -   Other minor changes included renaming variables and recoding a few outlier observations in the `party` variable.
 -   For the union data, we filtered the datasets down to not include union sectors (ex. construction-unions) and added columns to adjust for inflation.
@@
 -   Our last data manipulation was combining all of our tables into two data sets: one that stores data at a national level, and another that stores data at a state level.
--   some of the variables were only provided at the state level so they were summarized and added to the national data set.
+-   Some of the variables were only provided at the state level so they were summarized and added to the national data set.
 
@@
 ::: {.column width="40%"}
-![Plot 1](./Graph_Images/Income_by_year_adj.png)
-![Plot 2](./Graph_Images/Eff_wage_by_state.png)
+![Plot 1](./Graph_Images/Income_by_year_adj.png) ![Plot 2](./Graph_Images/Eff_wage_by_state.png)
 :::
@@
 ::: {.column width="40%"}
-![Plot 6](./Graph_Images/CB_by_decade1.png){width=80%}
-![Plot 7](./Graph_Images/CB_by_decade2.png){width=80%}
-
+![Plot 6](./Graph_Images/CB_by_decade1.png){width="80%"} ![Plot 7](./Graph_Images/CB_by_decade2.png){width="80%"}
 :::
@@
 ## Models {.smaller}
+
 ::: columns
 ::: {.column width="60%"}
--   Our regression analysis shows that the south and north central regions in the united states are the main regions that have seen union membership rates and political leanings show increased correlation.
+-   Our regression analysis shows that the south and north central regions in the United States are the main regions that have seen union membership rates and political leanings show increased correlation.
 -   T-test between union and non union wages
@@
 ![Plot 8](./Graph_Images/Regression_plot.png)
-
 :::
@@
 
-From our data analysis it can be infered that collective bargaining agreements have become a very political concept.
+From our data analysis it can be inferred that collective bargaining agreements have become a very political concept.
 The charts in our exploratory data analysis and regression models show this.
@@
 ## Where we would take our project in the future {.smaller}
-- Wage data at the state level
+
+-   Union wage data at the state level
 -   Data past 2019

### 2024-01-12 Code/Full R Code.qmd

@@
 # 1. Comparing interactions between collective bargaining rates, wages, and House of Representative elections
-
-**Insert abstract of project, where we got our data, what our data looks like, and other information here**
 
@@
 ```
+
 In this table, we can see that there are a few NA values in the tables, and some of the variables have a noticable skew. After some basic data cleaning is done we will see if these values remain
@@
 ```
+
 Only one column has NA values in this data frame, and in that column only 15 out of 2862 observations are missing values. Again, we will see if these values remain after our data cleaning
@@
 ```
+
 There are no more NA values
@@
 ```
+
 After combining the dataset, we can see that not all of the variables are tracked every year, causing there to be NA values. We will keep this in mind as we continue our data analysis.
@@
 ```
+
 There are no NA values in our national data set.
@@
 Next, we can see that there is a significant shift post 2014 of raising the minimum wage at the state level.
-Before that time almost no state had a adjusted minimum wage over $10.
+Before that time almost no state had a adjusted minimum wage over \$10.
 Now, there are numerous states that surpass this value.
@@
 ```
+
 These two graphs show very interesting results.
@@
 
-## Modeling
+## 2.4 Modeling
 
-t test between union and non union wages
+### 2.4.1 T-test between union and non union wages
+
 ```{r}
@@
 
-Linear regression of percent collective bargaining and party state average
+### 2.4.2 Linear regression of percent collective bargaining and party state average
+
 ```{r}
@@
 
-Graph of regression estimate by decade and year
+#### 2.4.2.1 Graph of regression estimate by decade and year
+
 ```{r}
@@
 ```
-
-## Conclusion

### 2024-01-12 Documents/Presentation.qmd
refined presentation

@@
 title: "WIDS 2024 Final Presentation"
+subtitle: "Comparing Collective Bargaining, Income, and House Election data"
 author: "Michael Taffe, Melanie Klein, Ramiro Arevalo"
@@
 ## Literature Review
+
+## Project Question
+
+The research questions we used to guide our analysis:
+
+1.  Is there any relationship between collective bargaining rates among workers to house delegation makeup?
+2.  Is there any relationship between collective bargaining rates among workers and wages
+3.  Is there any relationship between collective bargaining rates among workers and wages?
+
+## Hypotheses
+
+Guided by our questions we have two hypotheses:
+
+1.  There is a significant correlation between collective bargaining rates and higher amounts of Democrat representation.
+
+2.  Unions will lead to higher wages.
 
@@
 The charts in our exploratory data analysis and regression models show this.
+This partially supports our hypothesis.
+While the correlation is a newer phenomenon, the relationship appears to be strong today.
 The relationship between unions and wages/income in the United States is much harder to draw inferences.
 While our t-tests and some charts show that unions can help workers earn more, other graphs show that unions have a negative effect on incomes.
+Our second hypothesis thus remains without strong support.
 

### 2024-01-12 Code/Full R Code.qmd
worked on paper

@@
 This will be examined in closer detail in the exploratory data analysis.
-only two columns have NA values and each of those columns NA values represent less than 10% of the data.
+Only two columns have NA values and each of those columns NA values represent less than 10% of the data.
 There is not an immediate cause to take action on these NA values, although they will be examined in greater detail in the EDA.
@@
 
-#### 2.4.2.1 Graph of regression estimate by decade and year
+#### 2.4.2.1 Graph of regression estimate by decade and region
 

### 2024-01-12 Documents/Paper.qmd

@@
 ---
-title: "Comparing Interactions Between Collective Bargaining Rates, Wages, and House of Representative Elections"
+title: "Comparing Interactions Between Collective Bargaining, Wages and Income, and House of Representative Elections"
 author: 
@@
 
+library(parameters)
 library(tidyverse)
@@
 
-# read data, clean the data, create new variables, etc.
+# Read data
 
-# Union Data: obtained from Tidy Tuesday
+## Union Data: obtained from Tidy Tuesday
 union_wages <- read_rds("../Data/union_wages.rds")
@@
 
-# Election Data: Obtained from Tidy Tuesday
+## Election Data: Obtained from Tidy Tuesday
 house <- read_rds("../Data/house.rds")
 
-# Wage Data: Obtained from Tidy Tuesday
+## Wage Data: Obtained from Tidy Tuesday
 income_distribution <- read_rds("../Data/income_distribution.rds")
 
-# Minimum Wage Data: Obtained from U.S. Department of Labor
+## Minimum Wage Data: Obtained from U.S. Department of Labor
 minimum_wage <- read_csv("../Data/minimum_wage_data.csv",show_col_types = FALSE)
+
+# Clean data
+
+## house data
+house <- house |>
+  group_by(year, state, district) |> # group for next line
+  filter(candidatevotes == max(candidatevotes)) |> #only include candidates that won
+  ungroup() |> 
+  mutate(state = str_to_sentence(state)) |> # change from all caps to sentence case) 
+  select(!(state_fips:office)) |> #Remove unwanted columns
+  select(!(stage:candidate)) |> 
+  select(!(writein:fusion_ticket)) |> 
+  rename(state_abbreviation = state_po) |>  # rename state abbreviation column
+  mutate(party = case_when( # Recode of party variable
+    str_detect(party, "DEMOCRAT") ~ "DEMOCRAT",
+    str_detect(party, "REPUBLICAN") ~ "REUBLICAN",
+    str_detect(party, "INDEPENDENT") ~ "INDEPENDENT"
+  )) |> 
+  filter(state_abbreviation != "DC") # Removes the single line of DC data
+
+## union wages data
+union_wages <- union_wages |> 
+  filter(facet == "all wage and salary workers") |> 
+  distinct(year, .keep_all = TRUE) |> 
+  select(-sample_size,
+         -at_cap) |> 
+  mutate(wage_percent_difference = union_wage_premium_raw, .keep = "unused") |> 
+  mutate(wage_percent_difference_adjusted = union_wage_premium_adjusted, .keep = "unused") |> 
+  mutate(union_type = facet, .keep = "unused")
+
+## union states data
+union_states <- union_states |> 
+  select(-state_census_code,
+         -observations) |> # remove unwanted columns
+  mutate(employment_count = employment * 1000, .keep = "unused") |> # column measured in 1000s, this fixes it
+  mutate(union_employment_count = members * 1000, .keep = "unused") |> 
+  mutate(collective_bargain_employment_count = covered * 1000,
+         .keep = "unused") |> 
+  rename(
+    percent_union = p_members,
+    percent_collective_bargain = p_covered
+  ) #renamed columns
+
+## income distribution data
+income_distribution <- income_distribution |> 
+  select(-number) |> # removing unwanted columns
+  select(!ends_with("moe")) |> 
+  select(year:income_mean) |> 
+  distinct(year, race, income_median, income_mean, .keep_all = TRUE) |> # only include one line per year. needed bc of multiple sectors that are not being used in data analysis
+  rename(median_income = income_median,
+         mean_income = income_mean) |> # renaming columns
+  filter(race == "All Races") |> 
+  select(-race)
+
+## minimum wage data
+minimum_wage <- minimum_wage |> 
+  select(Year:CPI.Average) |> # selecting relevant columns
+  rename(year = Year,
+         state = State,
+         state_min_wage = State.Minimum.Wage,
+         state_min_wage_2020_adj = State.Minimum.Wage.2020.Dollars,
+         fed_min_wage = Federal.Minimum.Wage,
+         fed_min_wage_2020_adj = Federal.Minimum.Wage.2020.Dollars,
+         effective_min_wage = Effective.Minimum.Wage,
+         effective_min_wage_2020_adj = Effective.Minimum.Wage.2020.Dollars,
+         CPI_avg = CPI.Average
+  ) |> # renaming variables
+  filter(!state %in% c("U.S. Virgin Islands", "Puerto Rico", "Guam")) |>  # remove territory
+  mutate(state_abbreviation = 
+           ifelse(is.na(state.abb[match(state, state.name)]),
+                  "DC",
+                  state.abb[match(state, state.name)]),
+         .keep = "unused", .after = year) # replaces state with state abbreviation, NA values coded as DC as DC is the only NA value
+
+# Merging data frames
+
+## state data frame
+full_state_data <- union_states |> 
+  filter(sector == "Total") |> 
+  select(-sector,
+         -state) |> 
+  left_join(house, by = c("year", "state_abbreviation")) |> 
+  mutate(decade = paste0((year - year %% 10), "s"),
+         party_binary = ifelse(party == "DEMOCRAT", 1, 0)) |> 
+  left_join(minimum_wage, by = c("year", "state_abbreviation")) |> 
+  group_by(year, state) |> 
+  mutate(party_state_average = mean(party_binary)) |>
+  ungroup() |> 
+  mutate(state = state.name[match(state_abbreviation, state.abb)], # fix state encoding
+         regions = state.region[match(state_abbreviation, state.abb)]) |>  # fix region encoding
+  filter(year >= 1984 & year <= 2019) |> 
+  filter(state_abbreviation != "DC") # Removes the single line of DC data
+
+state_column_list = c("year", "decade", "state", "state_abbreviation",
+                      "regions", "district", "party", "party_binary",
+                      "party_state_average",
+                      "employment_count", "union_employment_count",
+                      "collective_bargain_employment_count", "percent_union",
+                      "percent_collective_bargain", "state_min_wage",
+                      "state_min_wage_2020_adj", "fed_min_wage",
+                      "fed_min_wage_2020_adj", "effective_min_wage",
+                      "effective_min_wage_2020_adj", "CPI_avg")
+full_state_data <- full_state_data[, state_column_list] # reorder columns
+full_state_data$party[is.na(full_state_data$party)] <- "No election" # replace party NAs
+full_state_data$district[is.na(full_state_data$district)] <- "No election" # replace district NAs
+full_state_data$party_binary[is.na(full_state_data$party_binary)] <- "No election" # replace district NAs
+
+## national data frame
+state_to_ntl_adjusted <- full_state_data |>
+  mutate(prop_2020_adj = fed_min_wage_2020_adj / fed_min_wage, .after = year) |>
+  distinct(year, .keep_all = TRUE) |>
+  select(year, prop_2020_adj)
+
+state_to_ntl_employment <- full_state_data |>
+  group_by(year) |>
+  summarise_at(vars(employment_count:collective_bargain_employment_count), sum)
+
+full_ntl_data <- income_distribution |> 
+  left_join(union_wages, by = "year") |>
+  left_join(state_to_ntl_adjusted, by = "year") |>
+  left_join(state_to_ntl_employment, by = "year") |>
+  mutate(decade = paste0((year - year %% 10), "s"), .after = year) |>
+  mutate(mean_median_income_difference = mean_income - median_income, .after = mean_income) |>
+   mutate(median_income_2020_adj = median_income * prop_2020_adj,
+         mean_income_2020_adj = mean_income * prop_2020_adj,
+         mean_median_income_difference_2020_adj = mean_median_income_difference * prop_2020_adj,
+         wage_2020_adj = wage * prop_2020_adj,
+         union_wage_2020_adj = union_wage * prop_2020_adj,
+         nonunion_wage_2020_adj = nonunion_wage * prop_2020_adj,
+         percent_union = union_employment_count / employment_count,
+         percent_collective_bargain = collective_bargain_employment_count / employment_count) |> 
+  select(-prop_2020_adj) |> 
+  filter(year >= 1984 & year <= 2019)
+
+ntl_column_list <- c("year", "decade", "mean_income",
+                     "mean_income_2020_adj", "median_income",
+                     "median_income_2020_adj", "mean_median_income_difference",
+                     "mean_median_income_difference_2020_adj", "wage",
+                     "wage_2020_adj", "union_wage", "union_wage_2020_adj",
+                     "nonunion_wage", "nonunion_wage_2020_adj",
+                     "wage_percent_difference", 
+                     "wage_percent_difference_adjusted",
+                     "employment_count", "union_employment_count", 
+                     "collective_bargain_employment_count",
+                     "percent_union", "percent_collective_bargain",
+                     "union_type")
+full_ntl_data <- full_ntl_data[, ntl_column_list] # reorder columns
+
+rm(house, income_distribution, minimum_wage, state_to_ntl_adjusted,
+   state_to_ntl_employment, union_states, union_wages, state_column_list,
+   ntl_column_list)
 
@@
 House of Representatives election results, and economic variables including the incomes and wages of Americans.
+These project topics are motivated by the recent surge in union activity, increased political polarization, and a rise in wealth inequality.
 
-In our research, we seek to understand how union membership and economic factors influence the level of representation by the Democratic Party in the House of Representatives.
-We hypothesize that there will be an association between the proportion of the House delegation that is Democratic and the employment of union and collective bargaining workers.
+In this paper, we use three research questions to guide our analysis:
 
-In this section, we introduce the reader to the phenomenon we investigate.
-We describe the way in which our analysis contributes to an important intellectual debate, or how it answers a pressing political or social question.
-We introduce our research question, hypotheses, data, and results.
-We signpost for the reader what's coming in the rest of the paper.
+1.    Is there any relationship between collective bargaining rates among workers to House delegation makeup?
+2.    Is there any relationship between collective bargaining rates among workers and wages?
+3.    Is there any relationship between collective bargaining rates among workers and wages?
 
-We remember that our paper is not a mystery novel. We note our core results early and often.
+Guided by our research questions, we have two hypothesis:
 
-Throughout our paper, we use active, first-person language and avoid the passive voice.
-For example, we write "we examine the relationship between $X$ and $Y$"; we do not write "the relationship between $X$ and $Y$ was examined." Where we do the analysis, we speak about it transparently.
-We use the present tense; for example, "In this paper, we argue \ldots" and "Paper XYZ demonstrates the relationship between \ldots".
+1.    There is a significant correlation between collective bargaining rates and higher amounts of Democrat representation in the House of Representatives. 
+2.    Unions will lead to higher wages.
+
+We use five data sets that are joined into two data frames to explore our research questions.
+We create a data frame that stores information at the national level about income and union wages, and a data frame that stores information at the state level about minimum wages, union employment, and House election results.
+
+When conducting a regression analysis, it demonstrates that the correlation between collective bargaining and political leanings has gotten stronger in the last four decades from 1984 to 2019.
+Additionally, when conducting t-tests among union and non union wages, there is a statistically significant difference that demonstrates an increasing wage differential over the same time period.
 
@@
 
-# \[Our Results Section Title Here\]
+# Exploring Correlations: Linear Regression Analyses and T-Tests on Collective Bargaining and Wage Disparities
 
-Here, we explain and interpret our results.
-We try to learn as much as we can about our question as possible, given the data and analysis.
-We present our results clearly.
-We interpret them for the reader with precision and circumspection.
-We avoid making claims that are not substantiated by our data.
-We are careful about causality.
-When we describe associations, we avoid language like "effects" and "increases"; we only describe "effects" or "impacts" when we have a causally well-identified research design.
+Based on our exploratory data analysis, we are conducting linear regression analyses and t-tests.
+The linear regressions use data from `full_state_data` and explore how collective bargaining agreements are correlated with Democratic representatives being elected to the House in all four regions.
+The results are used to examine if the north central region and southern region are driving this stronger correlation over time.
+Additionally, the t-test uses data from `full_ntl_data` to understand if the observed variations in wages between unionized and non-unionized workers are statistically meaningful.
+By looking at different decades, we aim to analyze potential shifts in wage consistencies or differences over time.
+This provides insight into the dynamics of the labor market.
 
-Note that this section may be integrated into @sec-data, if joining the two improves the overall presentation.
+## Predicting Democratic House Delegations with Percentage of Workers in Collective Bargaining Argreements
 
-## Predicting Distance with Speed
-
-Our results for the `cars` data include estimating the linear model
-
-$$\text{Distance}_i = \beta_0 + \beta_1 (\text{Speed}_i) + \epsilon_i.$$ Perhaps we start by plotting the data, as in @fig-cars.
-
-```{r fig-cars}
-#| echo: false
-#| fig-cap: Distance on Speed
-
-ggplot(cars, aes(speed, dist)) +
-  geom_point()
-```
-
-The data may be roughly linear, though there may be some non-linearity we should incorporate.
+@tbl-linear performs a linear regression between the percent of workers in a collective bargaining agreement and the proportion of the House delegation that is Democratic for the 1980s and 2010s in each region, as well as for all of the regions combined
 
@@
 
-# Estimate a linear model:
-lm_out <- lm(dist ~ speed, data = cars)
-# Extract the coefficient on speed:
-cars_speed_coef <- coef(lm_out)["speed"]
+decade_region <- list(
+  c("1980s", "Northeast"),
+  c("1980s", "West"),
+  c("1980s", "North Central"),
+  c("1980s", "South"),
+  c("2010s", "Northeast"),
+  c("2010s", "West"),
+  c("2010s", "North Central"),
+  c("2010s", "South")
+)
+
+regression_decade_region <- function(decade1, region1) {
+  full_state_data_filtered <- subset(full_state_data, regions == region1 & decade == decade1)
+  model <- lm(full_state_data_filtered$party_state_average~full_state_data_filtered$percent_collective_bargain) 
+  summary(model)$coefficients |> 
+  as_tibble() |> 
+  mutate(Reggression_Run = c("Intercept", "percent_collective_bargain"),
+         Decade = c(decade1, decade1),
+         Region = c(region1, region1),
+         .before = Estimate)
+}
+
+full_state_data_temp <- full_state_data |>
+  filter(year >= 1984 & year < 1990)
+CB_party_regression <- summary(lm(full_state_data_temp$party_state_average~full_state_data_temp$percent_collective_bargain))$coefficients |> 
+  as_tibble() |> 
+  mutate(Reggression_Run = c("Intercept", "percent_collective_bargain"),
+         Decade = c("1980s", "1980s"),
+         Region = c("all", "all"),
+         .before = Estimate)
+
+for (x in decade_region) {
+  CB_party_regression <- full_join(CB_party_regression, regression_decade_region(x[1], x[2]),
+                                   by = join_by(Reggression_Run, Decade, Region, Estimate, `Std. Error`, `t value`, `Pr(>|t|)`))
+}
+
+full_state_data_temp <- full_state_data |>
+  filter(year >= 2010 & year < 2020)
+CB_party_temp <- summary(lm(full_state_data_temp$party_state_average~full_state_data_temp$percent_collective_bargain))$coefficients |> 
+  as_tibble() |> 
+  mutate(Reggression_Run = c("Intercept", "percent_collective_bargain"),
+         Decade = c("2010s", "2010s"),
+         Region = c("all", "all"),
+         .before = Estimate)
+
+CB_party_regression <- full_join(CB_party_regression, CB_party_temp,
+                                   by = join_by(Reggression_Run, Decade, Region, Estimate, `Std. Error`, `t value`, `Pr(>|t|)`)) |>
+  filter(Reggression_Run == "percent_collective_bargain") |>
+  rename(P_Value = "Pr(>|t|)")
+
 ```
 
-Below we show the model estimates. The first table uses `xtable()`, the second uses `stargazer()` [@hlavac18].
-
-```{r linearxtable}
+```{r tbl-linear}
 #| echo: false
@@
 #| results: 'asis'
+#| tbl-cap: Linear Regression
 
-# We can print regression tables with xtable or stargazer:
-regr_table <- xtable::xtable(lm_out,
-                             digits = 2,
-                             caption = "Our Informative Caption")
+regr_table <- xtable::xtable(CB_party_regression,
+                             digits = 2)
 
@@
 
-```{r linearstargazer}
+For all linear regressions, $p < 0.05$, except for in the 1980s when examining `all` regions, $p = 0.659$.
+
+```{r fig-estimates}
 #| echo: false
-#| message: false
-#| results: 'asis'
+#| fig-cap: Linear Regression Estimates
 
-# We can print regression tables with xtable or stargazer:
-stargazer::stargazer(lm_out, 
-                     title = "Our Informative Title",
-                     dep.var.caption = "Outcome",
-                     digits = 2,
-                     header = FALSE)
+ggplot(CB_party_regression, aes(x = Decade, y = Estimate, fill = Region)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  geom_text(aes(label = sprintf("P=%.3f", P_Value)),
+             position = position_dodge(width = 0.9),
+             vjust = -0.2,
+             size = 2) +
+  theme_minimal() +
+  labs(title = "Linear Regression Estimate by Decade",
+       subtitle = "Model of Percent Collective Bargaining vs Democratic House Proportion")
 ```
 
-Using the `cars` data, we find that each unit of speed is associated with `r round(cars_speed_coef, 1)` more units of distance.
-We draw out what this really means, and what it implies.
-For example, if a typical difference among our observations is 7 units of speed, then our model estimates that a typical difference in distance among our observations is $7 \times `r round(cars_speed_coef, 1)` = `r 7 * round(cars_speed_coef, 1)`$ units of distance.
-We describe the substantive relevance of this number.
+@fig-estimates plots the `Estimate` values calculated in the linear regressions.
+The regression analysis shows that the south and north central regions in the United States are the main regions that have seen union membership rates and political leanings show increased correlation.
+The south and north central regions have the greatest increase in their estimate values from the 1980s to the 2010s among the four regions.
 
-We do not report estimates like `p = 3.242e-15`, since these are computational zeros. Instead, we write $p < 0.001$ or $p \approx 0$, as appropriate.
+## Comparing Wages between Union and Non Union Workers
 
-We do not report quantities to unhelpful degrees of precision.
-Although there were 112,030,874 votes cast from voting-eligible population of 242,690,810 in the U.S.
-in 2022, it is not helpful to report turnout as `r (112030874 / 242690810) * 100`%; writing `r round((112030874 / 242690810) * 100, 1)`% suffices.
-
-## Comparing Distances between High- and Low-Speed Cars
+Our results for these t-tests determine if there is a statistically significant difference between union wages and non union wages in the 1980s and the 2010s.
 
@@
 
-# Create a binary variable for "high-speed":
-cars <- cars |> mutate(hs = ifelse(speed > mean(speed), 1, 0))
+# 1980s
+full_ntl_data_1980s <- full_ntl_data |>
+  filter(decade == "1980s")
+t_out_1980s <- t.test(full_ntl_data_1980s$nonunion_wage_2020_adj, full_ntl_data_1980s$union_wage_2020_adj)
 
-# Conduct t-test:
-t_out <- t.test(dist ~ hs, cars)
+t_lower_ci_1980s <- t_out_1980s$conf.int[1] |> round(2)
+t_upper_ci_1980s <- t_out_1980s$conf.int[2] |> round(2)
 
-t_lower_ci <- t_out$conf.int[1] |> round(2)
-t_upper_ci <- t_out$conf.int[2] |> round(2)
+# 2010s
+full_ntl_data_2010s <- full_ntl_data |>
+  filter(decade == "2010s")
+t_out_2010s <- t.test(full_ntl_data_2010s$nonunion_wage_2020_adj, full_ntl_data_2010s$union_wage_2020_adj)
+
+t_lower_ci_2010s <- t_out_2010s$conf.int[1] |> round(2)
+t_upper_ci_2010s <- t_out_2010s$conf.int[2] |> round(2)
 ```
 
-To report the results of a $t$-test, we do so in text, and perhaps in a well-formatted table as well, such as @tbl-ttesttable.
-Here, as above, we report the important details in text.
-For example, when we define "high-speed" cars as those traveling above the mean speed, the difference between the high-speed and low-speed group means is `r t_out$estimate[2] - t_out$estimate[1]`, with a 95% confidence interval that covers $(`r -t_upper_ci`, `r -t_lower_ci`)$.
-
-```{r tbl-ttesttable}
-#| echo: FALSE
-#| warning: FALSE
-#| tbl-cap: Distance by Speed Group
-
-parameters::model_parameters(t_out) |> 
-  parameters::print_md(
-    footer = "")
-```
-
-If I have tests of two outcomes from the same data, I can bind them together, as in @tbl-ttesttable2:
+The difference between union wages and non union wages means in the 1980s is `r t_out_1980s$estimate[2] - t_out_1980s$estimate[1]`, with a 95% confidence interval that covers $(`r -t_upper_ci_1980s`, `r -t_lower_ci_1980s`)$.
+The difference between union wages and non union wages means in the 2010s is `r t_out_2010s$estimate[2] - t_out_2010s$estimate[1]`, with a 95% confidence interval that covers $(`r -t_upper_ci_2010s`, `r -t_lower_ci_2010s`)$.
+The 95% confidence interval for the 1980s has a smaller range than the 95% confidence interval for the 2010s, demonstrating an increasing wage differential over time.
 
@@
 #| warning: FALSE
-#| tbl-cap: Distance and Square-root Distance by Speed Group
+#| tbl-cap: Wages by Union and Non Union Workers
 
+tab <- rbind(
+  model_parameters(t_out_1980s),
+  model_parameters(t_out_2010s)
+)
+
+tab |> 
+  print_md(
+    footer = "")
 ```

### 2024-01-12 Documents/Paper.qmd
added EDA graphs

@@
 
+```{r fig-cb}
+#| echo: false
+#| fig-cap: Collective Bargaining
+
+per_union_wage_data <- full_state_data |> # This object will be used to create the graph and also in the data subset
+  mutate(
+    map_color = case_when(
+      state_abbreviation == "CA" ~ "CA", # States we would like to highlight. first "" is the state abb, second "" is what will appear in legend
+      state_abbreviation == "AK" ~ "AK",
+      state_abbreviation == "OK" ~ "OK",
+      state_abbreviation == "FL" ~ "FL",
+      state_abbreviation == "HI" ~ "HI",
+      .default = NA
+    )
+  ) 
+States_by_union <- per_union_wage_data |> # use created table here
+  ggplot(aes(x = year)) +
+  geom_line(aes(y = percent_union, group = state_abbreviation),
+            color = "grey75", alpha =0.6) + # using group means the groupings wont appear in legend
+  geom_line(data = subset(per_union_wage_data, # use created object here
+                          !is.na(map_color)), #subset used to keep out non highlighted states
+            aes(y = percent_union, color = map_color),
+            linewidth = .75) +
+  geom_line(data = subset(full_ntl_data, year >= 1984),
+            aes(y = percent_union), color = "orange", linewidth = 1.5) +
+  annotate("text", x=1988, y=.15, label= "Federal Avg", color = "orange", size = 5) +
+  theme_minimal() +
+  scale_x_continuous(breaks = seq(min(per_union_wage_data$year), max(per_union_wage_data$year), by = 4)) + 
+  labs(title = "States by % of Labor Market Covered by Union",
+       x = "Year",
+       y = "% of Labor Market In Union",
+       color = "Selected States")
+States_by_union
+ggsave("../Documents/Graph_Images/States_by_union.png", width=7.29, height=5.41)
+```
+
+```{r fig-a}
+#| echo: false
+#| fig-cap: a
+
+per_collective_wage_data <- full_state_data |> # This object will be used to create the graph and also in the data subset
+  mutate(
+    map_color = case_when(
+      state_abbreviation == "CA" ~ "CA", # States we would like to highlight. first "" is the state abb, second "" is what will appear in legend
+      state_abbreviation == "AK" ~ "AK",
+      state_abbreviation == "HI" ~ "HI",
+      state_abbreviation == "FL" ~ "FL",
+      .default = NA
+    )
+  )
+
+per_collective_wage_data |> # use created table here
+  ggplot(aes(x = year)) +
+  geom_line(aes(y = percent_collective_bargain, group = state_abbreviation),
+            color = "grey75", alpha =0.6) + # using group means the groupings wont appear in legend
+  geom_line(data = subset(per_collective_wage_data, # use created object here
+                          !is.na(map_color)), #subset used to keep out non highlighted states
+            aes(y = percent_collective_bargain, color = map_color),
+            linewidth = .75) +
+  geom_line(data = subset(full_ntl_data, year >= 1984),
+            aes(y = percent_collective_bargain), color = "orange", linewidth = 1.5) +
+  annotate("text", x=1988, y=.17, label= "Federal Avg", color = "orange", size = 5) +
+  theme_minimal() +
+  scale_x_continuous(breaks = seq(min(per_collective_wage_data$year), max(per_collective_wage_data$year), by = 4)) + 
+  labs(title = "States by % of Labor Market Covered by a Collective Bargaining Agreement",
+       x = "Year",
+       y = "% of Labor Market in a Collective Bargaining Agreement",
+       color = "Selected States")
+```
+
+```{r fig-wages}
+#| echo: false
+#| fig-cap: Wages
+
+full_ntl_data |> 
+  filter(year >= 1984) |> 
+  ggplot(aes(x = year)) +
+  geom_line(aes(y = mean_income), color = "red") +
+  annotate("text", x=1995, y=85000, label= "Mean income", color = "red", size = 5) +
+  geom_line(aes(y = median_income), color = "blue") +
+  annotate("text", x=1995, y=65000, label= "Median income", color = "blue", size = 5) +
+  scale_x_continuous(breaks = seq(min(full_ntl_data$year), max(full_ntl_data$year), by = 4)) + 
+  theme_minimal() +
+  labs(
+    title = "Mean and Median Income in the United States Between 1984 and 2019",
+    x = "Year",
+    y = "Income in USD (Adjusted to 2020 values)"
+  )
+```
+
+```{r fig-b}
+#| echo: false
+#| fig-cap: b
+
+full_ntl_data |> 
+  filter(year >= 1984) |> 
+  ggplot(aes(x = year)) +
+  geom_line(aes(y = mean_income_2020_adj), color = "red") +
+  annotate("text", x=1995, y=145000, label= "Mean income", color = "red", size = 5) +
+  geom_line(aes(y = median_income_2020_adj), color = "blue") +
+  annotate("text", x=1995, y=115000, label= "Median income", color = "blue", size = 5) +
+  scale_x_continuous(breaks = seq(min(full_ntl_data$year), max(full_ntl_data$year), by = 4)) + 
+  theme_minimal() +
+  labs(
+    title = "Mean and Median Income in the United States Between 1984 and 2019 (2020 USD)",
+    x = "Year",
+    y = "Income in USD (Adjusted to 2020 values)"
+  )
+```
+
+```{r fig-d}
+#| echo: false
+#| fig-cap: d
+
+state_minimum_wage_data <- full_state_data |> # This object will be used to create the graph and also in the data subset
+  mutate(
+    map_color = case_when(
+      state_abbreviation == "CA" ~ "CA", # States we would like to highlight. first "" is the state abb, second "" is what will appear in legend
+      state_abbreviation == "AK" ~ "AK",
+      state_abbreviation == "OK" ~ "OK",
+      state_abbreviation == "FL" ~ "FL",
+      .default = NA
+    )
+  )
+
+state_minimum_wage_data |> # use created table here
+  ggplot(aes(x = year)) +
+  geom_line(aes(y = state_min_wage_2020_adj, group = state_abbreviation),
+            color = "grey75", alpha =0.6) + # using group means the groupings wont appear in legend
+  geom_line(data = subset(state_minimum_wage_data, # use created object here
+                          !is.na(map_color)), #subset used to keep out non highlighted states
+            aes(y = state_min_wage_2020_adj, color = map_color),
+            linewidth = .75) +
+  geom_line(aes(y = fed_min_wage_2020_adj), color = "orange", linewidth = 1.5) + # federal minimum wage
+  annotate("text", x=1990, y=6, label= "Federal Avg", color = "orange", size = 5) +
+  theme_minimal() +
+  scale_x_continuous(breaks = seq(min(state_minimum_wage_data$year), max(state_minimum_wage_data$year), by = 4)) + 
+  labs(title = "State Minimum Wage by Year (Adjusted for 2020 $)",
+       subtitle = "For states with an adjusted effective minimum wage above the federal level",
+       x = "Year",
+       y = "Minimum Wage (Adjusted to 2020 values)",
+       color = "Selected States")
+```
+
+```{r fig-e}
+#| echo: false
+#| fig-cap: e
+
+effective_minimum_wage_data <- full_state_data |> 
+  filter(effective_min_wage_2020_adj != fed_min_wage_2020_adj) |> 
+  mutate(
+    map_color = case_when(
+      state_abbreviation == "CA" ~ "CA",
+      state_abbreviation == "AK" ~ "AK",
+      state_abbreviation == "CT" ~ "CT",
+      state_abbreviation == "HI" ~ "HI",
+      .default = NA
+    )
+  )
+
+effective_minimum_wage_data |> 
+  ggplot(aes(x = year)) +
+  geom_line(aes(y = effective_min_wage_2020_adj, group = state_abbreviation),
+            color = "grey75", alpha =0.6) +
+  geom_line(data = subset(effective_minimum_wage_data, !is.na(map_color)),
+            aes(y = effective_min_wage_2020_adj, color = map_color),
+            linewidth = .75) +
+  geom_line(aes(y = fed_min_wage_2020_adj), color = "orange", linewidth = 1.5) +
+  annotate("text", x=1990, y=6, label= "Federal Avg", color = "orange", size = 5) +
+  theme_minimal() +
+  scale_x_continuous(breaks = seq(min(effective_minimum_wage_data$year), max(effective_minimum_wage_data$year), by = 4)) + 
+  labs(title = "State Effective Minimum Wage by Year (Adjusted for 2020 $)",
+       subtitle = "For states with an adjusted effective minimum wage above the federal level",
+       x = "Year",
+       y = "Effective Minimum Wage (Adjusted to 2020 values)",
+       color = "Selected States")
+```
+
+## House Elections
+
+```{r fig-c}
+#| echo: false
+#| fig-cap: c
+
+map_data <- full_state_data |> 
+  select(c("year", "state", "party_state_average")) |> 
+  distinct(year, state, .keep_all = TRUE) |> 
+  filter(year %% 4==0) |> 
+  filter(!is.na(party_state_average)) |> 
+  full_join(
+    tibble(
+      "year" = seq(from=1984, to=2016, by=4),
+      "state" = rep("District of Columbia", 9),
+      "party_state_average" = rep(.5, 9)),
+    by = c("year", "state", "party_state_average")
+    )
+
+plot_usmap(data = map_data, values = "party_state_average", color = "black") +
+  facet_wrap(~year, nrow = 4, ncol= 3) +
+  theme(legend.position = "bottom") +
+  scale_fill_gradient(low = "red", high = "blue", name = "Proportion") +
+  labs(
+    title = "Proportion of House Delegation that is Democratic"
+  )
+```
+
+## Collective Bargaining and Wages
+
+```{r fig-w}
+#| echo: false
+#| fig-cap: w
+
+full_ntl_data |> 
+  ggplot(aes(x = year)) +
+  geom_line(aes(y = union_wage_2020_adj), color = "blue") +
+  annotate("text", x=1993, y=28, label= "Union Wages", color = "blue", size = 5) +
+  geom_line(aes(y = nonunion_wage_2020_adj), color = "red") +
+  annotate("text", x=1993, y=22, label= "Non Union Wages", color = "red", size = 5) +
+  scale_x_continuous(breaks = seq(min(full_ntl_data$year), max(full_ntl_data$year), by = 4)) + 
+  theme_minimal() +
+  labs(
+    title = "Union vs. Non Union Wages Over Time",
+    x = "Year",
+    y = "Wages (Adjusted to 2020 values)"
+  )
+```
+
+```{r fig-x}
+#| echo: false
+#| fig-cap: x
+
+full_ntl_data |> 
+  ggplot(aes(x = percent_collective_bargain)) +
+  geom_point(aes(y = nonunion_wage_2020_adj), color = "blue") +
+  geom_jitter(aes(y = union_wage_2020_adj), color = "red") +
+  theme_light() +
+  facet_wrap(~decade) +
+  labs(
+    title = "Union vs. Non-Union Wages by Decade",
+    subtitle = "Red represents union wages and blue represents non-union wages",
+    x = "Percent of Workers Under Collective Bargaining Agreement",
+    y = "Wages (Adjusted to 2020 values)"
+  )
+```
+
+## Collective Bargaining and House Elections
+
+```{r fig-y}
+#| echo: false
+#| fig-cap: y
+
+full_state_data |> 
+  filter(year %% 4==0) |>
+  ggplot(aes(x = percent_collective_bargain, y = party_state_average)) +
+  geom_jitter(aes(color = regions), alpha = 0.5, width = 0.1, height = 0.1) +
+  geom_smooth(method = "lm", color = "black") +
+  facet_wrap(~decade) +
+  theme_light() +
+  labs(
+    title = "% Collective Bargaining vs Proportion of House Delegation That is Democratic",
+    x = "Percent of Workers Covered by Collective Bargaining",
+    y = "Proportion of House Delegation that is Democratic",
+    color = "Region"
+  )
+```
+
+```{r fig-z}
+#| echo: false
+#| fig-cap: z
+
+full_state_data |> 
+  filter(year %% 4==0) |>
+  ggplot(aes(x = percent_collective_bargain, y = party_state_average)) +
+  geom_jitter(aes(color = regions), alpha = 0.15, width = 0.1, height = 0.1) +
+  geom_smooth(method = "lm", aes(group = regions, color = regions), se = FALSE) +
+  facet_wrap(~decade) +
+  theme_light() +
+  labs(
+    title = "% Collective Bargaining vs Proportion of House Delegation That is Democratic",
+    subtitle = "Regressed by Region",
+    x = "Percent of Workers Covered by Collective Bargaining",
+    y = "Proportion of House Delegation that is Democratic",
+    color = "Region"
+  )
+```
+
 # Exploring Correlations: Linear Regression Analyses and T-Tests on Collective Bargaining and Wage Disparities

### 2024-01-12 Documents/Paper.qmd
work on describing data

@@
 
-When conducting a regression analysis, it demonstrates that the correlation between collective bargaining and political leanings has gotten stronger in the last four decades from 1984 to 2019.
+When conducting a regression analysis, it demonstrates that the correlation between collective bargaining and political representation in the House has gotten stronger in the last four decades from 1984 to 2019.
 Additionally, when conducting t-tests among union and non union wages, there is a statistically significant difference that demonstrates an increasing wage differential over the same time period.
@@
 
+The data being analyzed is from a collection of five data sets.
+The first two data sets, `union_wages` and `union_states` are gathered from Tidy Tuesday, although the original data comes from the Union Membership and Coverage Database at Georgia State University.
+The  tables relevant to this paper detailed union membership counts by state and year, as well as union and non union wages nationally.
+The third data set, `house`, is also from Tidy Tuesday, although the original data comes from the MIT Election Data and Science Lab.
+The data includes every person who ran for a seat in the House of Representatives, but we only selected the election winner.
+The fourth data set, `income_distribution`, is also from Tidy Tuesday, although the original data comes from the U.S.
+Census.
+The data tracks the mean and median income nationally in the United States.
+The fifth data set, `minimum_wage`, comes from the U.S.
+Department of Labor and tracks minimum wage at the state level and the federal minimum wage.
+
+ADD SPECIFICS ABOUT CLEANING EACH OF THE DATA SETS, HOW PARTY STATE AVERAGE WAS MADE, ETC
+
+The five data sets are combined into two data sets: one that stores data at the state level and another that stores data at the national level.
+`full_state_data` joins the `union_states`, `house`, and `minimum_wage` data sets.
+There are `8730` observations and each row represents a year and a state.
+`full_ntl_data` joins the `union_wages` and `income_distribution` data sets.
+There are `36` observations and each row represents a distinct year.
+Variables regarding union employment are not provided at the national level, only at the state level, so they were summarized and added to the national data set.
+
+Among the original five data sets, there is data recorded from 1967 to 2022.
+Not all of the data sets cover that time frame, so once joined into two data sets, the data was filtered to only include the years 1984 to 2019.
+
 This section describes the data we analyze.
@@
 Of course, we do not cite software that we do not use.
+
+## Wage Data
 
@@
 
-## House Elections
+## House of Representatives Data
 
@@
 
-## Collective Bargaining and Wages
+## Collective Bargaining and Wage Data
 
@@
 
-## Collective Bargaining and House Elections
+## Collective Bargaining and House of Representatives Data
 

### 2024-01-12 Documents/Presentation.qmd
fixed third question
added biblgiography

@@
 1.  Is there any relationship between collective bargaining rates among workers to house delegation makeup?
-2.  Is there any relationship between collective bargaining rates among workers and wages
-3.  Is there any relationship between collective bargaining rates among workers and wages?
+2.  Is there any relationship between collective bargaining rates among workers and wages?
+3.  Do collective bargaining rates effect wealth inequality?
 
@@
 ## Bibliography {.smaller}
+
+"Johnthegeek". 2021. “Wealth and income over time.” GitHub. https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-02-09/readme.md.
+
+"Johnthegeek". 2023. “Union Membership in the United States.” GitHub. https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-09-05/readme.md.
+
+"Johnthegeek". 2023. “US House Election Results.” GitHub. https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-11-07/readme.md.
+
+"LisleJoem". 2021. “US Minimum Wage by State from 1968 to 2020 and 2020 Equivalent Dollars.” Kaggle. https://www.kaggle.com/datasets/lislejoem/us-minimum-wage-by-state-from-1968-to-2017.

### 2024-01-12 Documents/Presentation.qmd
fixed t test in pres

@@
 -   **Wage**
-    -   Our wage data was obtained through the U.S. Department of Labor and Tidy Tuesday (originally from U.S. census data).
-    -   This data tracked the mean and median income nationally in the United States.
-    -   Another table tracked minimum wage at the state level.
+    -   Our wage data was obtained through Kaggle that pulls data from the U.S. Department of Labor and Tidy Tuesday (originally from U.S. census data).
+    -   The data obtained through Kaggle data tracked tracked minimum wage at the state level.
+    -    The Tidy Tuesday data contained the mean and median income nationally in the United States.
 
@@
     -   Other minor changes included renaming variables and recoding a few outlier observations in the `party` variable.
--   For the union data, we filtered the datasets down to not include union sectors (ex. construction-unions) and added columns to adjust for inflation.
+-   For the union data, we filtered the data sets down to not include union sectors (ex. construction-unions).
 -   For our wage data, we filtered out race, as this was outside of the scope of our study.
@@
 -   Our last data manipulation was combining all of our tables into two data sets: one that stores data at a national level, and another that stores data at a state level.
+-   Inflation adjusted values added to national data set.
 -   Some of the variables were only provided at the state level so they were summarized and added to the national data set.
@@
     -   2010s: t = 22.846, df = 9.4096, p-value = 1.462e-09
-        -   mean of x = 3.786; mean of y = 25.664
+        -   mean of x = 30.533; mean of y = 25.664
         -   95% CI: (3.786, 5.952)

### 2024-01-12 Documents/Presentation.qmd
ramiros lit review and bib

@@
 ## Literature Review
+
+-   The decline of organized labor in the United States coincided with a large increase in wage inequality.
+From 1973 to 2007, men’s wage inequality increased by 40 percent, from .25 to .35, with most of the rise unfolding from 1978 to 2000.
+
+-   ..even in a period of labor weakness (2000-2014), union activism increases the wages of union members.
+
+-   “Particularly since the early 2000s, when the Bush administration’s NLRB appointments led to a series of decisions that were especially hostile to organized labor ([Liebman 2007](https://muse.jhu.edu/article/659928#b25); [Farber 2014](https://muse.jhu.edu/article/659928#b15)).
 
@@
 "LisleJoem". 2021. “US Minimum Wage by State from 1968 to 2020 and 2020 Equivalent Dollars.” Kaggle. https://www.kaggle.com/datasets/lislejoem/us-minimum-wage-by-state-from-1968-to-2017.
+
+## Bibliography pt 2 {.smaller}
+
+Western, B., & Rosenfeld, J. (2011). Unions, norms, and the rise in U.S. wage inequality. American Sociological Review, 76(4), 513-537. \
+Retrieved from <https://www.proquest.com/scholarly-journals/unions-norms-rise-u-s-wage-inequality/docview/908813088/se-2>
+
+Wilmers, Nathan.
+“Labor Unions as Activist Organizations: A Union Power Approach to Estimating Union Wage Effects.” Social Forces, vol.
+95, no.
+4, 2017, pp.
+1451–77.
+JSTOR, <http://www.jstor.org/stable/26166881> Accessed 9 Jan.
+2024.
+
+Freeman, Richard.
+et al.
+How Does Declining Unionism Affect the American Middle Class and Intergenerational Mobility?
+/ Richard Freeman, Eunice Han, David Madland, Brendan V.
+Duke.
+Cambridge, Mass: National Bureau of Economic Research, 2015.
+Print.

### 2024-01-12 Documents/Presentation.qmd
update t test values

--- 2024-01-12 Documents/Presentation.qmd
@@
 ## Literature Review
+
+-   The decline of organized labor in the United States coincided with a large increase in wage inequality.
+From 1973 to 2007, men's wage inequality increased by 40 percent, from .25 to .35, with most of the rise unfolding from 1978 to 2000.
+
+-   ..even in a period of labor weakness (2000-2014), union activism increases the wages of union members.
+
+-   "Particularly since the early 2000s, when the Bush administration's NLRB appointments led to a series of decisions that were especially hostile to organized labor ([Liebman 2007](https://muse.jhu.edu/article/659928#b25); [Farber 2014](https://muse.jhu.edu/article/659928#b15)).
 
@@
         -   95% CI: (5.347, 6.514)
-    -   2010s: t = 22.846, df = 9.4096, p-value = 1.462e-09
+    -   2010s: t = 9.5488, df = 15.629, p-value = 6.393e-08
         -   mean of x = 30.533; mean of y = 25.664
@@
 
-"Johnthegeek". 2021. “Wealth and income over time.” GitHub. https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-02-09/readme.md.
+"Johnthegeek". 2021. "Wealth and income over time." GitHub. https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-02-09/readme.md.
 
-"Johnthegeek". 2023. “Union Membership in the United States.” GitHub. https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-09-05/readme.md.
+"Johnthegeek". 2023. "Union Membership in the United States." GitHub. https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-09-05/readme.md.
 
-"Johnthegeek". 2023. “US House Election Results.” GitHub. https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-11-07/readme.md.
+"Johnthegeek". 2023. "US House Election Results." GitHub. https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-11-07/readme.md.
 
-"LisleJoem". 2021. “US Minimum Wage by State from 1968 to 2020 and 2020 Equivalent Dollars.” Kaggle. https://www.kaggle.com/datasets/lislejoem/us-minimum-wage-by-state-from-1968-to-2017.
+"LisleJoem". 2021. "US Minimum Wage by State from 1968 to 2020 and 2020 Equivalent Dollars." Kaggle. https://www.kaggle.com/datasets/lislejoem/us-minimum-wage-by-state-from-1968-to-2017.
+
+## Bibliography pt 2 {.smaller}
+
+Western, B., & Rosenfeld, J. (2011). Unions, norms, and the rise in U.S. wage inequality. American Sociological Review, 76(4), 513-537.\
+Retrieved from <https://www.proquest.com/scholarly-journals/unions-norms-rise-u-s-wage-inequality/docview/908813088/se-2>
+
+Wilmers, Nathan.
+"Labor Unions as Activist Organizations: A Union Power Approach to Estimating Union Wage Effects." Social Forces, vol.
+95, no.
+4, 2017, pp.
+1451--77.
+JSTOR, <http://www.jstor.org/stable/26166881> Accessed 9 Jan.
+2024.
+
+Freeman, Richard.
+et al.
+How Does Declining Unionism Affect the American Middle Class and Intergenerational Mobility?
+/ Richard Freeman, Eunice Han, David Madland, Brendan V.
+Duke.
+Cambridge, Mass: National Bureau of Economic Research, 2015.
+Print.

### 2024-01-12 Documents/Paper.qmd
fix graph labels

@@
 
-1.    Is there any relationship between collective bargaining rates among workers to House delegation makeup?
+1.  Is there any relationship between collective bargaining rates among workers to house delegation makeup?
 2.  Is there any relationship between collective bargaining rates among workers and wages?
-3.    Is there any relationship between collective bargaining rates among workers and wages?
+3.  Do collective bargaining rates effect wealth inequality?
 
@@
 
-## Wage Data
+## Union Data
 
-```{r fig-cb}
+```{r fig-uniondata}
 #| echo: false
-#| fig-cap: Collective Bargaining
-
-per_union_wage_data <- full_state_data |> # This object will be used to create the graph and also in the data subset
-  mutate(
-    map_color = case_when(
-      state_abbreviation == "CA" ~ "CA", # States we would like to highlight. first "" is the state abb, second "" is what will appear in legend
-      state_abbreviation == "AK" ~ "AK",
-      state_abbreviation == "OK" ~ "OK",
-      state_abbreviation == "FL" ~ "FL",
-      state_abbreviation == "HI" ~ "HI",
-      .default = NA
-    )
-  ) 
-States_by_union <- per_union_wage_data |> # use created table here
-  ggplot(aes(x = year)) +
-  geom_line(aes(y = percent_union, group = state_abbreviation),
-            color = "grey75", alpha =0.6) + # using group means the groupings wont appear in legend
-  geom_line(data = subset(per_union_wage_data, # use created object here
-                          !is.na(map_color)), #subset used to keep out non highlighted states
-            aes(y = percent_union, color = map_color),
-            linewidth = .75) +
-  geom_line(data = subset(full_ntl_data, year >= 1984),
-            aes(y = percent_union), color = "orange", linewidth = 1.5) +
-  annotate("text", x=1988, y=.15, label= "Federal Avg", color = "orange", size = 5) +
-  theme_minimal() +
-  scale_x_continuous(breaks = seq(min(per_union_wage_data$year), max(per_union_wage_data$year), by = 4)) + 
-  labs(title = "States by % of Labor Market Covered by Union",
-       x = "Year",
-       y = "% of Labor Market In Union",
-       color = "Selected States")
-States_by_union
-ggsave("../Documents/Graph_Images/States_by_union.png", width=7.29, height=5.41)
-```
-
-```{r fig-a}
-#| echo: false
-#| fig-cap: a
+#| fig-cap: Union Data
 
@@
 
-```{r fig-wages}
+## Income Data
+```{r fig-incomedata}
 #| echo: false
-#| fig-cap: Wages
+#| fig-cap: Income Data
 
@@
     x = "Year",
-    y = "Income in USD (Adjusted to 2020 values)"
+    y = "Income in USD"
   )
@@
 
-```{r fig-b}
+```{r fig-incomedataadj}
 #| echo: false
-#| fig-cap: b
+#| fig-cap: Income Data (adjusted)
 
@@
 
-```{r fig-d}
+```{r fig-wagedata}
 #| echo: false
-#| fig-cap: d
-
-state_minimum_wage_data <- full_state_data |> # This object will be used to create the graph and also in the data subset
-  mutate(
-    map_color = case_when(
-      state_abbreviation == "CA" ~ "CA", # States we would like to highlight. first "" is the state abb, second "" is what will appear in legend
-      state_abbreviation == "AK" ~ "AK",
-      state_abbreviation == "OK" ~ "OK",
-      state_abbreviation == "FL" ~ "FL",
-      .default = NA
-    )
-  )
-
-state_minimum_wage_data |> # use created table here
-  ggplot(aes(x = year)) +
-  geom_line(aes(y = state_min_wage_2020_adj, group = state_abbreviation),
-            color = "grey75", alpha =0.6) + # using group means the groupings wont appear in legend
-  geom_line(data = subset(state_minimum_wage_data, # use created object here
-                          !is.na(map_color)), #subset used to keep out non highlighted states
-            aes(y = state_min_wage_2020_adj, color = map_color),
-            linewidth = .75) +
-  geom_line(aes(y = fed_min_wage_2020_adj), color = "orange", linewidth = 1.5) + # federal minimum wage
-  annotate("text", x=1990, y=6, label= "Federal Avg", color = "orange", size = 5) +
-  theme_minimal() +
-  scale_x_continuous(breaks = seq(min(state_minimum_wage_data$year), max(state_minimum_wage_data$year), by = 4)) + 
-  labs(title = "State Minimum Wage by Year (Adjusted for 2020 $)",
-       subtitle = "For states with an adjusted effective minimum wage above the federal level",
-       x = "Year",
-       y = "Minimum Wage (Adjusted to 2020 values)",
-       color = "Selected States")
-```
-
-```{r fig-e}
-#| echo: false
-#| fig-cap: e
+#| fig-cap: Wage Data
 
@@
 
-```{r fig-c}
+```{r fig-housedata}
 #| echo: false
-#| fig-cap: c
+#| fig-cap: House Data
 
@@
 
-```{r fig-w}
+```{r fig-unionwagedata}
 #| echo: false
-#| fig-cap: w
+#| fig-cap: Union Wage Data
 
@@
 
-```{r fig-x}
+```{r fig-collectivebargain}
 #| echo: false
-#| fig-cap: x
+#| fig-cap: Collective Bargain Data
 
@@
 
-```{r fig-y}
+```{r fig-collectivebargaindecade}
 #| echo: false
-#| fig-cap: y
+#| fig-cap: Collective Bargain Data by Decade
 
@@
 
-```{r fig-z}
+```{r fig-collectivebargainregion}
 #| echo: false
-#| fig-cap: z
+#| fig-cap: Collective Bargain Region by Decade and Region
 
@@
 The difference between union wages and non union wages means in the 2010s is `r t_out_2010s$estimate[2] - t_out_2010s$estimate[1]`, with a 95% confidence interval that covers $(`r -t_upper_ci_2010s`, `r -t_lower_ci_2010s`)$.
-The 95% confidence interval for the 1980s has a smaller range than the 95% confidence interval for the 2010s, demonstrating an increasing wage differential over time.
+The 95% confidence interval for the 1980s has a smaller range than the 95% confidence interval for the 2010s.
 
-```{r tbl-ttesttable2}
+```{r}
 #| echo: FALSE
@@
 
-tab <- rbind(
-  model_parameters(t_out_1980s),
-  model_parameters(t_out_2010s)
-)
+tab <- tibble(
+  "metric" = c("t", "df", "p-value", "95%-CI-LB", "95%-CI-UB", "Mean_of_X", 
+               "Mean_of_Y"),
+  "t_out_1980s" = c(t_out_1980s$statistic, t_out_1980s$parameter, t_out_1980s$p.value, t_out_1980s$conf.int, t_out_1980s$estimate),
+  "t_out_2010s" = c(t_out_2010s$statistic, t_out_2010s$parameter, t_out_2010s$p.value, t_out_2010s$conf.int, t_out_2010s$estimate))
 

### 2024-01-12 Documents/Paper.qmd
update data description

@@
 The first two data sets, `union_wages` and `union_states` are gathered from Tidy Tuesday, although the original data comes from the Union Membership and Coverage Database at Georgia State University.
-The  tables relevant to this paper detailed union membership counts by state and year, as well as union and non union wages nationally.
+The  tables relevant to this paper detail union membership counts by state and year, as well as union and non union wages nationally.
 The third data set, `house`, is also from Tidy Tuesday, although the original data comes from the MIT Election Data and Science Lab.
@@
 
-ADD SPECIFICS ABOUT CLEANING EACH OF THE DATA SETS, HOW PARTY STATE AVERAGE WAS MADE, ETC
+All of the data sets undergo basic cleaning so that the columns are correctly formatted, variable names are consistent, so there are no NA values, and relevant variables are selected.
+For the house election data, the table is manipulated so that only rows of candidates who won their election remain.
+A binary variable is created in which we code the outcome as 1 if the winner was a Democrat and code the outcome as 0 if the winner was a Republican.
+An additional variable is created to show the proportion of the House delegation that is democratic for every combination of years and states.
+For the union data, it is filtered to not include union sectors, such as construction unions, and instead look at all unions.
+For the income distribution data, race is filtered out because it is outside of the scope of this study.
 
@@
 Variables regarding union employment are not provided at the national level, only at the state level, so they were summarized and added to the national data set.
-
+Economic variables in the national data are not originally adjusted to 2020 values, so we calculate the rate from the state data that contains adjusted and non-adjusted values for federal minimum wage, and then apply that rate to the national data.
+This allows all of the economic data to be standardized and account for inflation.
 Among the original five data sets, there is data recorded from 1967 to 2022.
@@
 
-This section describes the data we analyze.
-We describe the source of the data, and its primary features.
-We cite our data.
-We describe the methods we use to answer our question and to test our hypotheses.
+With two cleaned data sets, we conduct our exploratory data analysis using R version `r paste(R.version$major, R.version$minor, sep = ".")` [@rcoreteam23].
+We rely on several elements of the `tidyverse` [@wickhamtidyverse].
 
-If our data were `cars`, loaded in a chunk above, we note that our data have `r nrow(cars)` observations. Our unit of analysis is the cars; each row represents a different car that was measured.
+We cite our data. We describe the methods we use to answer our question and to test our hypotheses.
 
-We refer to concepts and label them appropriately.
-We state our outcome and how it is measured: "Our outcome is stopping distance measured in feet." We state our key predictor and how it is measured: "Our key predictor is the car's speed, measured in miles per hour."
+"Our outcome is stopping distance measured in feet." We state our key predictor and how it is measured: "Our key predictor is the car's speed, measured in miles per hour."
 
@@
 These particular names are almost never of interest or use to the reader.
-
-We explain important decisions and codings in this section.
-"We collect our data from source X.
-We code the outcome as 1 if the registrant turned out in the 2022 election, 0 if they did not." A table can serve as an efficient way to detail several such decisions.
-
-Where there are less-critical details that we implement to improve our analysis or presentation, we do not explain them in the paper.
-Our paper does not say, for example, "we reorder the levels of the factor variable from alphabetical 'high, low, medium' to the sensible 'low, medium, high.'" Of course, we implement that reordering, but it should not appear in our paper.
-
-We cite the software we use.
-For example, we conduct our analysis using R version `r paste(R.version$major, R.version$minor, sep = ".")` [@rcoreteam23].
-We rely on several elements of the `tidyverse` [@wickhamtidyverse].
-Of course, we do not cite software that we do not use.
 
@@
 
+From our data analysis it can be inferred that collective bargaining agreements have become a very political concept.
+The charts in our exploratory data analysis and regression models show this.
+This partially supports our hypothesis.
+While the correlation is a newer phenomenon, the relationship appears to be strong today.
+The relationship between unions and wages/income in the United States is much harder to draw inferences.
+While our t-tests and some charts show that unions can help workers earn more, other graphs show that unions have a negative effect on incomes.
+Our second hypothesis thus remains without strong support.
+
+-   Union wage data at the state level
+-   Data past 2019
+-   A greater amount and complexity of models
+
 We remind the reader what this paper was about, why it was important, and what we found.

### 2024-01-12 Documents/Presentation.qmd
revised pres

@@
 ::: columns
-::: {.column width="60%"}
+::: {.column width="30%"}
 -   Union and collective bargaining rates have collapsed over the last 40 years, although may just beginning to see a come back in the last two.
@@
 
-::: {.column width="40%"}
+::: {.column width="70%"}
 ![Plot 3](./Graph_Images/States_by_CB.png)
@@
 ::: columns
-::: {.column width="60%"}
+::: {.column width="30%"}
 -   The country has generally grown more republican in the last 40 years, although data in the last 10 years show the pendulum might now be swinging the other way.
@@
 
-::: {.column width="40%"}
+::: {.column width="70%"}
 ![Plot 4](./Graph_Images/House_by_year.png)
@@
 However, for individual workers, being part of a union remains a way to increase wages.
--   We do not have data to compare wages and union rates by state, which would likely tell a different story.
+-   This graph should not be interpreted as proving that decreased union activity leads to increased wages.
+-   It is interesting that it appears that as union percentages decrease, the gap between union and non union data appears to be shrinking.
+Still, this correlation does not take into account the counts of workers in each group and other confounding variables could be effecting it as well.
 :::
@@
 ::: columns
-::: {.column width="60%"}
+::: {.column width="70%"}
 -   Our regression analysis shows that the south and north central regions in the United States are the main regions that have seen union membership rates and political leanings show increased correlation.
+    - Plot 8 measures the regression estimates for the 10 SLR models that we ran. The increases in the 2010s grouping shows the increased politicization of unions in the U.S.
 -   T-test between union and non union wages
@@
 
-::: {.column width="40%"}
+::: {.column width="30%"}
 ![Plot 8](./Graph_Images/Regression_plot.png)

### 2024-01-12 Documents/Paper.qmd
fix graphs

@@
 #| echo: false
+#| warning: FALSE
 #| fig-cap: Union Data
@@
 #| echo: false
+#| warning: FALSE
 #| fig-cap: Income Data
@@
 #| echo: false
+#| warning: FALSE
 #| fig-cap: Income Data (adjusted)
@@
 #| echo: false
+#| warning: FALSE
 #| fig-cap: Wage Data
@@
 #| echo: false
+#| warning: FALSE
 #| fig-cap: House Data
@@
 #| echo: false
+#| warning: FALSE
 #| fig-cap: Union Wage Data
@@
 #| echo: false
+#| warning: FALSE
 #| fig-cap: Collective Bargain Data
@@
 #| echo: false
+#| warning: FALSE
 #| fig-cap: Collective Bargain Data by Decade
@@
 #| echo: false
+#| warning: FALSE
 #| fig-cap: Collective Bargain Region by Decade and Region
@@
 
-Our results for these t-tests determine if there is a statistically significant difference between union wages and non union wages in the 1980s and the 2010s.
+Our results seen in @tbl-ttest for these t-tests determine if there is a statistically significant difference between union wages and non union wages in the 1980s and the 2010s.
 
@@
 
-```{r}
+```{r tbl-ttest}
 #| echo: FALSE

### 2024-01-12 Documents/Paper.qmd
update labels

@@
 
-Here we go deeper into the intellectual debate, the political and social context of our investigation.
-To give the reader a clear sense of why we are writing this paper, we describe the relevant scholarly, technical, or popular literature.
-We give this section a meaningful *substantive* title; it is not entitled "Literature Review", for example.
+Analyzing Union Datasets along with US House of Representative Data allowed us to examine and search for correlation between wages and the increase/decrease of Union participation in national regions.
+We investigated decreases in wages and union participation, along with widening gaps in income between non union, and union members, and the overall increases in wages.
+Our study found strong correlation with Western and Rosenfeld's (2011) Unions, norms and the rise in the U.S.
+wage inequality which found “The decline of organized labor in the United States coincided with a large increase in wage inequality.
+From 1973 to 2007”.
+
+Our graphs convey, and affirm Nathan Wilmers (2017) reference point that “even in a period of labor weakness (2000 -2014), union activism increases the wages of union members.
+Utilizing the House of Representative data set of both Democrats and Republicans we also were able to find correlation between the downward slope in union wages and enrollment during a Republican presidency which also came with a large republican majority in the house during the Bush administration.
+Particularly since the early 2000s, when the Bush administration’s NLRB appointments led to a series of decisions that were especially hostile to organized labor (Liebman 2007; Farber 2014) Milkman, Ruth (2017)
+
 We cite at least three published, peer-reviewed scholarly works.
@@
   scale_x_continuous(breaks = seq(min(per_collective_wage_data$year), max(per_collective_wage_data$year), by = 4)) + 
-  labs(title = "States by % of Labor Market Covered by a Collective Bargaining Agreement",
+  labs(title = "States by % of Labor Market Covered by a CB Agreement",
        x = "Year",
-       y = "% of Labor Market in a Collective Bargaining Agreement",
+       y = "% of Labor Market in a CB Agreement",
        color = "Selected States")
@@
   labs(
-    title = "Mean and Median Income in the United States Between 1984 and 2019",
+    title = "Mean and Median Income in the United States",
     x = "Year",
@@
   labs(
-    title = "Mean and Median Income in the United States Between 1984 and 2019 (2020 USD)",
+    title = "Mean and Median Income in the United States (2020 USD)",
     x = "Year",
@@
   scale_x_continuous(breaks = seq(min(effective_minimum_wage_data$year), max(effective_minimum_wage_data$year), by = 4)) + 
-  labs(title = "State Effective Minimum Wage by Year (Adjusted for 2020 $)",
+  labs(title = "State Effective Minimum Wage by Year (2020 USD)",
        subtitle = "For states with an adjusted effective minimum wage above the federal level",
@@
   labs(
-    title = "% Collective Bargaining vs Proportion of House Delegation That is Democratic",
+    title = "% CB vs Proportion of House Delegation That is Democratic",
     x = "Percent of Workers Covered by Collective Bargaining",
@@
   labs(
-    title = "% Collective Bargaining vs Proportion of House Delegation That is Democratic",
+    title = "% CB vs Proportion of House Delegation That is Democratic",
     subtitle = "Regressed by Region",
@@
 
+```{r fig-qqplot1980s}
+#| echo: false
+#| fig-cap: QQ Plot 1980s
+
+state_data_1980 <- full_state_data |> 
+  filter(year >= 1984 & year < 1990)
+regression_1980s <- lm(party_state_average~percent_collective_bargain, data = state_data_1980)
+qqnorm(regression_1980s$resi)
+qqline(regression_1980s$resi)
+```
+
+```{r fig-qqplot2010s}
+#| echo: false
+#| fig-cap: QQ Plot 2010s
+
+state_data_2010 <- full_state_data |> 
+  filter(year >= 2010 & year < 2020)
+regression_2010s <- lm(party_state_average~percent_collective_bargain, data = state_data_2010)
+qqnorm(regression_2010s$resi)
+qqline(regression_2010s$resi)
+```
+The QQ plot in @fig-qqplot1980s from the 1980s shows a very bimodal distribution, while the QQ plot in @fig-qqplot2010s from the 2010s regression model shows a more normal, slightly light-tailed distribution.
+
 ## Comparing Wages between Union and Non Union Workers

### 2024-01-12 Documents/Paper.qmd
refined paper

@@
 Using regression testing we show that for unions and political leanings, this relationship has surged and deepened in the last four decades from 1984 to 2019"
-#bibliography: main.bib
+bibliography: myref.bib
 ---
@@
 
-# \[Our Substance and Context Section Title Here\]
+# Historical Reference and Overlap Points of Union Wages and Effects of Collective Bargaining Agreements
 
-Analyzing Union Datasets along with US House of Representative Data allowed us to examine and search for correlation between wages and the increase/decrease of Union participation in national regions.
+Analyzing Union data sets along with US House of Representative Data allowed us to examine and search for correlation between wages and the increase/decrease of Union participation in national regions.
 We investigated decreases in wages and union participation, along with widening gaps in income between non union, and union members, and the overall increases in wages.
-Our study found strong correlation with Western and Rosenfeld's (2011) Unions, norms and the rise in the U.S.
+Our study found strong correlation with @western2011unions.
+Unions, norms and the rise in the U.S.
 wage inequality which found “The decline of organized labor in the United States coincided with a large increase in wage inequality.
-From 1973 to 2007”.
+From 1973 to 2007.”
 
-Our graphs convey, and affirm Nathan Wilmers (2017) reference point that “even in a period of labor weakness (2000 -2014), union activism increases the wages of union members.
+Our graphs convey, and affirm @wilmers2017labor, reference point that “even in a period of labor weakness (2000 -2014), union activism increases the wages of union members.
 Utilizing the House of Representative data set of both Democrats and Republicans we also were able to find correlation between the downward slope in union wages and enrollment during a Republican presidency which also came with a large republican majority in the house during the Bush administration.
-Particularly since the early 2000s, when the Bush administration’s NLRB appointments led to a series of decisions that were especially hostile to organized labor (Liebman 2007; Farber 2014) Milkman, Ruth (2017)
-
-We cite at least three published, peer-reviewed scholarly works.
-For example, we could cite @mooree20 or @moorav12[^1], which we discussed in class.[^2] We only cite others' work in our paper when it enhances the reader's understanding of what we, the authors of this paper, are doing.
-We connect everything we cite to *our* investigation; this is our original research, not a book report or an annotated bibliography.
-
-[^1]: There should always be a space before the "(" in a citation date.
-
-[^2]: To cite a paper within parentheses, use, e.g., [@moore12].
-
-We do not cite paper titles or journal names, unless our paper is about someone else's paper or about the set of articles in a journal.
-We do not cite authors' first, given names.
-We can refer to either what an author does, or what a paper does, but we should be consistent.
-For example, "@moorav12 argue that we should \ldots" refers to what the authors do; "@moorav12 argues that we should \ldots" refers to what the paper -- @moorav12 -- does.
-
-In order to integrate citations into the References section below, we add entries into our file `main.bib`.
-This is a plain-text file that we edit in RStudio.
-We store `main.bib` in the same folder as our paper's `.qmd` and `.pdf` files.
-Its entries are formatted so that they can be knit to `.pdf`; see [https://j.mp/2UzTXEZ](https://www.overleaf.com/learn/latex/Bibliography_management_with_bibtex#The_bibliography_file) for example entries for articles, books, and miscellaneous.
-We can get these entries automatically from Google Scholar by turning on BibTeX in the Google Scholar Settings - Bibliography Manager.
-Perhaps we use a tool like free, open-source BibDesk to help us manage the `.bib` file.
+Particularly since the early 2000s, when the Bush administration’s NLRB appointments led to a series of decisions that were especially hostile to organized labor (Liebman 2007; Farber 2014) @milkman2017labor.
 
@@
 The first two data sets, `union_wages` and `union_states` are gathered from Tidy Tuesday, although the original data comes from the Union Membership and Coverage Database at Georgia State University.
-The  tables relevant to this paper detail union membership counts by state and year, as well as union and non union wages nationally.
+@Uniondata The  tables relevant to this paper detail union membership counts by state and year, as well as union and non union wages nationally.
 The third data set, `house`, is also from Tidy Tuesday, although the original data comes from the MIT Election Data and Science Lab.
-The data includes every person who ran for a seat in the House of Representatives, but we only selected the election winner.
+@Housedata The data includes every person who ran for a seat in the House of Representatives, but we only selected the election winner.
 The fourth data set, `income_distribution`, is also from Tidy Tuesday, although the original data comes from the U.S.
 Census.
-The data tracks the mean and median income nationally in the United States.
+@Wealthdata The data tracks the mean and median income nationally in the United States.
 The fifth data set, `minimum_wage`, comes from the U.S.
 Department of Labor and tracks minimum wage at the state level and the federal minimum wage.
+@MinWagedata
 
@@
 
-With two cleaned data sets, we conduct our exploratory data analysis using R version `r paste(R.version$major, R.version$minor, sep = ".")` [@rcoreteam23].
-We rely on several elements of the `tidyverse` [@wickhamtidyverse].
-
-We cite our data. We describe the methods we use to answer our question and to test our hypotheses.
-
-"Our outcome is stopping distance measured in feet." We state our key predictor and how it is measured: "Our key predictor is the car's speed, measured in miles per hour."
-
-We almost never refer to specific variable, object, function, or data frame names (such as `var_x`, `ourdata`, `this_useful_func`, or `df`).
-These particular names are almost never of interest or use to the reader.
-
+\clearpage
 ## Union Data
@@
 ```
+@fig-uniondata shows what is commonly known about collective bargaining agreements in the United States: they have been on the decline since at least the 1980s.
+Unfortunately our data does not extend to the last four years where some reports indicate union membership is on the rise.
+Early on, we can see that southern and more Republican leaning states have lower collective bargaining rates than more Democrat and northern / western states.
+There is a very small bump in membership in 2008, which we hypothesize happened in relation to the 2008 financial crisis.
+
+\clearpage
 
@@
 
+\clearpage
+
 ```{r fig-incomedataadj}
@@
 ```
+
+@fig-incomedata and @fig-incomedataadj show that unadjusted income increases, but when accounting for inflation, Americans are earning less per year now than they did 4 decades ago.
+As we will see later on when examining wage data at a national level, there are numerous external reasons for this.
+Most notably, an aging population that is no longer working is likely driving the average down.
+It is possible that the average worker is earning more but we do not have the data to verify this.
+
+\clearpage
 
@@
 
+@fig-wagedata removes all data points where the state effective minimum wage falls below the federal amount.
+From this we can see that more states are taking it into their own hands to protect workers rights.
+We see a large jump in 2007 which may be related to the financial crisis.
+There is another spike around 2014.
+
+\clearpage
+
 ## House of Representatives Data
@@
 
+In @fig-housedata we see that, especially in the northern-central states, the country has grown to be more Republican in the last 4 decades.
+However, in the last 10 years this has begun to shift again.
+During this whole time, it appears that New England has become more Democratic.
+It is important to note that the states in the north-central region typically have only one to three representatives which can make changes appeaer much more drastic than they actually are.
+
+\clearpage
+
 ## Collective Bargaining and Wage Data
@@
 
+@fig-unionwagedata is rather unintuitive, as we would expect adjusted wages to be decreasing in this time, based on @fig-incomedataadj.
+There are a number of potential causes for this.
+First, Americans work less hours today than they did in 1980.
+Second, and more likely, is that the income variables we used includes those who are not working.
+As the U.S.
+experiences a graying of society, this could be driving down mean incomes.
+
+\clearpage
+
 ```{r fig-collectivebargain}
@@
 ```
+
+@fig-collectivebargain is incredibly interesting.
+First, a quick view of this graph shows the exact opposite of our hypothesis.
+Looking at each decade, we see collective bargaining rates decreasing and wage increasing.
+Then, in each decade that relationship continues.
+Still, we can see that in every decade, a collective bargaining agreement is associated with higher wages than for a worker over non-collective-bargaining-jobs.
+However, by the 2010s, it appears that the distance between union and non union wages are shrinking.
+Still, this finding should be taken in the context that the graph does not account for the number of workers in each category and other confounding variables.
+
+\clearpage
 
@@
 
+@fig-collectivebargaindecade demonstrates the political association with collective bargaining agreements over the last 4 decades In the 1980s, there was almost no relationship between collective bargaining agreements and party affiliation, as seen by the horizontal regression line.
+However, by the 2010s, there is a significant increase in slope.
+
+\clearpage
+
 ```{r fig-collectivebargainregion}
@@
 
+In @fig-collectivebargainregion we can see that the North Central and South regions are the main drivers behind this, seen in the steepening of the regression slope.
+
+\clearpage
+
 # Exploring Correlations: Linear Regression Analyses and T-Tests on Collective Bargaining and Wage Disparities
@@
 
-@tbl-linear performs a linear regression between the percent of workers in a collective bargaining agreement and the proportion of the House delegation that is Democratic for the 1980s and 2010s in each region, as well as for all of the regions combined
+Building on the results in @fig-collectivebargaindecade and @fig-collectivebargainregion, @tbl-linear performs a linear regression between the percent of workers in a collective bargaining agreement and the proportion of the House delegation that is Democratic for the 1980s and 2010s in each region, as well as for all of the regions combined
 
@@
 
+\clearpage
+
 ```{r fig-qqplot2010s}
@@
 
+\clearpage
+
 ## Comparing Wages between Union and Non Union Workers
 
-Our results seen in @tbl-ttest for these t-tests determine if there is a statistically significant difference between union wages and non union wages in the 1980s and the 2010s.
+Building on the results in @fig-collectivebargain, our results seen in @tbl-ttest for these t-tests determine if there is a statistically significant difference between union wages and non union wages in the 1980s and the 2010s.
 
@@
 
--   Union wage data at the state level
--   Data past 2019
--   A greater amount and complexity of models
+There are multiple limitations to our project that we found during our research, that we hope to address in future research.
+First, our regression models could be improved by adding a time series element to the model.
+Furthermore, using a multiple linear regression model instead of multiple SLR models.
+For instance, in an MLR model with a time series element, we could take into account the success of unions in a given time period and state.
+We could also include other socioeconomic factors at the state level such as housing prices and general costs of living.
+This would provide a more accurate image of what trends are occurring in the data.
+Next, data past 2019 would provide useful insights into the union work done by the Biden administration.
+Finally, with more variables we may be able to expose more nuanced relationships.
+Two example of this would be union wages and other income indicators at the state level and election data for the U.S.
+Senate and individual state legislatures.
 
-We remind the reader what this paper was about, why it was important, and what we found.
-We reflect on limitations of the data or methods.
-If we have specific advice for someone picking up where we leave off, we provide that guidance.
-We avoid making trite statements like "more research should be done".
+Overall, in our project, we were able to provide merit to our first hypothesis, showing a growing relationship between collective bargaining agreement rates and house delegation makeup.
+We were not able to substantiate our second hypothesis as some of our graphs showed the opposite of our predictions while others and our t-test did indicate our predictions were correct.
 

### 2024-01-12 Documents/Paper.qmd
paper

@@
 House of Representatives election results, and economic variables including the incomes and wages of Americans.
-These project topics are motivated by the recent surge in union activity, increased political polarization, and a rise in wealth inequality.
+The recent surge in union activity, increased political polarization, and a rise in wealth inequality motivate these project topics.
 
@@
 
-Guided by our research questions, we have two hypothesis:
+Our research questions guide our two hypotheses:
 
 1.    There is a significant correlation between collective bargaining rates and higher amounts of Democrat representation in the House of Representatives. 
-2.    Unions will lead to higher wages.
+2.    Unions lead to higher wages.
 
@@
 
-When conducting a regression analysis, it demonstrates that the correlation between collective bargaining and political representation in the House has gotten stronger in the last four decades from 1984 to 2019.
-Additionally, when conducting t-tests among union and non union wages, there is a statistically significant difference that demonstrates an increasing wage differential over the same time period.
+When conducting a regression analysis, it demonstrates that the correlation between collective bargaining and Democratic representation in the House has gotten stronger in the last four decades from 1984 to 2019.
+Additionally, when conducting t-tests among union and non union wages, there is a statistically significant difference that demonstrates a decreasing wage differential over the same time period.
 
@@
 
-Analyzing Union data sets along with US House of Representative Data allowed us to examine and search for correlation between wages and the increase/decrease of Union participation in national regions.
+Analyzing union data sets along with U.S.
+House of Representative data allows us to examine and search for correlation between wages and the increase/decrease of Union participation in national regions.
 We investigated decreases in wages and union participation, along with widening gaps in income between non union, and union members, and the overall increases in wages.
@@
 Our graphs convey, and affirm @wilmers2017labor, reference point that “even in a period of labor weakness (2000 -2014), union activism increases the wages of union members.
-Utilizing the House of Representative data set of both Democrats and Republicans we also were able to find correlation between the downward slope in union wages and enrollment during a Republican presidency which also came with a large republican majority in the house during the Bush administration.
+Utilizing the House of Representative data set of both Democrats and Republicans we also were able to find correlation between the downward slope in union wages and enrollment during a Republican presidency which also came with a large republican majority in the House during the Bush administration.
 Particularly since the early 2000s, when the Bush administration’s NLRB appointments led to a series of decisions that were especially hostile to organized labor (Liebman 2007; Farber 2014) @milkman2017labor.
@@
 The data being analyzed is from a collection of five data sets.
-The first two data sets, `union_wages` and `union_states` are gathered from Tidy Tuesday, although the original data comes from the Union Membership and Coverage Database at Georgia State University.
-@Uniondata The  tables relevant to this paper detail union membership counts by state and year, as well as union and non union wages nationally.
-The third data set, `house`, is also from Tidy Tuesday, although the original data comes from the MIT Election Data and Science Lab.
-@Housedata The data includes every person who ran for a seat in the House of Representatives, but we only selected the election winner.
+The first two data sets, `union_wages` and `union_states` are gathered from Tidy Tuesday, although the original data comes from the Union Membership and Coverage Database at Georgia State University @Uniondata.
+The  tables relevant to this paper detail union membership counts by state and year, as well as union and non union wages nationally.
+The third data set, `house`, is also from Tidy Tuesday, although the original data comes from the MIT Election Data and Science Lab @Housedata.
+The data includes every person who ran for a seat in the House of Representatives, but we only selected the election winner.
 The fourth data set, `income_distribution`, is also from Tidy Tuesday, although the original data comes from the U.S.
-Census.
-@Wealthdata The data tracks the mean and median income nationally in the United States.
+Census @Wealthdata.
+The data tracks the mean and median income nationally in the United States.
 The fifth data set, `minimum_wage`, comes from the U.S.
-Department of Labor and tracks minimum wage at the state level and the federal minimum wage.
-@MinWagedata
+Department of Labor and tracks minimum wage at the state level and the federal minimum wage @MinWagedata.
 
 All of the data sets undergo basic cleaning so that the columns are correctly formatted, variable names are consistent, so there are no NA values, and relevant variables are selected.
-For the house election data, the table is manipulated so that only rows of candidates who won their election remain.
+For the House election data, the table is manipulated so that only rows of candidates who won their election remain.
 A binary variable is created in which we code the outcome as 1 if the winner was a Democrat and code the outcome as 0 if the winner was a Republican.
@@
     x = "Year",
-    y = "Income in USD (Adjusted to 2020 values)"
+    y = "Income in USD (Adjusted to 2020 Values)"
   )
@@
        x = "Year",
-       y = "Effective Minimum Wage (Adjusted to 2020 values)",
+       y = "Effective Minimum Wage (Adjusted to 2020 Values)",
        color = "Selected States")
@@
     x = "Year",
-    y = "Wages (Adjusted to 2020 values)"
+    y = "Wages (Adjusted to 2020 Values)"
   )
@@
     x = "Percent of Workers Under Collective Bargaining Agreement",
-    y = "Wages (Adjusted to 2020 values)"
+    y = "Wages (Adjusted to 2020 Values)"
   )
@@
 The relationship between unions and wages/income in the United States is much harder to draw inferences.
-While our t-tests and some charts show that unions can help workers earn more, other graphs show that unions have a negative effect on incomes.
+While our t-tests and some charts show that unions can help workers earn more, other graphs show that unions have a negative relationship on incomes.
 Our second hypothesis thus remains without strong support.
@@
 
-Overall, in our project, we were able to provide merit to our first hypothesis, showing a growing relationship between collective bargaining agreement rates and house delegation makeup.
+Overall, in our project, we were able to provide merit to our first hypothesis, showing a growing relationship between collective bargaining agreement rates and House delegation makeup.
 We were not able to substantiate our second hypothesis as some of our graphs showed the opposite of our predictions while others and our t-test did indicate our predictions were correct.

### 2024-01-12 Documents/Paper.qmd
DONE!

@@
 Analyzing union data sets along with U.S.
-House of Representative data allows us to examine and search for correlation between wages and the increase/decrease of Union participation in national regions.
+House of Representative data allows us to examine and search for correlation between wages and the increase/decrease of union participation in national regions.
 We investigated decreases in wages and union participation, along with widening gaps in income between non union, and union members, and the overall increases in wages.
-Our study found strong correlation with @western2011unions.
-Unions, norms and the rise in the U.S.
-wage inequality which found “The decline of organized labor in the United States coincided with a large increase in wage inequality.
-From 1973 to 2007.”
+Our study's findings are in line with @western2011unions.
+*Unions, Norms and the Rise in U.S.
+Wage Inequality* found “The decline of organized labor in the United States coincided with a large increase in wage inequality" from 1973 to 2007.
 
-Our graphs convey, and affirm @wilmers2017labor, reference point that “even in a period of labor weakness (2000 -2014), union activism increases the wages of union members.
-Utilizing the House of Representative data set of both Democrats and Republicans we also were able to find correlation between the downward slope in union wages and enrollment during a Republican presidency which also came with a large republican majority in the House during the Bush administration.
-Particularly since the early 2000s, when the Bush administration’s NLRB appointments led to a series of decisions that were especially hostile to organized labor (Liebman 2007; Farber 2014) @milkman2017labor.
+Our graphs convey, and affirm @wilmers2017labor, reference point that “even in a period of labor weakness (2000-2014), union activism increases the wages of union members.
+Utilizing the House of Representatives data set of both Democrats and Republicans, we also find a correlation between the downward slope in union wages and enrollment during a Republican presidency which also came with a large Republican majority in the House during the Bush administration.
+Particularly since the early 2000s, when the Bush administration’s NLRB appointments led to a series of decisions that were especially hostile to organized labor @milkman2017labor.
 
@@
 The data being analyzed is from a collection of five data sets.
-The first two data sets, `union_wages` and `union_states` are gathered from Tidy Tuesday, although the original data comes from the Union Membership and Coverage Database at Georgia State University @Uniondata.
+The first two data sets, `union_wages` and `union_states`, we collect from the source Tidy Tuesday, although the original data comes from the Union Membership and Coverage Database at Georgia State University @Uniondata.
 The  tables relevant to this paper detail union membership counts by state and year, as well as union and non union wages nationally.
-The third data set, `house`, is also from Tidy Tuesday, although the original data comes from the MIT Election Data and Science Lab @Housedata.
-The data includes every person who ran for a seat in the House of Representatives, but we only selected the election winner.
-The fourth data set, `income_distribution`, is also from Tidy Tuesday, although the original data comes from the U.S.
+The third data set, `house`, we also collect from the source Tidy Tuesday, although the original data comes from the MIT Election Data and Science Lab @Housedata.
+The data includes every person who ran for a seat in the U.S.
+House of Representatives, but we only look at the election winner.
+The fourth data set, `income_distribution`, we also collect from the source Tidy Tuesday, although the original data comes from the U.S.
 Census @Wealthdata.
 The data tracks the mean and median income nationally in the United States.
-The fifth data set, `minimum_wage`, comes from the U.S.
-Department of Labor and tracks minimum wage at the state level and the federal minimum wage @MinWagedata.
+The fifth data set, `minimum_wage`, we collect from the U.S.
+Department of Labor and tracks state minimum wage and the federal minimum wage @MinWagedata.
 
-All of the data sets undergo basic cleaning so that the columns are correctly formatted, variable names are consistent, so there are no NA values, and relevant variables are selected.
+All of the data sets undergo basic cleaning so that the columns are correctly formatted, variable names are consistent, there are no NA values, and relevant variables are selected.
 For the House election data, the table is manipulated so that only rows of candidates who won their election remain.
 A binary variable is created in which we code the outcome as 1 if the winner was a Democrat and code the outcome as 0 if the winner was a Republican.
-An additional variable is created to show the proportion of the House delegation that is democratic for every combination of years and states.
-For the union data, it is filtered to not include union sectors, such as construction unions, and instead look at all unions.
-For the income distribution data, race is filtered out because it is outside of the scope of this study.
+We create an additional variable to show the proportion of the House delegation that is Democratic for every combination of years and states.
+For the union data, we filter it to not include union sectors, such as construction unions, and instead look at all unions.
+For the income distribution data, we filter it to not include race because it is outside of the scope of this study.
 
-The five data sets are combined into two data sets: one that stores data at the state level and another that stores data at the national level.
+We combine the five data sets into two data sets: one that stores data at the state level and another that stores data at the national level.
 `full_state_data` joins the `union_states`, `house`, and `minimum_wage` data sets.
@@
 There are `36` observations and each row represents a distinct year.
-Variables regarding union employment are not provided at the national level, only at the state level, so they were summarized and added to the national data set.
+Variables regarding union employment are not provided at the national level, only at the state level, so we summarize and add them to the national data set.
 Economic variables in the national data are not originally adjusted to 2020 values, so we calculate the rate from the state data that contains adjusted and non-adjusted values for federal minimum wage, and then apply that rate to the national data.
@@
 Among the original five data sets, there is data recorded from 1967 to 2022.
-Not all of the data sets cover that time frame, so once joined into two data sets, the data was filtered to only include the years 1984 to 2019.
+Not all of the data sets cover that time frame, so once joined into two data sets, we filter the data to only include the years 1984 to 2019.
 
@@
 Unfortunately our data does not extend to the last four years where some reports indicate union membership is on the rise.
-Early on, we can see that southern and more Republican leaning states have lower collective bargaining rates than more Democrat and northern / western states.
+Early on, we see that southern and more Republican leaning states have lower collective bargaining rates than more Democrat and northern / western states.
 There is a very small bump in membership in 2008, which we hypothesize happened in relation to the 2008 financial crisis.
@@
   geom_line(aes(y = mean_income), color = "red") +
-  annotate("text", x=1995, y=85000, label= "Mean income", color = "red", size = 5) +
+  annotate("text", x=1992, y=85000, label= "Mean income", color = "red", size = 5) +
   geom_line(aes(y = median_income), color = "blue") +
@@
   geom_line(aes(y = median_income_2020_adj), color = "blue") +
-  annotate("text", x=1995, y=115000, label= "Median income", color = "blue", size = 5) +
+  annotate("text", x=1996, y=115000, label= "Median income", color = "blue", size = 5) +
   scale_x_continuous(breaks = seq(min(full_ntl_data$year), max(full_ntl_data$year), by = 4)) + 
@@
 @fig-wagedata removes all data points where the state effective minimum wage falls below the federal amount.
-From this we can see that more states are taking it into their own hands to protect workers rights.
+From this we see that more states are taking it into their own hands to protect workers rights.
 We see a large jump in 2007 which may be related to the financial crisis.
@@
 During this whole time, it appears that New England has become more Democratic.
-It is important to note that the states in the north-central region typically have only one to three representatives which can make changes appeaer much more drastic than they actually are.
+It is important to note that the states in the north-central region typically have only one to three representatives which can make changes appear much more drastic than they actually are.
 
@@
 First, Americans work less hours today than they did in 1980.
-Second, and more likely, is that the income variables we used includes those who are not working.
+Second, and more likely, our income variables include those who do not work.
 As the U.S.
-experiences a graying of society, this could be driving down mean incomes.
+experiences a "graying of society", this could be driving down mean incomes.
 
@@
 Then, in each decade that relationship continues.
-Still, we can see that in every decade, a collective bargaining agreement is associated with higher wages than for a worker over non-collective-bargaining-jobs.
+Still, we see that in every decade, a collective bargaining agreement is associated with higher wages than for a worker over non-collective-bargaining-jobs.
 However, by the 2010s, it appears that the distance between union and non union wages are shrinking.
@@
 @fig-collectivebargaindecade demonstrates the political association with collective bargaining agreements over the last 4 decades In the 1980s, there was almost no relationship between collective bargaining agreements and party affiliation, as seen by the horizontal regression line.
-However, by the 2010s, there is a significant increase in slope.
+However, by the 2010s, there is a significant increase in regression slope.
 
@@
 
-Based on our exploratory data analysis, we are conducting linear regression analyses and t-tests.
+Based on our exploratory data analysis, we conduct linear regression analyses and t-tests.
 The linear regressions use data from `full_state_data` and explore how collective bargaining agreements are correlated with Democratic representatives being elected to the House in all four regions.
-The results are used to examine if the north central region and southern region are driving this stronger correlation over time.
+The results are used to examine if the North Central region and Southern region are driving this stronger correlation over time.
 Additionally, the t-test uses data from `full_ntl_data` to understand if the observed variations in wages between unionized and non-unionized workers are statistically meaningful.
@@
 @fig-estimates plots the `Estimate` values calculated in the linear regressions.
-The regression analysis shows that the south and north central regions in the United States are the main regions that have seen union membership rates and political leanings show increased correlation.
-The south and north central regions have the greatest increase in their estimate values from the 1980s to the 2010s among the four regions.
+The regression analysis shows that the South and North Central regions in the United States are the main regions that have seen union membership rates and political leanings show increased correlation.
+The South and North Central regions have the greatest increase in their estimate values from the 1980s to the 2010s among the four regions.
 
@@
 
-From our data analysis it can be inferred that collective bargaining agreements have become a very political concept.
+From our data analysis, it can be inferred that collective bargaining agreements have become a very political concept.
 The charts in our exploratory data analysis and regression models show this.
@@
 There are multiple limitations to our project that we found during our research, that we hope to address in future research.
-First, our regression models could be improved by adding a time series element to the model.
-Furthermore, using a multiple linear regression model instead of multiple SLR models.
-For instance, in an MLR model with a time series element, we could take into account the success of unions in a given time period and state.
-We could also include other socioeconomic factors at the state level such as housing prices and general costs of living.
-This would provide a more accurate image of what trends are occurring in the data.
-Next, data past 2019 would provide useful insights into the union work done by the Biden administration.
+First, our regression models can be improved by adding a time series element to the model.
+Furthermore, we can use a multiple linear regression model instead of multiple SLR models.
+For instance, in an MLR model with a time series element, we can take into account the success of unions in a given time period and state.
+We can also include other socioeconomic factors at the state level such as housing prices and general costs of living.
+This can provide a more accurate image of what trends are occurring in the data.
+Next, data past 2019 can provide useful insights into the union work done by the Biden administration.
 Finally, with more variables we may be able to expose more nuanced relationships.
@@
 
-Overall, in our project, we were able to provide merit to our first hypothesis, showing a growing relationship between collective bargaining agreement rates and House delegation makeup.
+Overall, in our project, we are able to provide merit to our first hypothesis, showing a growing relationship between collective bargaining agreement rates and House delegation makeup.
 We were not able to substantiate our second hypothesis as some of our graphs showed the opposite of our predictions while others and our t-test did indicate our predictions were correct.
